---
title: "CFS trade flows between US states"
output:
  html_document:
    df_print: paged
  pdf_document:
    number_sections: yes
header-includes:
- \usepackage{amsmath}
- \usepackage{amssymb}
- \usepackage{pifont}
- \usepackage{mathpazo}
---

# General information

This code uses CFS plus proportionality to calculate $X_{ij,k}$ for $i\in US\;  \&\;   j \in US$, for all manufacturing sectors (all sectors except services and agriculture).

## Input files

1. `0-Raw_Data/regions.xlsx`
2. `0-Raw_Data/sectors.xlsx`
3. `1-Intermediate_Processed_Data/CFS_Xijk.dta`
4. `1-Intermediate_Processed_Data/country_country_step_`

## Output files

One txt files per year. The root of each file is: `1-Intermediate_Processed_Data/state_cfs_step_`


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
remove(list = ls())
```

```{r packages, warning=FALSE, include=FALSE, message=FALSE, results='hide', cache=FALSE}
# libraries
options( scipen=100, stringsAsFactors=FALSE );
# vector of libraries to be used
libs <- c( 'dplyr', 'tidyr', 'readstata13', 'haven', 'stringr', 'nleqslv', 'gdata', 'Statamarkdown', 'reshape2', 'read_excel', 'readxl');
# Install libraries in case they are not installed
for( i in length(libs)){ 
  if( !(libs[i] %in% installed.packages()) ) install.packages( libs ) 
};
# Carga de bibliotecas
lapply( libs, require, character.only=TRUE );
```

# Manufacturing bilateral trade between US states

## Intermediate variables

Using CFS data we construct: $$x_{ij,k}^{CFS}\equiv\dfrac{X_{ij,k}^{CFS}}{\sum_{h}\sum_{l}X_{hl,k}^{CFS}},$$ where $X_{ij,k}^{CFS}$ is the sales of state $i$ to state $j$ in sector $k$ according to CFS.


##Main variable 

The object of interest is: $$X_{ij,k}=x_{ij,k}^{CFS}X_{US,US,k}^{WIOD},$$ where $X_{US,US,k}^{WIOD}$ is the consumption of the US in sector $k$ that is produced in the US, according to WIOD. Note that proportionality implies that $$\sum_{i} \sum_{j} X_{ij,k}=X_{US,US,k}^{WIOD}$$ for each manufacturing sector. 

## Importing datasets to R

```{r data}
#CFS
CFS.base <- read_dta("1-Intermediate_Processed_Data//CFS_Xijk.dta")
#WIOD
WIOD.base <- c()
for (yr in 2000:2007) {
  wiot <- country.country <- read.csv(paste('1-Intermediate_Processed_Data//country_country_step_', yr, '.txt', sep= ""), header = TRUE, sep = ",", dec = ".")
  WIOD.base <- rbind(WIOD.base, wiot)
}

#General parameters
regions <- read_excel("0-Raw_Data/regions.xlsx")
s.names <- regions %>% filter(status=="state") 
s.names <- s.names$region
n.s <- length(s.names)

sector.file <- read_excel("0-Raw_Data/sectors.xlsx")
n.sec <- length(unique(sector.file$final_sector)) -3

epsilon <- 0.0001
```

## Calculations

```{r calculations}
#x_ijk

for (yr in 2000:2007) {

#################################Bases  
#Aproximation
c=c(2002,2007,2012)
ap=1000
for (i in 1:3) {
  p=yr-c[i]
  if(abs(p)<ap) {
    aprox=p
    ap=abs(p)
  }
}
yr. = yr-aprox

#CFS
CFS <- subset(CFS.base, year==yr.)
CFS <- subset(CFS, select = -year )


#WIOD
WIOD <- subset(WIOD.base, year==yr)
WIOD <- subset(WIOD, select = -year)  
  
  
##########################Construction of intermediate variables    
# Create data frame
df = data.frame(CFS)

# Dimensions of data frame
n_rows = dim(df)[1]
n_columns = dim(df)[2]

# Sectors
n_sectors = length(unique(df$sector))

# Unique sectors
unique_sectors = unique(df$sector)

# Create XijkCFS and xijkCFS
XijkCFS = df[,3:n_columns]
sum_hl_XhlkCFS = numeric(n_sectors)
for(k in 1:n_sectors) {
  sum_hl_XhlkCFS[k] = sum(XijkCFS[df$sector==k,])
}


XijkCFS_1 = c()
denominator = c()
for (i in 1:n.sec) {
  denominator = rbind(denominator,matrix(replicate(n.s*n.s,sum_hl_XhlkCFS[i]), 
                                         nrow = n.s, ncol = n.s))
  XijkCFS_1 = rbind(XijkCFS_1,XijkCFS[df$sector==i,])
}

# Compute xijkCFS
xijkCFS = XijkCFS_1/denominator  

##################################Main variable  
#X_ijk  
# Create data frame
df_WIOD = data.frame(WIOD)

# Ignore non-USA exporters (columns)
df_WIOD = df_WIOD %>% select(importer, sector, USA)

#Filter out non-USA importers
USA_exporter_indices = df_WIOD$importer == "USA"

# Take into account only the first 12 sectors 
X_USA_USA_k = df_WIOD$USA[USA_exporter_indices]
X_USA_USA_k = X_USA_USA_k[1:n.sec]

Xijk = c()
for(i in 1:n.sec) {
  Xijk = rbind(Xijk, xijkCFS[((i-1)*n.s + 1):(n.s*i),]*X_USA_USA_k[i])
}

###########################################Check that the following condition holds:
  
Xijk$sector = df$sector[1:(n.sec*n.s)]
Xijk$importer = df$importer[1:(n.sec*n.s)]
Xijk = Xijk %>% select(importer, sector, everything())

xijkCFS$sector = df$sector[1:(n.sec*n.s)]
xijkCFS$importer = df$importer[1:(n.sec*n.s)]
xijkCFS = xijkCFS %>% select(importer, sector, everything())

# Check for all  k
for(k in 1:n.sec) {
  if(abs(sum(Xijk[Xijk$sector==k,3:52])/X_USA_USA_k[k] -1)>epsilon) 
    stop(paste("states' exports to states != US WIOD exports to US for year", yr))
}

Step2 <- Xijk 

Step2 <- Step2 %>%
  mutate(year=yr) %>%
  select(year, importer, sector, everything())


#Exporting 
colnames(Step2)[4:dim(Step2)[2]]= CFS.base$importer[1:n.s]

  #Exporting
write.table(Step2, file = paste("1-Intermediate_Processed_Data//state_cfs_step_", yr,".txt", sep=""), sep = ",", row.names = FALSE)
}
```

