---
title: "Country-level value-added share"
output:
  html_notebook:
    toc: yes
    df_print: paged
    number_sections: yes    
  pdf_document:
    number_sections: yes
  word_document: default
header-includes:
- \usepackage{amsmath}
- \usepackage{amssymb}
- \usepackage{pifont}
- \usepackage{mathpazo}
---

# General information

This code calculates the share of value-added (mapped in the model to the labor share) for each country, each sector, and each year using data from WIOD. 

## Input files

1. `0-Raw_Data/regions.xlsx`
2. `0-Raw_Data/sectors.xlsx`
3. `1-Intermediate_Processed_Data/wiot_full.dta`

## Output files

One CSV file per year with root: `1-Intermediate_Processed_Data/labor_shares_countries`

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
remove(list = ls())
```

```{r packages, warning=FALSE, include=FALSE, message=FALSE, results='hide', cache=FALSE}
# libraries
options( scipen=100, stringsAsFactors=FALSE );
## vector of libraries to be used
libs <- c( 'dplyr', 'tidyr', 'readstata13', 'haven', 'stringr', 'nleqslv', 'gdata', 'Statamarkdown', 'readxl', 'reshape2');
# Install libraries in case they are not installed
for( i in length(libs)){
  if( !(libs[i] %in% installed.packages()) ) install.packages( libs )
};
# Carga de bibliotecas
lapply( libs, require, character.only=TRUE );
```


# Constructing Value-Added for countries ("Labor Shares")

## Importing datasets to R
```{r data, warning=FALSE}
regions <- read_excel("0-Raw_Data/regions.xlsx")
c.names <- regions %>% filter(status=="country")
c.names <- c.names$region

s.names <- regions %>% filter(status=="state") 
s.names <- s.names$region

n.s <- length(s.names)
n.c <- length(c.names) -1

sector.file <- read_excel("0-Raw_Data/sectors.xlsx")
id.sectors <- sector.file %>% 
  select(final_sector, WIOD_sector) %>%
  distinct(WIOD_sector, .keep_all = TRUE)

#WIOD
wiot.full <- read.dta13("1-Intermediate_Processed_Data//wiot_full.dta")
wiot.full <- wiot.full %>%
  left_join(id.sectors, by=c('row_item'='WIOD_sector')) %>%
  mutate(row_item = final_sector) %>%
  select(-final_sector) %>%
  left_join(id.sectors, by=c('col_item'='WIOD_sector')) %>%
  mutate(col_item = final_sector) %>%
  select(-final_sector) %>%
  arrange(year, row_country, col_country, row_item, col_item) %>%
  group_by(year, row_country, col_country, row_item, col_item) %>% #aggregating repeated r-c sequence
  mutate(value1= sum(value)) %>%
  ungroup() %>%
  distinct(year, row_country, col_country, row_item, col_item, .keep_all = TRUE) %>%
  select(-value) %>%
  dplyr::rename(value=value1)

wiot.full.inter <- wiot.full %>%
  filter(row_item!=0, col_item!=0) #NOT CONSUMPTION
```

## Calculations

Gross output for each region, for each sector

$$R_{i,k}= \sum_j X_{ij,k} \quad \forall j$$

```{r gross-output, warning=FALSE}
R <- wiot.full %>%
  group_by(year, row_country, row_item) %>%
  mutate(val= sum(value)) %>%
  ungroup() %>%
  distinct(year, row_country, row_item, .keep_all = TRUE) %>%
  select(-col_country, -col_item, -value) %>%
  dplyr::rename(sector=row_item, region=row_country) %>%
  arrange(year, sector, region)

```

Calculating value added:

$$VA_{i,k}= R_{i,k}- \sum_s X_{j,sk} $$
```{r value-added, warning=FALSE}
#calculating value added for countries
sum.s.X_jsk <- wiot.full.inter %>%
  group_by(year, col_country, col_item) %>%
  mutate(val= sum(value)) %>%
  ungroup() %>%
  distinct(year, col_country, col_item, .keep_all = TRUE) %>%
  select(-row_country, -row_item, -value) %>%
  dplyr::rename(region=col_country, sector=col_item) %>%
  arrange(year, sector, region)
VA <- R$val - sum.s.X_jsk$val
VA <- cbind(R[1:3], VA)
colnames(VA) <- c("year", "region", "sector", "value_added")

VA<-spread(VA, sector, value_added, fill = NA, convert = TRUE,  drop = TRUE, sep = NULL)
R <-spread(R, sector, val, fill = NA, convert = TRUE,  drop = TRUE, sep = NULL)
```

Calculating the share of value added in gross output:

$$\alpha_{i,k} = \frac{VA_{i,k}}{R_{i,k}}$$
```{r value-added-share, warning=FALSE}
labor_shares <- VA[, 3:dim(VA)[2]] / R[, 3:dim(R)[2]]
labor_shares[labor_shares>1] <- 1
labor_shares[labor_shares<(-1)] <- (-1)
labor_shares <- cbind(VA[, 1:2], labor_shares)

for (yr in 2000:2007) {
  final <- labor_shares %>%
    filter(year==yr)
  write.table(final, file = paste("1-Intermediate_Processed_Data//labor_shares_countries", yr, ".csv", sep=""), sep = ",", row.names = FALSE)
}


```
