---
title: "Country Gravity Services WIOD"
output:
  html_document:
    df_print: paged
    number_sections: yes    
  pdf_document:
    number_sections: yes
header-includes:
- \usepackage{amsmath}
- \usepackage{amssymb}
- \usepackage{pifont}
- \usepackage{mathpazo}
- \newcommand{\xmark}{\ding{55}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning=FALSE, include=FALSE, message=FALSE, results='hide', cache=FALSE}
# libraries
remove(list = ls())
# vector of libraries to be used
libs <- c( 'dplyr', 'tidyr', 'readstata13', 'haven', 'stringr', 'nleqslv', 'gdata', 'Statamarkdown', 'formattable', 'reshape', 'ggplot2', 'viridis', 'RColorBrewer', 'reshape2', 'foreign', 'readxl');
#Installing libraries in case they are not installed
for( i in length(libs)){ 
  if( !(libs[i] %in% installed.packages()) ) install.packages( libs ) 
};
#Loading libraries
lapply( libs, require, character.only=TRUE );
install.packages("https://www.ssc.wisc.edu/~hemken/Stataworkshops/Stata%20and%20R%20Markdown/Statamarkdown_0.3.9.zip", repos=NULL)
  library(Statamarkdown)
```

## Final Objective

To estimate
$$
\ln X_{ij,t}=\lambda_t + \delta_{i}^{o}+\delta_{j}^{d}+\beta_{0}\iota_{ij}+\beta_{1}\ln dist_{ij}+\xi_{ij,t}
$$
where $\lambda_t$ is a time FE, $\delta_{i}^{o}$ is an origin FE, $\delta_{j}^{d}$ is a destination FE, and $\iota_{ij}$ is an indicator variable equal to 1 if $i=j$, and zero otherwise. As usual $X_{ij,t}$ is what country $i$ sells to country $j$ in year $t$. The coefficients of interest are $\beta_0$ and $\beta_1$. Then, construct: $\phi_{ij}=\exp(\hat{\beta}_{0}\iota_{ij}+\hat{\beta}_{1}\ln dist_{ij})$.

We will focus on countries only (we forget about the states but care about USA as a country), and our focus is on sector 13 only.

```{r}
regions <- read_excel("0-Raw_Data/regions.xlsx")
c.names <- regions %>% filter(status=="country", region!="USA")
c.names <- c.names$region

s.names <- regions %>% filter(status=="state") 
s.names <- s.names$region

n.s <- length(s.names)
n.c <- length(c.names)

#WIOD
WIOD <- read_dta("1-Intermediate_Processed_Data\\wiot_full.dta")
WIOD=data.frame(WIOD)
WIOD <- WIOD %>% filter(year>=2000) #since distances are available from that year on  

#Trade distances
data_services <- read.csv("1-Intermediate_Processed_Data\\data_services.csv")
data_services=data.frame(data_services)
data_servicesUS <- data_services %>% filter(year==2000)

#US population in 2010
us_pop <- read_excel("0-Raw_Data\\Fips\\us_states_coordinates_counties.xlsx")
us_pop=data.frame(us_pop)

#Trade distances
data_agriculture <- read.csv("1-Intermediate_Processed_Data//data_agriculture.csv")
data_agriculture=data.frame(data_agriculture)
data_agricultureUS <- data_agriculture %>% filter(year==2000)

#US population in 2010
us_pop <- read_excel("0-Raw_Data//Fips//us_states_coordinates_counties.xlsx")
us_pop=data.frame(us_pop)
```
*	((NAICS 90) Construction (c18); 
  row_item=18
*	(NAICS 42-45) Wholesale and Retail Trade (c19-c21); 
  row_item=19 | row_item=20 | row_item=21 
*	(NAICS 481-488) Transport Services (c23-c26);
  row_item=23 | row_item=24 | row_item=25 | row_item=26 
*	(NAICS 511-518) Information Services (c27);
  row_item=27 
*	(NAICS 521-525) Finance and Insurance (c28); 
  row_item=28 
*	(NAICS 531-533)  Real Estate (c29-c30); 
  row_item=29 | row_item=30  
*	(NAICS 61) Education (c32); 
  row_item=32 
*	(NAICS 621-624) Health Care (c33);
  row_item=33 
*	(NAICS 721-722) Accommodation and Food Services (c22); 
  row_item=22 
*	(NAICS 493, 541, 55, 561, 562, 711-713, 811-814) Other Services (c34).
  row_item=34
  
```{r}

#only sector 13: origin
WIOD1 <- WIOD %>%
  filter(row_item>=18 & row_item<=34 & row_item!=31) %>%
  group_by(year, row_country, col_country) %>%
  mutate(ag_value=sum(value)) %>% 
  ungroup() %>% 
  distinct(year, row_country, col_country, .keep_all=TRUE ) %>% 
  select(-col_item, -value, -row_item) %>%
  mutate(iota=0) %>%
  mutate(iota=replace(iota, row_country==col_country, 1)) %>%
  unite(ori_dest, c("row_country", "col_country"), remove = FALSE) %>%
  mutate(dist=0)

#preparing distances 
data_services1 <- data_services %>%
  unite(ori_dest, c("iso_o", "iso_d"), remove = FALSE) %>%
  mutate(ch_o=nchar(as.character(iso_o))) %>%
  mutate(ch_d=nchar(as.character(iso_d))) %>%
  filter(ch_d<=3 & ch_o<=3) %>%
  select(year, dist, ori_dest)

#add distances (country-country, excluding US)
WIOD2=WIOD1
for (i in 1:dim(data_services1)[1]){  
  WIOD2 <- WIOD2 %>%
    mutate(dist=replace(dist, ori_dest==data_services1$ori_dest[i] & year==data_services1$year[i], data_services1$dist[i]))
}

```
adding US-country distances 
```{r}

#creating weights
us_pop1 <- us_pop %>%
  select(State, Population_2010) %>%
  group_by(State) %>%
  mutate(state_pop=sum(Population_2010)) %>%
  ungroup() %>%
  select(-Population_2010) %>%
  distinct(State, .keep_all=TRUE) %>%
  filter(State!="DC")

data_servicesUS1 <- data_servicesUS %>%
  mutate(ch_d=nchar(as.character(iso_d))) %>%
  filter(ch_d>3) %>%
  distinct(iso_d, .keep_all=TRUE)
us_pop1$State=data_servicesUS1$iso_d
us_pop1 <- us_pop1 %>% mutate(tot_pop=sum(state_pop))
us_pop1 <- us_pop1 %>% mutate(pond=(state_pop/tot_pop)) #pond sum==1

# adding US-country distances
data_servicesUS2 <- data_services %>%
  mutate(ch_o=nchar(as.character(iso_o))) %>%
  mutate(ch_d=nchar(as.character(iso_d))) %>%
  filter(ch_d>3 & ch_o==3) %>% #only countries in origin and only states in destination 
  mutate(pond=0)

data_servicesUS3=data_servicesUS2
for(i in 1:n.s){
  data_servicesUS3 <- data_servicesUS3 %>%
    mutate(pond=replace(pond, iso_d==us_pop1$State[i], us_pop1$pond[i]))
}

data_servicesUS3 <- data_servicesUS3 %>%
  mutate(component = pond*dist) %>%
  group_by(year, iso_o) %>%
  mutate(dist_US=sum(component)) %>%
  ungroup() %>%
  distinct(year, iso_o, .keep_all=TRUE) %>%
  mutate(iso_d="USA") %>%
  unite(name1, c("iso_o", "iso_d"), remove = FALSE) %>%
  unite(name2, c("iso_d", "iso_o"), remove = FALSE) %>%
  select(year, name1, name2, dist_US)

#add distances (US-country)
WIOD3 <- WIOD2
for (i in 1:dim(data_servicesUS3)[1]){  
  WIOD3 <- WIOD3 %>%
    mutate(dist=replace(dist, (ori_dest==data_servicesUS3$name1[i] | ori_dest==data_servicesUS3$name2[i]) & year==data_servicesUS3$year[i], data_servicesUS3$dist_US[i]))
}

```


Specific US_US distance using the following formula:
$$dist\left(US,US\right) = \left(\sum_{r\, \in\, US} \sum_{s \, \in \, US} \left(\dfrac{pop_r}{pop_{US}}\right) \left(\dfrac{pop_s}{pop_{US}}\right) d_{rs}^{-1}\right)^{-1},$$
```{r }
# US-US distance
USserv <- data_services %>%
  mutate(ch_o=nchar(as.character(iso_o))) %>%
  mutate(ch_d=nchar(as.character(iso_d))) %>%
  filter(ch_d>3 & ch_o>3) %>%
  mutate(pond_r=0, pond_s=0) %>%
  mutate(dist=(1/dist))
  

for(i in 1:n.s){
  USserv <- USserv %>%
    mutate(pond_r=replace(pond_r, iso_o==us_pop1$State[i], us_pop1$pond[i])) %>%
    mutate(pond_s=replace(pond_s, iso_d==us_pop1$State[i], us_pop1$pond[i]))
}

USserv.final <- USserv %>%
  mutate(component = pond_r*pond_s*dist) %>%
  group_by(year) %>%
  mutate(dist_US_US=sum(component)) %>%
  ungroup() %>%
  distinct(year, .keep_all=TRUE) %>%
  mutate(dist_US_US=1/dist_US_US) 

USserv.final$iso_o="USA"
USserv.final$iso_d="USA"
USserv <- USserv.final %>%
  unite(name1, c("iso_o", "iso_d"), remove = FALSE) 
  
#adding US_US  
for (i in 1:dim(USserv)[1]){  
  WIOD3 <- WIOD3 %>%
    mutate(dist=replace(dist, ori_dest==USserv$name1[i] & year==USserv$year[i], USserv$dist_US_US[i]))
}
```

Final base for regression
```{r }
#taking out countries not of interest
Final <- WIOD3 %>%
  filter(dist!=0) %>%
  dplyr::rename(X=ag_value) %>%
  arrange(year, row_country) %>%
  dplyr::rename(origin=row_country , destination=col_country)
write.csv(Final, "1-Intermediate_Processed_Data\\Country_Gravity_Services_WIOD.csv")
```

###Stata code

```{stata, collectcode=TRUE}
import delimited "1-Intermediate_Processed_Data\\Country_Gravity_Services_WIOD.csv"


gen log_X=ln(x)

gen log_dist=ln(dist)

drop ori_dest x dist

*ssc install reghdfe

*ssc install ftools

*reghdfe, compile

egen ori=group(origin)
egen des=group(destin)

reghdfe log_X iota log_dist , absorb(year ori des) vce(robust)
reghdfe log_X iota log_dist , absorb(year#ori year#des) vce(robust)


gen b0 = _b[iota]

gen b1 = _b[log_dist]

gen phi = exp((b0*iota) + (b1*log_dist))
```

##Agriculture
We will focus on countries only (we forget about the states but care about USA as a country), and our focus is on sector 14 only.

*	(NAICS 111-115) Agriculture (c1).
  row_item=1
  
```{r}
#only sector 14: origin
WIOD1 <- WIOD %>%
  filter(row_item==1) %>%
  group_by(year, row_country, col_country) %>%
  mutate(ag_value=sum(value)) %>% 
  ungroup() %>% 
  distinct(year, row_country, col_country, .keep_all=TRUE ) %>% 
  select(-col_item, -value, -row_item) %>%
  mutate(iota=0) %>%
  mutate(iota=replace(iota, row_country==col_country, 1)) %>%
  unite(ori_dest, c("row_country", "col_country"), remove = FALSE) %>%
  mutate(dist=0)

#preparing distances 
data_agriculture1 <- data_agriculture %>%
  unite(ori_dest, c("iso_o", "iso_d"), remove = FALSE) %>%
  mutate(ch_o=nchar(as.character(iso_o))) %>%
  mutate(ch_d=nchar(as.character(iso_d))) %>%
  filter(ch_d<=3 & ch_o<=3) %>%
  select(year, dist, ori_dest)

#add distances (country-country, excluding US)
WIOD2=WIOD1
for (i in 1:dim(data_agriculture1)[1]){  
  WIOD2 <- WIOD2 %>%
    mutate(dist=replace(dist, ori_dest==data_agriculture1$ori_dest[i] & year==data_agriculture1$year[i], data_agriculture1$dist[i]))
}

```
adding US-country distances 
```{r}

#creating weights
us_pop1 <- us_pop %>%
  select(State, Population_2010) %>%
  group_by(State) %>%
  mutate(state_pop=sum(Population_2010)) %>%
  ungroup() %>%
  select(-Population_2010) %>%
  distinct(State, .keep_all=TRUE) %>%
  filter(State!="DC")

data_agricultureUS1 <- data_agricultureUS %>%
  mutate(ch_d=nchar(as.character(iso_d))) %>%
  filter(ch_d>3) %>%
  distinct(iso_d, .keep_all=TRUE)
us_pop1$State=data_agricultureUS1$iso_d
us_pop1 <- us_pop1 %>% mutate(tot_pop=sum(state_pop))
us_pop1 <- us_pop1 %>% mutate(pond=(state_pop/tot_pop)) #pond sum==1

# adding US-country distances
data_agricultureUS2 <- data_agriculture %>%
  mutate(ch_o=nchar(as.character(iso_o))) %>%
  mutate(ch_d=nchar(as.character(iso_d))) %>%
  filter(ch_d>3 & ch_o==3) %>% #only countries in origin and only states in destination 
  mutate(pond=0)

data_agricultureUS3=data_agricultureUS2
for(i in 1:n.s){
  data_agricultureUS3 <- data_agricultureUS3 %>%
    mutate(pond=replace(pond, iso_d==us_pop1$State[i], us_pop1$pond[i]))
}

data_agricultureUS3 <- data_agricultureUS3 %>%
  mutate(component = pond*dist) %>%
  group_by(year, iso_o) %>%
  mutate(dist_US=sum(component)) %>%
  ungroup() %>%
  distinct(year, iso_o, .keep_all=TRUE) %>%
  mutate(iso_d="USA") %>%
  unite(name1, c("iso_o", "iso_d"), remove = FALSE) %>%
  unite(name2, c("iso_d", "iso_o"), remove = FALSE) %>%
  select(year, name1, name2, dist_US)

#add distances (US-country)
WIOD3 <- WIOD2
for (i in 1:dim(data_agricultureUS3)[1]){  
  WIOD3 <- WIOD3 %>%
    mutate(dist=replace(dist, (ori_dest==data_agricultureUS3$name1[i] | ori_dest==data_agricultureUS3$name2[i]) & year==data_agricultureUS3$year[i], data_agricultureUS3$dist_US[i]))
}

```


Specific US_US distance using the following formula:
$$dist\left(US,US\right) = \left(\sum_{r\, \in\, US} \sum_{s \, \in \, US} \left(\dfrac{pop_r}{pop_{US}}\right) \left(\dfrac{pop_s}{pop_{US}}\right) d_{rs}^{-1}\right)^{-1},$$
```{r }
# US-US distance
USagri <- data_agriculture %>%
  mutate(ch_o=nchar(as.character(iso_o))) %>%
  mutate(ch_d=nchar(as.character(iso_d))) %>%
  filter(ch_d>3 & ch_o>3) %>%
  mutate(pond_r=0, pond_s=0) %>%
  mutate(dist=(1/dist))
  

for(i in 1:n.s){
  USagri <- USagri %>%
    mutate(pond_r=replace(pond_r, iso_o==us_pop1$State[i], us_pop1$pond[i])) %>%
    mutate(pond_s=replace(pond_s, iso_d==us_pop1$State[i], us_pop1$pond[i]))
}

USagri.final <- USagri %>%
  mutate(component = pond_r*pond_s*dist) %>%
  group_by(year) %>%
  mutate(dist_US_US=sum(component)) %>%
  ungroup() %>%
  distinct(year, .keep_all=TRUE) %>%
  mutate(dist_US_US=1/dist_US_US) 

USagri.final$iso_o="USA"
USagri.final$iso_d="USA"
USagri <- USagri.final %>%
  unite(name1, c("iso_o", "iso_d"), remove = FALSE) 
  
#adding US_US  
for (i in 1:dim(USagri)[1]){  
  WIOD3 <- WIOD3 %>%
    mutate(dist=replace(dist, ori_dest==USagri$name1[i] & year==USagri$year[i], USagri$dist_US_US[i]))
}
```

Final base for regression
```{r }
#taking out countries not of interest
Final <- WIOD3 %>%
  filter(dist!=0) %>%
  dplyr::rename(X=ag_value) %>%
  arrange(year, row_country) %>%
  dplyr::rename(origin=row_country , destination=col_country)
write.csv(Final, "1-Intermediate_Processed_Data//Country_Gravity_Agriculture_WIOD.csv")
```

###Stata code

```{stata, collectcode=TRUE}
import delimited "1-Intermediate_Processed_Data//Country_Gravity_Agriculture_WIOD.csv", clear

gen log_X=ln(x)

gen log_dist=ln(dist)

drop ori_dest x dist

egen ori=group(origin)
egen des=group(destin)

reghdfe log_X iota log_dist , absorb(year ori des) vce(robust)
reghdfe log_X iota log_dist , absorb(year#ori year#des) vce(robust)

```
