---
title: "Gravity for Agriculture: Inputs for the System"
author: "JP"
date: "April 2, 2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##General description

This file generates the inputs of the gravity system for agriculture.

Start with the standard gravity equation,
$$
X_{ij}=\left(\frac{w_{i}\tau_{ij}}{P_{j}}\right)^{-\varepsilon}X_{j}.
$$
Define: $\Pi_{i}^{-\varepsilon}=\sum_{j}\tau_{ij}^{-\varepsilon}P_{j}^{\varepsilon}X_{j}$. Let $\tilde{P}_{j}\equiv P_{j}^{-\varepsilon}$ and $\tilde{\Pi}_{i}\equiv\Pi_{i}^{-\varepsilon}, and \tilde{\tau}_{ij}\equiv\tau_{ij}^{-\varepsilon}$. Then we have the system
$$
\tilde{P}_{j}=\sum_{i}\tilde{\tau}_{ij}\tilde{\Pi}_{i}^{-1}Y_{i} \nonumber
$$

$$
\tilde{\Pi}_{i}=\sum_{j}\tilde{\tau}_{ij}\tilde{P}_{j}^{-1}X_{j}
$$

So far this is the same we had before. Given $\{X_{j}\}$ and $\{Y_{i}\}$ and $\{\tilde{\tau}_{ij}\}$ then from here one can get $\{\tilde{P}_{j}\}$ and $\{\tilde{\Pi}_{i}\}$. Then we get $\{X_{ij}\}$ from 
$$
X_{ij}=\tilde{\tau}_{ij}\tilde{\Pi}_{i}^{-1}\tilde{P}_{j}^{-1}Y_{i}X_{j} \label{eq:Xij}
$$
We don't know $\{X_{j}\}_{j\in US}$. But we know: 1) $\{X_{ij}\}\quad\forall i,j\notin US$ (from WIOD), 2) $\{X_{ij}\}\quad\forall i\notin US\;or\:j\notin US$ (from CENSUS). 

##Data and Libraries
```{r warning=FALSE, include=FALSE, message=FALSE, results='hide', cache=FALSE}
# libraries
remove(list = ls())
options( scipen=100, stringsAsFactors=FALSE );
# vector of libraries to be used
libs <- c( 'dplyr', 'readstata13', 'nleqslv', 'haven', 'tidyverse', 'optimx', 'optimization',
           'optimx', 'tidyr', 'readxl', 'Rcpp', 'DEoptim', 'parallel', 'reshape2');
# Install libraries in case they are not installed
for( i in length(libs)){ 
  if( !(libs[i] %in% installed.packages()) ) install.packages( libs ) 
};
# Carga de bibliotecas
lapply( libs, require, character.only=TRUE );
```

```{r warning=FALSE, include=FALSE, message=FALSE, results='hide', cache=FALSE}
# OPTIONS AND LIBRARIES
options( scipen=100 );
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
#example vs real
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
## real
# number of countries (n.c) and number of states (n.s)
n.c <- 37; 
n.s <- 50; 
WIOD.base <- read.dta13( file="1-Intermediate_Processed_Data//WIOD_countries.dta" )
data_gravity_agriculture.base <- read.csv( file="1-Intermediate_Processed_Data//data_agriculture.csv", header=TRUE, sep="," )
#state level imports and exports
CENSUS.exp <- read.csv("1-Intermediate_Processed_Data//census.exp.agri.proportionality.csv")
CENSUS.imp <- read.csv("1-Intermediate_Processed_Data//census.imp.agri.proportionality.csv")
wiot.full <- read.dta13("1-Intermediate_Processed_Data//wiot_full.dta")
wiot.full=data.frame(wiot.full)
wiot.full <- wiot.full %>% 
  filter(year>=2000, col_country!="LUX", col_country!="LVA", col_country!="MLT", row_country!="LUX", row_country!="LVA", row_country!="MLT") %>% #since distances are available from that year on
  mutate(row_country= ifelse(row_country=="ROM", "ROU", row_country)) %>%
  mutate(col_country= ifelse(col_country=="ROM", "ROU", col_country)) 


# # ##################example  (do not run creation of dist.matrix in section 3.1)
# ####################Select the block of code that you want to comment. Press Ctrl + Shift + C.
# example <- read_excel( "2-Final_Data//ejemplo.xlsx")
# WIOD.base <- WIOD.base [1:3, 1:7]
# WIOD.base$value_AUS[2] <- example$`X_{ij}`[1]
# WIOD.base$value_AUS[3] <- example$`X_{ij}`[2]
# WIOD.base$value_AUT[2] <- example$`X_{ij}`[5]
# WIOD.base$value_AUT[3] <- example$`X_{ij}`[6]
# WIOD.base$value_AUS[1] <- example$`X_{ij}`[3] + example$`X_{ij}`[4]
# WIOD.base$value_AUT[1] <- example$`X_{ij}`[7] + example$`X_{ij}`[8]
# WIOD.base$value_USA[2] <- example$`X_{ij}`[9] + example$`X_{ij}`[13]
# WIOD.base$value_USA[3] <- example$`X_{ij}`[10] + example$`X_{ij}`[14]
# WIOD.base$value_USA[1] <- NA
# WIOD.base <- WIOD.base %>%
#   mutate(sector=14)
# data_gravity_agriculture.base <- data_gravity_agriculture.base %>%
#   filter(year==2000, (iso_o=="Alabama" | iso_o=="Alaska" | iso_o=="AUS" | iso_o=="AUT"), (iso_d=="Alabama" | iso_d=="Alaska" | iso_d=="AUS" | iso_d=="AUT") ) %>%
#   mutate(X_j= ifelse(iso_d=="AUS", example$X_j[1], X_j)) %>%
#   mutate(X_j= ifelse(iso_d=="AUT", example$X_j[2], X_j)) %>%
#   mutate(Y_i= ifelse(iso_o=="AUS", example$Y_i[1], Y_i)) %>%
#   mutate(Y_i= ifelse(iso_o=="AUT", example$Y_i[5], Y_i)) %>%
#   mutate(Y_i= ifelse(iso_o=="Alabama", example$Y_i[9], Y_i)) %>%
#   mutate(Y_i= ifelse(iso_o=="Alaska", example$Y_i[13], Y_i))
# data_gravity_agriculture.base$Y_i <- as.numeric((data_gravity_agriculture.base$Y_i))
# data_gravity_agriculture.base$X_j <- as.numeric(as.character(data_gravity_agriculture.base$X_j))
# dist.matrix <- matrix(c(example$phi_ij[11], example$phi_ij[12], example$phi_ij[9], example$phi_ij[10], example$phi_ij[15], example$phi_ij[16], example$phi_ij[13], example$phi_ij[14], example$phi_ij[3], example$phi_ij[4], example$phi_ij[1], example$phi_ij[2], example$phi_ij[7], example$phi_ij[8], example$phi_ij[5], example$phi_ij[6]), nrow=4, ncol=4)
# dist.matrix <- data.frame(dist.matrix)
# colnames(dist.matrix) <- c("1", "2", "3", "4")
# dist.matrix <- matrix(unlist(dist.matrix), ncol = 4, byrow = TRUE)
# n.c <- 2;
# n.s <- 2;
# yr=2000
# CENSUS.exp <- CENSUS.exp %>%
#   filter(year==2000, (origin=="Alabama" | origin=="Alaska"), (importer=="AUS" | importer=="AUT")) %>%
#   mutate(value= ifelse(origin=="Alabama" & importer=="AUS", example$`X_{ij}`[9], value)) %>%
#   mutate(value= ifelse(origin=="Alabama" & importer=="AUT", example$`X_{ij}`[10], value)) %>%
#   mutate(value= ifelse(origin=="Alaska" & importer=="AUS", example$`X_{ij}`[13], value)) %>%
#   mutate(value= ifelse(origin=="Alaska" & importer=="AUT", example$`X_{ij}`[14], value))
# wiot.full <- wiot.full %>%
#   filter(year==2000, (row_country=="AUS" | row_country=="AUT" | row_country=="USA"), (col_country=="AUS" | col_country=="AUT" | col_country=="USA"), row_item==1, col_item==1) %>%
#   mutate(value= ifelse(row_country=="AUS" & col_country=="AUS", WIOD.base$value_AUS[2], value)) %>%
#   mutate(value= ifelse(row_country=="AUS" & col_country=="AUT", WIOD.base$value_AUS[3], value)) %>%
#   mutate(value= ifelse(row_country=="AUS" & col_country=="USA", WIOD.base$value_AUS[1], value)) %>%
#   mutate(value= ifelse(row_country=="AUT" & col_country=="AUS", WIOD.base$value_AUT[2], value)) %>%
#   mutate(value= ifelse(row_country=="AUT" & col_country=="AUT", WIOD.base$value_AUT[3], value)) %>%
#   mutate(value= ifelse(row_country=="AUT" & col_country=="USA", WIOD.base$value_AUT[1], value)) %>%
#   mutate(value= ifelse(row_country=="USA" & col_country=="AUS", WIOD.base$value_USA[2], value)) %>%
#   mutate(value= ifelse(row_country=="USA" & col_country=="AUT", WIOD.base$value_USA[3], value)) %>%
#   mutate(value= ifelse(row_country=="USA" & col_country=="USA", WIOD.base$value_USA[1], value))
# CENSUS.exp$value <- as.numeric(CENSUS.exp$value)
# # ##############################

#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
```


Looping over years
```{r warning=FALSE}
for (yr in 2000:2007) {
# 1. LOAD DATA
WIOD<-WIOD.base
## WIOD
WIOD <- WIOD.base %>% 
  filter( year==yr, sector==14 ) %>% 
  filter( importer_country!='USA' ) %>% 
  select( -year, -value_USA, -sector, -importer_country, -importer  ) ;
## Gravity data csv format with comma as the separator
data_gravity_agriculture <- data_gravity_agriculture.base %>% 
  filter( year==yr ) %>% 
  select( -year ) %>% 
  mutate( iso_o=as.character(iso_o), iso_d=as.character(iso_d) ) %>% 
  mutate( len_o=ifelse(nchar(iso_o)<4, 'COUNTRY', 'STATE'), 
          len_d=ifelse(nchar(iso_d)<4, 'COUNTRY', 'STATE') );

#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
# 2. NUMBER STATES AND COUNTRIES
# States
id.states <- data_gravity_agriculture %>% 
  filter( len_o=='STATE' ) %>% 
  select( iso_o ) %>% 
  distinct( iso_o ) %>% 
  arrange( iso_o ) %>% 
  mutate( id=row_number() ) %>% 
  dplyr::rename( region=iso_o ) %>% 
  select( id, region );
# Countries
id.countries <- data_gravity_agriculture %>% 
  filter( len_o=='COUNTRY' ) %>% 
  select( iso_o ) %>% 
  distinct( iso_o ) %>% 
  arrange( iso_o ) %>% 
  mutate( id=n.s+row_number() ) %>% 
  dplyr::rename( region=iso_o ) %>% 
  select( id, region );
# Total
id.total <- rbind( id.states, id.countries );
#t(id.total)

#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
# 3. ELEMENTARY MATRICES FOR KRONECKER PRODUCTS

# 3.1. DISTANCE MATRIX
dist.matrix <- data_gravity_agriculture %>% 
  select( iso_o, iso_d, dist ) %>% 
  left_join( y=id.total, by=c('iso_o'='region') ) %>% 
  dplyr::rename( id_o=id ) %>% 
  left_join( y=id.total, by=c('iso_d'='region') ) %>% 
  dplyr::rename( id_d=id ) %>% 
  select( id_o, id_d, dist ) %>% 
  arrange( id_o, id_d ) %>% 
  spread( key=id_d, value=dist );
dist.matrix <- as.matrix( dist.matrix[,2:(n.c+n.s+1)] );

#we use the coefficients from 4-4-Country_Gravity_WIOD.
#own dummy=4.1, Distance elasticity=-1.8
phi<-diag(x = 4.1, n.c+n.s, n.c+n.s);
phi<-exp(phi);
phi<-phi/dist.matrix^1.8; 
dist.matrix<-phi

# 3.2. GDP VECTOR 
gdp.vector <- data_gravity_agriculture %>% 
  select( iso_o, Y_i ) %>% 
  distinct( iso_o, .keep_all=TRUE ) %>% 
  left_join( y=id.total, by=c('iso_o'='region') ) %>% 
  select( id, Y_i ) %>% 
  arrange( id );
gdp.vector <- as.vector( gdp.vector$Y_i );

# 3.3. EXPENDITURE VECTOR 
exp.vector <- data_gravity_agriculture %>% 
  select( iso_d, X_j ) %>% 
  distinct( iso_d, .keep_all=TRUE ) %>% 
  left_join( y=id.total, by=c('iso_d'='region') ) %>% 
  select( id, X_j ) %>% 
  arrange( id );
exp.vector <- as.vector( exp.vector$X_j );

#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
# 4. BIG MATRICES

# 4.1 Matrices X and Y (repeat gdp and expenditure vectors in ns+nc rows)
mat.Y <- matrix( data=rep(gdp.vector, (n.s+n.c)), nrow=(n.s+n.c), byrow=TRUE );
mat.X <- matrix( data=rep(exp.vector, (n.s+n.c)), nrow=(n.s+n.c), byrow=TRUE );

# 4.2 Submatrices
##phi*Y submatrices
phi.Y.ss <- dist.matrix[(1:n.s), (1:n.s)] * mat.Y[(1:n.s), (1:n.s)];
phi.Y.sc <- dist.matrix[(1:n.s), ((n.s+1):(n.s+n.c))] * mat.Y[(1:n.s), ((n.s+1):(n.s+n.c))];
#phi.Y.cs <- dist.matrix[((n.s+1):(n.s+n.c)), (1:n.s)] * mat.Y[((n.s+1):(n.s+n.c)), (1:n.s)];
phi.Y.cc <- dist.matrix[((n.s+1):(n.s+n.c)), ((n.s+1):(n.s+n.c))] * 
  mat.Y[((n.s+1):(n.s+n.c)), ((n.s+1):(n.s+n.c))];
## phi*X submatrices
#phi.X.ss <- dist.matrix[(1:n.s), (1:n.s)] * mat.X[(1:n.s), (1:n.s)];
phi.X.sc <- dist.matrix[(1:n.s), ((n.s+1):(n.s+n.c))] * mat.X[(1:n.s), ((n.s+1):(n.s+n.c))];
#phi.X.cs <- dist.matrix[((n.s+1):(n.s+n.c)), (1:n.s)] * mat.X[((n.s+1):(n.s+n.c)), (1:n.s)];
phi.X.cc <- dist.matrix[((n.s+1):(n.s+n.c)), ((n.s+1):(n.s+n.c))] * 
  mat.X[((n.s+1):(n.s+n.c)), ((n.s+1):(n.s+n.c))];


#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
# 5. WIOD MATRIX AND RELATED QUANTITIES

## 5.1 WIOD matrix (d,o)
wiod.matrix <- as.matrix( WIOD );

## 5.2 lambda vector
### imports and exports of countries to countries
exports.to.countries <- as.matrix( apply(X=wiod.matrix, MARGIN=2, FUN=sum) )

wiot.full.c <- wiot.full %>%
  filter(col_country!="USA", row_country!="USA", row_item==1) %>%
  group_by(year, col_country) %>%
  mutate(consumption= sum(value)) %>%
  ungroup() %>%
  distinct(year, col_country, .keep_all = TRUE) 
wiot.full.c <- wiot.full.c %>%
  filter(year==yr) %>%
  arrange(col_country)
imports.from.countries <- matrix(wiot.full.c$consumption);


### shares of imports and exports of countries
X_c <- as.matrix( exp.vector[(n.s+1):(n.s+n.c)] );
Y_s <- as.matrix( gdp.vector[(1):(n.s)] );
Y_c <- as.matrix( gdp.vector[(n.s+1):(n.s+n.c)] );

state.exports.to.countries <- CENSUS.exp %>%
  filter(year==yr) %>%
  group_by(origin) %>%
  mutate(exp= sum(value)) %>%
  ungroup() %>%
  distinct(origin, .keep_all = TRUE) %>%
  arrange(origin)
state.exports.to.countries <- matrix(state.exports.to.countries$exp)


### lambdas and chis
lambda.cj <- imports.from.countries/ X_c;
chi.si <- state.exports.to.countries/Y_s;
chi.ci <- exports.to.countries/Y_c;
### lambda vector
ones.ns <- matrix( data=rep(1, n.s), nrow=n.s, ncol=1 );
lambda_vector <- rbind( ones.ns, lambda.cj, chi.si, chi.ci );

#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
## 6.1 Big zeros matrices
zeros.nc.ns <- matrix( data=rep(0, n.c*n.s), nrow=n.c, ncol=n.s );
zeros.ns.ns <- matrix( data=rep(0, n.s*n.s), nrow=n.s, ncol=n.s );
zeros.big <- matrix( data=rep(0, (n.s+n.c)*(n.s+n.c)), nrow=(n.s+n.c), ncol=(n.s+n.c) );

## 6.2 GIANT MATRIX 
### Block for system 10
block.10 <- rbind( cbind(phi.Y.ss, phi.Y.sc),
                   cbind(zeros.nc.ns, phi.Y.cc) );
### Block for system 11
block.11 <- rbind( cbind(zeros.ns.ns, phi.X.sc),
                   cbind(zeros.nc.ns, phi.X.cc) );
### Relevant matrix
BIG.MAT <- rbind( cbind(zeros.big, block.10),
                  cbind(block.11, zeros.big) );

### Saving input matrices for the matlab code
write.csv(BIG.MAT, file= paste('1-Intermediate_Processed_Data//agric_mat_B_',yr, '.csv', sep= ""), row.names=FALSE, na="" );
write.csv(lambda_vector, file= paste('1-Intermediate_Processed_Data//agric_vec_lambda_',yr, '.csv', sep= ""), row.names=FALSE, na="" );
}
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

for (yr in 2000:2007) {
## 7. NEW MATH FOR X_j, P_j, and Pi_j only
  
# 1. LOAD DATA
# WIOD<-WIOD.base
# ## WIOD
# WIOD <- WIOD.base %>% 
#   filter( year==yr, sector==14 ) %>% 
#   filter( importer_country!='USA' ) %>% 
#   select( -year, -value_USA, -sector, -importer_country, -importer  ) ;
## Gravity data csv format with comma as the separator
data_gravity_agriculture <- data_gravity_agriculture.base %>% 
  filter( year==yr ) %>% 
  select( -year ) %>% 
  mutate( iso_o=as.character(iso_o), iso_d=as.character(iso_d) ) %>% 
  mutate( len_o=ifelse(nchar(iso_o)<4, 'COUNTRY', 'STATE'), 
          len_d=ifelse(nchar(iso_d)<4, 'COUNTRY', 'STATE') );

#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
# 2. NUMBER STATES AND COUNTRIES
# States
id.states <- data_gravity_agriculture %>% 
  filter( len_o=='STATE' ) %>% 
  select( iso_o ) %>% 
  distinct( iso_o ) %>% 
  arrange( iso_o ) %>% 
  mutate( id=row_number() ) %>% 
  dplyr::rename( region=iso_o ) %>% 
  select( id, region );
# Countries
id.countries <- data_gravity_agriculture %>% 
  filter( len_o=='COUNTRY' ) %>% 
  select( iso_o ) %>% 
  distinct( iso_o ) %>% 
  arrange( iso_o ) %>% 
  mutate( id=n.s+row_number() ) %>% 
  dplyr::rename( region=iso_o ) %>% 
  select( id, region );
# Total
id.total <- rbind( id.states, id.countries );
#t(id.total)

# 3. ELEMENTARY MATRICES FOR KRONECKER PRODUCTS

# 3.1. DISTANCE MATRIX
dist.matrix <- data_gravity_agriculture %>% 
  select( iso_o, iso_d, dist ) %>% 
  left_join( y=id.total, by=c('iso_o'='region') ) %>% 
  dplyr::rename( id_o=id ) %>% 
  left_join( y=id.total, by=c('iso_d'='region') ) %>% 
  dplyr::rename( id_d=id ) %>% 
  select( id_o, id_d, dist ) %>% 
  arrange( id_o, id_d ) %>% 
  spread( key=id_d, value=dist );
dist.matrix <- as.matrix( dist.matrix[,2:(n.c+n.s+1)] );

#we use the coefficients from 4-4-Country_Gravity_WIOD.
#own dummy=4.1, Distance elasticity=-1.8
phi<-diag(x = 4.1, n.c+n.s, n.c+n.s);
phi<-exp(phi);
phi<-phi/dist.matrix^1.8; 
dist.matrix<-phi

# 3.2. GDP VECTOR 
gdp.vector <- data_gravity_agriculture %>% 
  select( iso_o, Y_i ) %>% 
  distinct( iso_o, .keep_all=TRUE ) %>% 
  left_join( y=id.total, by=c('iso_o'='region') ) %>% 
  select( id, Y_i ) %>% 
  arrange( id );
gdp.vector <- matrix( gdp.vector$Y_i );

# 3.3. EXPENDITURE VECTOR 
# exp.vector <- data_gravity_agriculture %>% 
#   select( iso_d, X_j ) %>% 
#   distinct( iso_d, .keep_all=TRUE ) %>% 
#   left_join( y=id.total, by=c('iso_d'='region') ) %>% 
#   select( id, X_j ) %>% 
#   arrange( id );
# exp.vector <- as.vector( exp.vector$X_j );  
  
#state import and export to countries 
state.exports.to.countries <- CENSUS.exp %>%
  filter(year==yr) %>%
  group_by(origin) %>%
  mutate(exp= sum(value)) %>%
  ungroup() %>%
  distinct(origin, .keep_all = TRUE) %>%
  arrange(origin)
state.exports.to.countries <- matrix(state.exports.to.countries$exp)
state.imports.from.countries <- CENSUS.imp %>%
  filter(year==yr) %>%
  group_by(importer) %>%
  mutate(imp= sum(value)) %>%
  ungroup() %>%
  distinct(importer, .keep_all = TRUE) %>%
  arrange(importer)
state.imports.from.countries <- matrix(state.imports.from.countries$imp)
 
X <- matrix( data=rep(1, n.s*3), nrow=n.s*3, ncol=1);
P_j <- matrix(X[(1):(n.s)]) 
Pi_i <- matrix(X[(n.s+1):(n.s*2)])  
X_j <- matrix(X[( (n.s*2)+1 ):(n.s*3)])
Y_i <- as.matrix( gdp.vector[(1):(n.s)] );

#Right hand side
X.inv.new <- matrix(rev(X^(-1)))

X.X <- rbind(X_j, Y_i, X_j)

X.X.inv <- X.X^(-1)

ones.vec <- matrix( data=rep(1, n.s*3), nrow=n.s*3, ncol=1)

forall <- state.imports.from.countries

forall.tilde <- state.exports.to.countries

forall.tilde2 <- Y_i + forall - forall.tilde

big.forall <- rbind(forall, forall.tilde, forall.tilde2)

RHS <- X*( ones.vec- (big.forall*X.X.inv))


#left hand side
zeros.ns.ns <- matrix( data=rep(0, n.s), nrow=n.s, ncol=n.s);

tau.ns.ns <- dist.matrix[(1:n.s), (1:n.s)]

tau.ns.ns.diag <- tau.ns.ns

diag(tau.ns.ns.diag)<- 0

big.tau <- rbind( cbind(zeros.ns.ns, tau.ns.ns, zeros.ns.ns), cbind(zeros.ns.ns, zeros.ns.ns, tau.ns.ns))

ones <- matrix( data=rep(1, n.s), nrow=1, ncol=n.s)
kron.1 <- ( ((P_j^(-1))*X_j) %x% ones )*tau.ns.ns.diag
kron.2 <- ( ((Pi_i^(-1))*Y_i) %x% ones )*((-1)*tau.ns.ns.diag)   
Big.Matrix <- rbind(big.tau, cbind(zeros.ns.ns, kron.1, kron.2))

LHS <- Big.Matrix %*%(X.inv.new*X.X)



}
```
