---
title: "Bilateral Flows Between Countries"
output:
  html_document:
    df_print: paged
  pdf_document:
    number_sections: yes
header-includes:
- \usepackage{amsmath}
- \usepackage{amssymb}
- \usepackage{pifont}
- \usepackage{mathpazo}
- \newcommand{\xmark}{\ding{55}}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
remove(list = ls())
```

```{r warning=FALSE, include=FALSE, message=FALSE, results='hide', cache=FALSE}
# libraries
options( scipen=100, stringsAsFactors=FALSE );
# vector of libraries to be used
libs <- c( 'dplyr', 'tidyr', 'readstata13', 'haven', 'stringr', 'nleqslv', 'gdata', 'Statamarkdown');
# Install libraries in case they are not installed
for( i in length(libs)){ 
if( !(libs[i] %in% installed.packages()) ) install.packages( libs ) 
};
# load library
lapply( libs, require, character.only=TRUE );
```

##Notation
The dataset "final_matrix_2008" has elements $X_{ij,k}^{Base}$. There are three dimensions: $i$ is the exporter region (in the columns), $j$ is the importer region (in the rows, under the variable "importer"), and $k=1,...,14$ represents the sector. Each region are either US States or Countries (countries are written using 3 letters only. States have their full names).

## General Instructions

To use WIOD data to calculate $X_{ij,k}^{2000}$ for $i\notin US\;  \&\;  j \notin US$, and $\forall k=1,...,14$. I call the output of this part $X_{ij,k}^{Step1}$.


## Importing datasets to R
```{r}
#WIOD
WIOD.base <- read_dta("2-Final_Data//WIOD_countries.dta")
```

## Country-Country Step

For the dataset "WIOD", the elements are defined as $X_{ij,k}^{WIOD}$. As usual, there are three dimensions: $i$ is the exporter country (in the columns), $j$ is the importer country (in the rows, under the variable "importer"), and $k=1,...,14$ represents the sector. 

The matrix we care about is $X_{ij,k}^{Step1}=X_{ij,k}^{WIOD} \quad \forall i\notin US\; \& \; j\notin US$. This means that I only need to take out the first row and the variable "value_USA".
```{r}

for (yr in 2000:2007) {
  WIOD <- subset(WIOD.base, year==yr)
  WIOD <- subset(WIOD, select = -year )
  
  X_ijk_step1 <- data.frame(WIOD)
    
  Step1 <- X_ijk_step1 %>%
    select(-value_USA, -importer) %>%
    dplyr::rename(importer=importer_country) %>%
    filter(importer!= 'USA') %>%
    mutate(year=yr) %>%
    select(year, importer, sector, everything())
  
  colnames(Step1)[4:(dim(Step1)[2])]= matrix(WIOD$importer_country[2:38])

  #Exporting
write.table(Step1, file = paste("1-Intermediate_Processed_Data//country_country_step_", yr,".txt", sep=""), sep = ",", row.names = FALSE)
}
```



