## General information

This script does the following:

1) Constructs Revenue and Expenditures data for agriculture by state.

2) Sets the gravity system for state-region agriculture flows and solves it.

## Input files

1. `0-Raw_Data/regions.csv`
2. `0-Raw_Data/sectors.csv`
3. `1-Intermediate_Processed_Data//wiot_full.csv`
4. `1-Intermediate_Processed_Data/state_imports_exports.csv`
5. `1-Intermediate_Processed_Data/country_country_step_.csv`
6. `0-Raw_Data//Agriculture_Census/data_agriculture.csv`
7. `0-Raw_Data//Agriculture_Census//data_fish.csv`
8. `1-Intermediate_Processed_Data/data_services.csv`
9. `1-Intermediate_Processed_Data/state_country_step_y_.csv`
10. `1-Intermediate_Processed_Data/state_country_step_e_.csv`
11. `1-Intermediate_Processed_Data/io_shares.csv`
12. `1-Intermediate_Processed_Data/labor_shares_countries.csv`
13. `1-Intermediate_Processed_Data//state_cfs_step_.csv`
## Output files

1. `1-Intermediate_Processed_Data/data_agriculture.csv`
2. `1-Intermediate_Processed_Data/agric_mat_B_.csv`
3. `1-Intermediate_Processed_Data/agric_vec_lambda_.csv`
4. `1-Intermediate_Processed_Data//vec_agric_solution_.csv`
5. `1-Intermediate_Processed_Data//Xij_matrix_agric_.csv`


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
remove(list = ls())
```

```{r packages, warning=FALSE, include=FALSE, message=FALSE, results='hide', cache=FALSE}
# libraries
set.seed(2023)
options(digits = 15, scipen = 100, stringsAsFactors = FALSE)
# vector of libraries to be used
libs <- c( 'dplyr', 'tidyr', 'readstata13', 'haven', 'stringr', 'nleqslv', 'gdata', 'Statamarkdown', 'formattable', 'ggplot2', 'viridis', 'RColorBrewer', 'reshape2', 'foreign', 'readxl', 'writexl', 'rjson', 'read_excel', 'Matrix', 'geodist', 'pracma');
# Install libraries in case they are not installed
for( i in length(libs)){ 
  if( !(libs[i] %in% installed.packages()) ) install.packages( libs ) 
};
# load libraries
lapply( libs, require, character.only=TRUE );
```

# Revenue and Expenditures data for agriculture
## Importing data

```{r data-importing}
#General parameters (number of countries, n_c, and number of states, n_s).
regions <- read.csv("0-Raw_Data/regions.csv")
regions$region <- as.character(regions$region)
c_names <- regions %>% filter(status == "country", region != "USA")
c_names <- c_names$region

s_names <- regions %>% filter(status == "state") 
s_names <- s_names$region

n_s <- length(s_names)
n_c <- length(c_names)

#sectors and the number of sectors (specifically, the number of sectors minus three to ease referencing sectors 13 and 14: services and agriculture)
reclass_sectors <- read.csv("0-Raw_Data/sectors.csv")
n_sec <- length(unique(reclass_sectors$final_sector)) -3
reclass_sectors <- reclass_sectors %>% 
  dplyr::select(final_sector, wiod_sector) %>%
  distinct(wiod_sector, .keep_all = TRUE)

#States' exports to countries
state_to_country <- c()
for (yr in 2000:2007) {
  temp <- read.csv(paste('1-Intermediate_Processed_Data//state_country_step_y_', yr, '.csv', sep= ""), header = TRUE, sep = ",", dec = ".")
  state_to_country <- rbind(state_to_country, temp)
}
census_exp <- state_to_country %>%
  filter(sector == n_sec + 2) %>%
  dplyr::select(-sector)
census_exp <- melt(census_exp, id.var=c("year", "importer"), variable = "origin")

#States' imports from countries
country_to_state <- c()
for (yr in 2000:2007) {
  temp <- read.csv(paste('1-Intermediate_Processed_Data//state_country_step_e_', yr, '.csv', sep= ""), header = TRUE, sep = ",", dec = ".")
  country_to_state <- rbind(country_to_state, temp)
}
census_imp <- country_to_state %>%
  filter(sector == n_sec + 2) %>%
  dplyr::select(-sector)
census_imp <- melt(census_imp, id.var=c("year", "importer"), variable = "origin")

#WIOD
wiod_base <- c()
for (yr in 2000:2007) {
  wiot <- read.csv(paste('1-Intermediate_Processed_Data//country_country_step_', yr, '.csv', sep= ""), header = TRUE, sep = ",", dec = ".")
  wiod_base <- rbind(wiod_base, wiot)
}

wiod_full <- read.csv("1-Intermediate_Processed_Data//wiot_full.csv")

#CENSUS
census <- read.csv("1-Intermediate_Processed_Data//state_imports_exports.csv")
state_imp_exp_0 <- census %>%
  filter(origin != "AllStates", destination != "AllStates")

#### Agriculture and fish
#Agriculture production (states)
data_agri <- read.csv("0-Raw_Data//Agriculture_Census//data_agriculture.csv")

#Fish production (states)
data_fish <- read.csv("0-Raw_Data//Agriculture_Census//data_fish.csv")

#Trade distances.
#Here, we are taking advantage of the fact that this services data set is already structured including trade distances. We just delete the revenue and expenditure values.
data_agriculture <- read.csv("1-Intermediate_Processed_Data//data_services.csv")
data_agriculture <- data_agriculture %>%
  mutate(R_i=NA, E_j=NA)

#### small (tolerance) value to check between calculations
epsilon <- 0.0001

```


## State's agricultural revenue (by year)

The agricultural production for each state is defined as the sum of: i) exports to other countries; and ii) the total exports to states, which we obtain distributing WIOD US to US exports between states according to the Census Agriculture relative state production (so that the sum of these exports across states equals US to US exports).
```{r}

## Exports to countries
census_exp <- census_exp %>%
  group_by(year, origin) %>%
  mutate(R_i_1 = sum(value)) %>%
  ungroup() %>%
  distinct(year, origin, .keep_all = TRUE) %>%
  arrange(year, origin) %>%
  dplyr::select(-importer, -value) %>%
  dplyr::rename(iso_o = origin)
#Join to distances data set.
data_agriculture <- data_agriculture %>%
  left_join(census_exp, by=c('year'='year', 'iso_o'='iso_o'))

## Exports to states

# sum of agriculture, and fish
data_agri<- merge(data_agri,data_fish, by=c("iso_o", "year"))
data_agri <- data_agri %>% 
  arrange(year) %>%
  mutate(R_i = fish_prod + R_i) %>%
  dplyr::select(-fish_prod)

# calculating relative productions for each state
data_agri <- data_agri %>%
  dplyr::rename(R_i_2 = R_i) %>%
  group_by(year) %>%
  mutate(tot_US = sum(R_i_2)) %>%
  ungroup() %>%
  mutate(rel = R_i_2/tot_US) 

# Assigning Census data to years. Joining relative productions for all states for all years in one data set.

agri <- c() 
# We use Agriculture Census 2002 for years 2000-2004, and Agriculture Census 2007 for years 2005-2007 (we delete years 2008-2011 later).
agri_2002 <- data_agri %>% filter(year == 2002)
agri_2007 <- data_agri %>% filter(year == 2007)
for (yr in 2000:2011) {
  if (yr<2005) {
    num  <- agri_2002
  } else {
    num  <- agri_2007  
  }  
  num    <- num %>%
            dplyr::select(iso_o, rel,R_i_2) %>%
            mutate(year = yr)
  
  agri <- rbind(agri, num)
}
agri$R_i_2  <- NA

# APPLYING PROPORTIONALITY TO WIOD

## WIOD US to US exports
wiod_agri <- wiod_full %>%
  filter(row_country == "USA", col_country == "USA", row_item == (n_sec + 2)) %>%
  group_by(year) %>%
  mutate(production = sum(value)) %>%
  ungroup() %>%
  distinct(year, .keep_all = TRUE) %>%
  dplyr::select(-value, -col_item, -col_country)

## Distributing US to US exports according to the relative importance of each state. 

for (yr in 2000:2011) {
  wiod_value  <- wiod_agri %>%
                 filter(year == yr)
  wiod_value  <- wiod_value$production[1]
  agri        <- agri %>%
                 mutate(R_i_2= ifelse(year == yr, rel*wiod_value, R_i_2))
}
agri <- agri %>% dplyr::select(-rel)

## Final agricultural production for each state
data_agriculture <- data_agriculture %>%
  left_join(agri, by=c('year'='year', 'iso_o'='iso_o'))
data_agriculture <- data_agriculture %>%
  mutate(R_i = R_i_1 + R_i_2) %>%
  dplyr::select(-R_i_1, -R_i_2)

#Check that revenue by states sums up to US exports. 
check_1     <- data_agriculture %>%
               filter(iso_o %in% s_names, year <= 2007) %>%
               distinct(year, iso_o, .keep_all = TRUE) %>%
               group_by(year) %>%
               mutate(total = sum(R_i)) %>%
               ungroup() %>% 
               distinct(year, .keep_all = TRUE) %>%
               dplyr::select(total, year)
check_2     <- wiod_full %>%
               filter(row_country == "USA", row_item == (n_sec + 2), year <= 2007) %>%
               group_by(year) %>%
               mutate(production = sum(value)) %>%
               ungroup() %>%
               distinct(year, .keep_all = TRUE) %>%
               dplyr::select(production, year)
if(abs(sum(check_2$production / check_1$total-1))>epsilon)
   stop(paste("states' production does not add to WIOD for"))


```


## Agricultural revenue and expenditure for countries

```{r prod-exp-countries}
#Agricultural production for countries 
wiod_p <- wiod_full %>%
  filter(row_item == n_sec + 2) %>%
  group_by(year, row_country) %>%
  mutate(production = sum(value)) %>%
  ungroup() %>%
  distinct(year, row_country, .keep_all = TRUE) %>%
  dplyr::rename(iso_o = row_country) %>%
  dplyr::select(year, iso_o, production)
# join with states' production.
data_agriculture <- data_agriculture %>%
  left_join(wiod_p, by=c('year'='year', 'iso_o'='iso_o')) %>%
  mutate(R_i = ifelse(iso_o %in% c_names, production, R_i))
  
#Agricultural consumption for countries 
wiod_c <- wiod_full %>%
  filter(row_item == n_sec + 2) %>%
  group_by(year, col_country) %>%
  mutate(consumption = sum(value)) %>%
  ungroup() %>%
  distinct(year, col_country, .keep_all = TRUE) %>% 
  dplyr::rename(iso_d = col_country) %>%
  dplyr::select(year, iso_d, consumption)
# join with production data-set
data_agriculture <- data_agriculture %>%
  left_join(wiod_c, by=c('year'='year', 'iso_d'='iso_d')) %>%
  mutate(E_j = ifelse(iso_d %in% c_names, consumption, E_j)) %>%
  dplyr::select(-production, -consumption)
  
#saving
data_agriculture <- data_agriculture %>%
  arrange(year, iso_o, iso_d)
write.table(data_agriculture, file = paste("1-Intermediate_Processed_Data//data_agriculture.csv", sep=""), sep = ",", row.names = FALSE)
```


# Gravity system for services

## Theory

Start with the standard gravity equation: 

$$X_{ij}=\left(\frac{w_{i}\tau_{ij}}{P_{j}}\right)^{-\varepsilon}E_{j} \;\; (1)$$

where $P_{j}^{-\varepsilon}=\sum_{i}\left(w_{i}\tau_{ij}\right)^{-\varepsilon}$.
Consider
$$\Pi_{i}^{-\varepsilon}= \sum_{j} \tau_{ij}^{ -\varepsilon}P_{j}^{\varepsilon}E_{j} \;\; (2)$$
Since $\sum_{j}X_{ij}=R_{i}$, (1) implies $\sum_{j}\left(\frac{w_{i}\tau_{ij}}{P_{j}} \right)^{-\varepsilon}E_{j}=R_{i}.$ Then, from (2), we conclude $w_{i}^{-\varepsilon} \Pi_{i}^{-\varepsilon}=R{i}$. Hence we have:

$$X_{ij}=\frac{\tau_{ij}^{-\varepsilon}}{\left(\Pi_{i}P_{j}\right)^{-\varepsilon}}R_{i}E_{j}=\tilde{\tau}_{ij}\tilde{\Pi}_{i}^{-1}\tilde{P}_{j}^{-1}R_{i}E_{j}  \;\; (3)$$

where $\tilde{P}_{j}\equiv P_{j}^{-\varepsilon}$ and $\tilde{\Pi}_{i}\equiv \Pi_{i}^{-\varepsilon}$, and $\tilde{\tau}_{ij}\equiv\tau_{ij}^{-\varepsilon}$.
We want to use (3) to calculate services bilateral flows for states, so we need to calculate  $\tilde{\Pi}_{i}^{-1}$ and $\tilde{P}_{j}$. To do that, first note that if we substitute $w_{i}^{-\varepsilon} \Pi_{i}^{-\varepsilon}=R{i}$ in $P_{j}^{-\varepsilon}=\sum_{i}\left(w_{i}\tau_{ij}\right)^{-\varepsilon}$, we have

$$\tilde{P}_{j}  =  \sum_{i}\tilde{\tau}_{ij}\tilde{\Pi}_{i}^{-1}R_{i} \;\; (4)$$

Equations (2) and (4) give us a system that, once solved, lets us know $\tilde{\Pi}_{i}^{-1}$ and $\tilde{P}_{j}$.
Specifically, using (2) and (4) we get:
$$
\tilde{P}_{j}  =  \sum_{i}\tilde{\tau}_{ij}\tilde{\Pi}_{i}^{-1}R_{i}\quad j\in US  \;\; (5)
$$
$$
\tilde{\Pi}_{i}  = \sum_{j}\tilde{\tau}_{ij}\tilde{P}_{j}^{-1}E_{j}\quad i\in US 
 \;\; (6)
$$
Now, from WIOD and  Import and Export Merchandise Trade Statistics we know $X_{ij}$ for some pairs. Define $S_{j}$ as the set of $i$'s such that we know $X_{ij}$; and define $S^{*}_{i}$ as the set of $j$'s such that we know $X_{ij}$. Letting 

$$X_{S_{j},j}\equiv\sum_{i\in S_{j}}X_{ij}, \quad\lambda_{j}\equiv1-X_{S_{j},j}/E_{j}, \quad X_{i,S_{i}^{*}}=\sum_{j\in S_{i}^{*}}X_{ij}, \quad\lambda_{i}^{*}\equiv1-X_{i,S_{i}^{*}}/R_{i}, $$
we conclude that
$$
\lambda_{j}\tilde{P}_{j} =  \sum_{i \in US}\tilde{\tau}_{ij}\tilde{\Pi}_{i}^{-1}R_{i}\quad j\in US 
\;\; (7)
$$

$$
\lambda_{i}^{*}\tilde{\Pi}_{i}  =  \sum_{j\in US}\tilde{\tau}_{ij}\tilde{P}_{j}^{-1}E_{j}\quad i\in US
\;\; (8)
$$
Note that equations (7) and (8) are analogous to equations (7) and (8) in `4-Gravity_services.Rmd`, with the difference that now $S_{j}$ and $S^{*}_{i}$ only excludes the bilateral flows from states to states. Hence, (7) and (8) are our full system now. We can represent it in matrix notation. 
Define $P_{s}=(\tilde{P}_{1},...,\tilde{P}_{50})'$ for the states, and similarly for $\Pi_{s}$,$\lambda_{s}$, and $\lambda_{s}^{*}$. Define $\lambda=\left(\lambda_{s},\lambda_{s}^{*}\right)'$. Define $S=\left(P_{s},\Pi_{s}\right)'$, and with some abuse of notation $S^{-1}=\left(P_{s}^{-1},\Pi_{s}^{-1}\right)'$ with dimensions $(s+s)\times(1)$. Define $\left(\mathcal{T} R\right)_{ss}$ as follows:

$$ \boldsymbol{\left(\mathcal{T} R\right)_{ss}}=\left(\begin{array}{ccc}
\tilde{\tau}_{s_{1}s_{1}}R_{s_{1}} & \cdots & \tilde{\tau}_{s_{k}s_{1}}R_{s_{k}}\\
\vdots & \ddots & \vdots\\
\tilde{\tau}_{s_{1}s_{k}}R_{s_{1}} & \cdots & \tilde{\tau}_{s_{k}s_{k}}R_{s_{k}}
\end{array}\right), $$

and define $\left(\mathcal{T} E\right)_{ss}$ analogously. 
The full system can be written as:

$$ \lambda\circ S=\left(\begin{array}{cc}
0 & \left(\mathcal{T} R\right)_{ss} \\
\left(\mathcal{T} E\right)_{ss} & 0\\
\end{array}\right)\cdot S^{-1}, $$

(again, remember that we are searching for states-states flows, so we ignore $cs$, $cc$, and $sc$ pairs). Or in a more compact representation:
$$ \lambda\circ S=B\cdot S^{-1}, $$

where $\circ$ is the element-by-element product and $B$ is the big matrix. Given $\left\{ E_{j}\right\}$  and $\left\{ R_{i}\right\}$  and $\left\{ \tilde{\tau}_{ij}\right\}$,  we can get $\left\{ \tilde{P}_{j}\right\}$  and $\left\{ \tilde{\Pi}_{i}\right\}$. The solution for $\left\{ \tilde{P}_{j}, \tilde{\Pi}_{i} \right\}$ is unique up to a constant. This indeterminacy requires a normalization. We thus impose $\tilde{P}_{1}=100$. Then we compute $\left\{ X_{ij}\right\}$  from (3) as $$X_{ij}=\tilde{\tau}_{ij}\tilde{\Pi}_{i}^{-1}\tilde{P}_{j}^{-1}R_{i}E_{j}.$$

However, we need to calculate $E_j$ for states first; as it is explained in **Appendix B.2 Step 4** of the paper, we do so using the following formula
$$
E_{j, A G}=\sum_{s} \tilde{\phi}_{j, A G s} R_{j, s}+\frac{\gamma_{A G}}{1-\gamma_{A G}} \sum_{r \neq A G}\left(E_{j, r}-\sum_{s} \tilde{\phi}_{j, r s} R_{j, s}\right)
$$
where $\gamma_{k} \equiv \frac{F_{u s, k}}{F_{U S}}$, where $F_{j, k}=\sum_{i} X_{i j, k F}$ is the final consumption in region $j$ of sector $k, and F_{j}=\sum_{k} F_{j, k}$; and $\tilde{\phi}_{j, k s}=\phi_{j, k s}\left(1-\phi_{j, s}\right)$, where $\phi_{j, k s}$ is the corresponding input-output coefficient for the US and $\phi_{j, s}$ is the share of value-added in gross production for the US in sector $k$ (i.e., we assume 
common input-output matrix and value-added shares across U.S. states and equal to the  ones of the U.S. as a whole).

## Importing data to R

```{r data, warning=FALSE}
# COEFFIENTS FROM REGRESSIONS
own_dummy <- 4.143
dist_coeff <- -1.745

#R_i, E_j (for agriculture, states need to be calculated), and trade distances for agriculture and services.
data_gravity_agriculture.base <- read.csv(file = "1-Intermediate_Processed_Data//data_agriculture.csv", header = TRUE, sep = ",")
data_gravity_serv.base <- read.csv(file = "1-Intermediate_Processed_Data//data_services.csv", header = TRUE, sep = ",")

wiot_full <- wiod_full

#States' exports to countries
state_to_country <- c()
for (yr in 2000:2007) {
  temp <- read.csv(paste('1-Intermediate_Processed_Data//state_country_step_y_', yr, '.csv', sep= ""), header = TRUE, sep = ",", dec = ".")
  state_to_country <- rbind(state_to_country, temp)
}
census_exp <- state_to_country %>%
  filter(sector == n_sec + 2) %>%
  dplyr::select(-sector)
census_exp <- melt(census_exp, id.var=c("year", "importer"), variable = "origin")

#States' imports from countries
country_to_state <- c()
for (yr in 2000:2007) {
  temp <- read.csv(paste('1-Intermediate_Processed_Data//state_country_step_e_', yr, '.csv', sep= ""), header = TRUE, sep = ",", dec = ".")
  country_to_state <- rbind(country_to_state, temp)
}
census_imp <- country_to_state %>%
  filter(sector == n_sec + 2) %>%
  dplyr::select(-sector)
census_imp <- melt(census_imp, id.var=c("year", "importer"), variable = "origin")

```

## Creating and solving gravity system for agriculture (and calculating $E_j$ for states)

```{r warning=FALSE}

for (yr in 2000:2007) {
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
# 1. LOAD DATA
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
  ## Gravity data set (distances, revenue, and expenditure)
  data_gravity_agriculture <- data_gravity_agriculture.base %>%
    filter(year == yr) %>%
    dplyr::select(-year) %>%
    mutate(iso_o = as.character(iso_o), iso_d = as.character(iso_d)) %>%
    mutate(
      len_o = ifelse(nchar(iso_o) < 4, "COUNTRY", "STATE"),
      len_d = ifelse(nchar(iso_d) < 4, "COUNTRY", "STATE")
    )
  data_gravity_serv <- data_gravity_serv.base %>%
    filter(year == yr) %>%
    dplyr::select(-year)
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
# 2. ID FOR STATES AND COUNTRIES
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  ## States
  id_states <- data_gravity_agriculture %>%
    filter(len_o == "STATE") %>%
    dplyr::select(iso_o) %>%
    distinct(iso_o) %>%
    arrange(iso_o) %>%
    mutate(id = row_number()) %>%
    dplyr::rename(region = iso_o) %>%
    dplyr::select(id, region)
  ## Countries
  id_countries <- data_gravity_agriculture %>%
    filter(len_o == "COUNTRY" & iso_o!= "USA") %>%
    dplyr::select(iso_o) %>%
    distinct(iso_o) %>%
    arrange(iso_o) %>%
    mutate(id = n_s + row_number()) %>%
    dplyr::rename(region = iso_o) %>%
    dplyr::select(id, region)
  ## Total
  id_total <- rbind(id_states, id_countries)
  state_names <- id_total$region[1:n_s]

#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
# 3. PRODUCTION AND EXPENDITURE: services and agriculture
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  # 3.1. PRODUCTION/REVENUE (R_i) VECTOR
  ##For services
  gdp_vector_serv <- data_gravity_serv %>%
    dplyr::select(iso_o, R_i) %>%
    distinct(iso_o, .keep_all = TRUE) %>%
    left_join(y = id_total, by = c("iso_o" = "region")) %>%
    dplyr::select(id, R_i) %>%
    arrange(id)
  gdp_vector_serv <- matrix(gdp_vector_serv$R_i)
  #Only states.
  gdp_vector_serv_s <- data.frame(gdp_vector_serv[1:n_s])
  #Add the corresponding columns (and their names).
  gdp_vector_serv_s$sector <- n_sec + 1
  gdp_vector_serv_s$exporter <- state_names
  colnames(gdp_vector_serv_s) <- c("export", "sector", "exporter")
  ##For agriculture
  gdp_vector_agri <- data_gravity_agriculture %>%
    dplyr::select(iso_o, R_i) %>%
    distinct(iso_o, .keep_all = TRUE) %>%
    left_join(y = id_total, by = c("iso_o" = "region")) %>%
    dplyr::select(id, R_i) %>%
    arrange(id)
  gdp_vector_agri <- matrix(gdp_vector_agri$R_i)
  #Only states
  gdp_vector_agri_s <- data.frame(gdp_vector_agri[1:n_s])
  ##Add the corresponding columns (and their names).
  gdp_vector_agri_s$sector <- n_sec + 2
  gdp_vector_agri_s$exporter <- state_names
  colnames(gdp_vector_agri_s) <- c("export", "sector", "exporter")
  #3.2. EXPENDITURE (E_j) VECTOR
  ##For services
  expenditure_vector_serv <- data_gravity_serv %>%
    dplyr::select(iso_d, E_j) %>%
    distinct(iso_d, .keep_all = TRUE) %>%
    left_join(y = id_total, by = c("iso_d" = "region")) %>%
    dplyr::select(id, E_j) %>%
    arrange(id)
  expenditure_vector_serv <- matrix(expenditure_vector_serv$E_j)
  #Only states
  expenditure_vector_serv_s <- data.frame(expenditure_vector_serv[1:n_s])
  ##Add the corresponding columns (and their names).
  expenditure_vector_serv_s$sector <- n_sec + 1
  expenditure_vector_serv_s$importer <- state_names
  colnames(expenditure_vector_serv_s) <- c("import", "sector", "importer")
  expenditure_vector_serv_s <- expenditure_vector_serv_s %>% dplyr::select(importer, sector, import)

#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
# 4. share of US agriculture final consumption in total US final consumption (\gamma_AG)
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  wiot_full_0 <- wiot_full %>%
    filter(year == yr, col_item == 0 & col_country == "USA")
  ##US total consumption.
  wiot_full_1 <- wiot_full_0 %>%
    group_by(year, col_country) %>%
    mutate(C_us = sum(value)) %>%
    ungroup() %>%
    distinct(year, col_country, .keep_all = TRUE)
  ##US agriculture consumption
  wiot_full_2 <- wiot_full_0 %>%
    filter(row_item == n_sec + 2) %>%
    group_by(year, col_country) %>%
    mutate(C_us_agri = sum(value)) %>%
    ungroup() %>%
    distinct(year, col_country, .keep_all = TRUE)
  ##gamma_AG and gamma_AG/(1-gamma_AG)
  C_us <- wiot_full_1$C_us
  C_us_agri <- wiot_full_2$C_us_agri
  gamma <- C_us_agri / C_us
  frac_gamma <- gamma / (1 - gamma)

#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
# 5. CALCULATING E_j FOR AGRICULTURE
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  
  ## 5.1 Prepare R_i's and \tilde{\phi}_{j, AG s} (to calculate the first term in E_j in the next step).
  io_shares <- read.csv(file = paste("1-Intermediate_Processed_Data//io_shares", yr, ".csv", sep = ""))
  labor_shares <- read.csv(file = paste("1-Intermediate_Processed_Data//labor_shares_countries", yr, ".csv", sep = ""))
  ##\phi_{j, AG s}
  io_0 <- io_shares %>%
    filter(region == "USA", sector == n_sec + 2) %>%
    dplyr::select(-year, -region, -sector)
  ##\phi_{j, s}
  labor_shares_0 <- labor_shares %>%
    filter(region == "USA") %>%
    dplyr::select(-year, -region)
  ##\tilde{\phi}_{j, AG s}
  phi_14 <- io_0 * (1 - labor_shares_0)
  ##bilateral trade between US states
  state_state <- read.csv(paste("1-Intermediate_Processed_Data//state_cfs_step_", yr, ".csv", sep = ""), header = TRUE, sep = ",", dec = ".")
  ##States' exports to countries
  state_country_y <- data.frame(read.csv(paste("1-Intermediate_Processed_Data//state_country_step_y_", yr, ".csv", sep = ""), header = TRUE, sep = ",", dec = ".")) %>% filter(sector <= n_sec)
  ##States' imports from countries
  state_country_e <- data.frame(read.csv(paste("1-Intermediate_Processed_Data//state_country_step_e_", yr, ".csv", sep = ""), header = TRUE, sep = ",", dec = ".")) %>% filter(sector <= n_sec)
  ##States' exports to states and countries.
  state_exports <- rbind(state_state, state_country_y)
  ##Total R_i's per state without agriculture and services.
  exp_v <- c()
  for (sec in 1:n_sec) {
    state_exports_1 <- state_exports %>%
      filter(sector == sec) %>%
      dplyr::select(-year, -importer, -sector)
    export <- matrix(colSums(state_exports_1))
    export <- data.frame(export)
    export$sector <- sec
    export$exporter <- state_names
    exp_v <- rbind(exp_v, export)
  }
  ##TOTAL R_i's PER STATE
  gdp_tot_v <- rbind(exp_v, gdp_vector_serv_s, gdp_vector_agri_s)
  US_sum <- gdp_tot_v %>%
    group_by(sector) %>%
    mutate(total_R = sum(export)) %>%
    ungroup() %>%
    distinct(sector, .keep_all = TRUE) %>%
    dplyr::select(sector, total_R)
  ##Check that WIOD US total R_i's coincide with the sum over states of our R_i's, for each sector.
  US_totals <- wiod_base %>%
    filter(year == yr) %>%
    dplyr::select(year, importer, sector, USA) %>%
    group_by(year, sector) %>%
    mutate(total = sum(USA)) %>%
    ungroup() %>%
    distinct(year, sector, .keep_all = TRUE) %>%
    dplyr::select(sector, total)
  if (sum(abs(US_sum$total_R / US_totals$total - 1)) > epsilon) {
    stop(paste("sum(R_i) over states != US WIOD for year", yr))
  }

  ##5.2 CALCULATING \sum_{s}\tilde{\phi}_{j, AG s}R_{j,s} (SUM OF INTERMEDIATE USES IN AGRICULTURE) FOR EACH STATE
  ##reshaping vector of total R_i's per state to matrix form
  gdp_tot_v_mat <- spread(gdp_tot_v, sector, export, fill = NA, convert = TRUE, drop = TRUE, sep = NULL)
  gdp_tot_v_mat <- t(gdp_tot_v_mat[, -1])
  class(gdp_tot_v_mat) <- "numeric"
  ##product of vector of \tilde{\phi}_{j, AG s}'s and the matrix of R_i's to get the sum of intermediate uses in agriculture.
  sum_14 <- as.matrix(phi_14) %*% gdp_tot_v_mat
  sum_14 <- t(sum_14) 
  
  ## 5.3 second component in E_j
  # E_j for all sectors except agriculture
  state_imports <- cbind(state_state[4:dim(state_state)[2]], state_country_e[4:dim(state_country_e)[2]])
  expenditure_state <- rowSums(state_imports)
  expenditure_state <- cbind(state_state[2:3], expenditure_state)
  colnames(expenditure_state) <- c("importer", "sector", "import")
  expenditure_vector <- rbind(expenditure_state, expenditure_vector_serv_s)
  colnames(expenditure_vector) <- c("importer", "r", "import")
  expenditure_mat <- spread(expenditure_vector, importer, import, fill = NA, convert = FALSE, drop = TRUE, sep = NULL) %>% dplyr::select(-r)
  expenditure_mat <- t(expenditure_mat)
  ## checking that expenditure adds up to WIOD totals
  exp_states <- colSums(expenditure_mat)
  US_totals <- wiod_base %>%
    filter(year == yr, importer == "USA", sector != 14) %>%
    dplyr::select(-year, -importer, -sector)
  US_totals <- rowSums(US_totals)
  ## check
  if (sum(abs(exp_states / US_totals - 1)) > epsilon) {
    stop(paste("sum(E_j) over states (except agric) != US WIOD for year", yr))
  }

  ##\phi_{j, k s}
  io_shares <- read.csv(file = paste("1-Intermediate_Processed_Data//io_shares", yr, ".csv", sep = "")) %>%
    filter(region == "USA") %>%
    dplyr::select(-year, -region, -sector)
  ##\phi_{j, s}
  labor_shares <- read.csv(file = paste("1-Intermediate_Processed_Data//labor_shares_countries", yr, ".csv", sep = "")) %>%
    filter(region == "USA") %>%
    dplyr::select(-year, -region)
  labor_shares <- matrix(data = rep(as.matrix(labor_shares), (n_sec + 2)), nrow = (n_sec + 2), byrow = TRUE)

  ##\tilde{\phi}_{j, k s}
  phi_ks <- io_shares * (1 - labor_shares)

  ##sum \tilde{\phi}_{j, k s}*R_{j,s} over all receiving sectors
  sum_not_14_v <- as.matrix(phi_ks) %*% gdp_tot_v_mat
  sum_not_14_v <- as.data.frame(t(sum_not_14_v)) %>%
    dplyr::select(-assign(paste("V", n_sec + 2, sep = ""), n_sec + 2)) 
  ##Sum (for all sectors excluding agriculture) the difference between the previous sums and the corresponding E_{j,s} 
  sum_not_14_vector <- as.data.frame(rowSums(expenditure_mat - sum_not_14_v))

  ## first term of E_{j,AG} + US total agriculture consumption must be equal to WIOD US expenditure in agriculture
  check1 <- wiod_base %>%
    filter(year == yr, importer == "USA", sector == (n_sec + 2))
  c1 <- check1[4:dim(check1)[2]]
  c2 <- sum(sum_14) + C_us_agri
  if (abs(sum(c1) / c2 - 1) > epsilon) {
   stop(paste("problem with sum.14 year", yr))
  }
  ## second term of E_{j,AG} must coincide with total US consumption in agriculture
  if (abs(C_us_agri / sum(frac_gamma * sum_not_14_vector) - 1) > epsilon) {
    stop(paste("second part of the summ does not add to total consumption year", yr))
  }
  
  ## 5.4 final computation of E_j
  E_j <- sum_14 + (frac_gamma * sum_not_14_vector)
  ## check sum of E_j is consistent with WIOD
  c1 <- sum(c1)
  c2 <- sum(E_j)
  if (abs(c1 / c2 - 1) > epsilon) {
    stop(paste("sum(E_j) over states != US WIOD for year", yr))
  }

#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
# 6. Calculating lambda and B matrix for the system
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

  ##6.1 DISTANCE MATRIX
  dist_matrix <- data_gravity_agriculture %>%
    dplyr::select(iso_o, iso_d, dist) %>%
    filter(iso_d != "USA" & iso_o != "USA") %>%
    left_join(y = id_total, by = c("iso_o" = "region")) %>%
    dplyr::rename(id_o = id) %>%
    left_join(y = id_total, by = c("iso_d" = "region")) %>%
    dplyr::rename(id_d = id) %>%
    dplyr::select(id_o, id_d, dist) %>%
    arrange(id_o, id_d) %>%
    spread(key = id_d, value = dist)
  dist_matrix <- as.matrix(dist_matrix[, 2:(n_c + n_s + 1)])
  
  ##Matrix of taus (we refer to it as "phi" from now on). This step eases the calculation of TE and TR matrices.
  #we use the coefficients from 1-WIOD_VA_shares_io_shares_distances.RMD
  phi <- diag(x = own_dummy, n_c + n_s, n_c + n_s)
  phi <- exp(phi)
  phi <- phi * dist_matrix^dist_coeff
  dist_matrix <- phi

  ## 6.2 lambda vector
  ## imports and exports
  state_imports_from_countries <- spread(census_imp, origin, value, fill = NA, convert = FALSE, drop = TRUE, sep = NULL)
  state_imports_from_countries <- state_imports_from_countries %>%
    filter(year == yr) %>%
    dplyr::select(-year, -importer)
  state_exports_to_countries <- spread(census_exp, importer, value, fill = NA, convert = FALSE, drop = TRUE, sep = NULL)
  state_exports_to_countries <- state_exports_to_countries %>%
    filter(year == yr) %>%
    dplyr::select(-year, -origin)
  ## revenue and expenditure (to calculate shares)
  expenditure_vector_agri_s <- (as.matrix(E_j))
  gdp_vector_agri_s <- (as.matrix(gdp_vector_agri_s$export))
  ## lambdas
  state_imports_from_countries <- state_imports_from_countries / expenditure_vector_agri_s
  lambda_j <- 1 - rowSums(state_imports_from_countries)
  state_exports_to_countries <- state_exports_to_countries / gdp_vector_agri_s
  chi_i <- 1 - rowSums(state_exports_to_countries)
  lambda_vector <- rbind(as.matrix(lambda_j), as.matrix(chi_i))

  ## 6.3 BIG MATRIX B
  ## Matrices for E and R (repeat gdp and expenditure vectors in n_s+n_c rows). This step eases the calculation of TE and TR matrices.
  mat_Y <- matrix(data = rep(gdp_vector_agri_s, (n_s)), nrow = (n_s), byrow = TRUE)
  mat_X <- matrix(data = rep(expenditure_vector_agri_s, (n_s)), nrow = (n_s), byrow = TRUE)
  ##TR sub-matrices
  phi_Y_ss <- dist_matrix[(1:n_s), (1:n_s)] * mat_Y
  ##TE sub-matrices
  phi_X_ss <- dist_matrix[(1:n_s), (1:n_s)] * mat_X
  ## Big zeros matrices
  zeros_big <- matrix(data = rep(0, (n_s) * (n_s)), nrow = (n_s), ncol = (n_s))
  ## Relevant matrix
  BIG_MAT <- rbind(
    cbind(zeros_big, phi_Y_ss),
    cbind(phi_X_ss, zeros_big)
  )

#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
# 7. SOLVING THE SYSTEM
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

  ### Saving input matrices
  write.csv(BIG_MAT, file = paste("1-Intermediate_Processed_Data//agric_mat_B_", yr, ".csv", sep = ""), row.names = FALSE, na = "")
  write.csv(lambda_vector, file = paste("1-Intermediate_Processed_Data//agric_vec_lambda_", yr, ".csv", sep = ""), row.names = FALSE, na = "")

  ## Gravity Systems

# Agriculture gravity

BIG.MAT <- as.matrix(read.csv(paste('1-Intermediate_Processed_Data//agric_mat_B_',yr, '.csv', sep= "")))
lambda_vector <- as.matrix(read.csv(paste('1-Intermediate_Processed_Data//agric_vec_lambda_',yr, '.csv', sep= "")))


# This function solves the equilibrium system using non-linear least squares.
gravity_system <- function(x){
    Bfirst <- ( ((lambda_vector*x) / (BIG.MAT%*%(x^(-1)))) -1)*100
    Bsecond <- x[1] - 100
    B <- as.matrix(rbind(Bfirst, Bsecond))
}

start <- 100*matrix(0.6, 2*n_s, 1)
if (yr > 2000){start <- as.matrix(sol)}
sol <- lsqnonlin(gravity_system, start)
sol <- as.matrix(sol$x)

write.csv(sol, file= paste('1-Intermediate_Processed_Data//vec_agric_solution_',yr, '.csv', sep= ""), row.names=FALSE, na="" );


#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
###FROM SOLUTION TO X_ij
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  SOL <- as.matrix(read.csv(paste("1-Intermediate_Processed_Data//vec_agric_solution_", yr, ".csv", sep = "")))
  SOL <- (as.matrix(SOL))[1:(n_s * 2)]
  ## checking if the solver worked.
  value <- lambda_vector * SOL - BIG.MAT %*% SOL^{-1}
  if (norm(value, "2") > epsilon) {
    stop(paste("Solution does not seem to work for year", yr))}
  x <- SOL
  
  ### Basic vector
  P.s <- as.matrix(x[1:n_s])
  Pi.s <- as.matrix(x[(n_s + 1):(2 * n_s)])

  ### Inverted basic
  P.s.inv <- P.s^(-1)
  Pi.s.inv <- Pi.s^(-1)

  ##dist matrix
  dist_matrix <- dist_matrix[(1:n_s), (1:n_s)]

  ##Pi^(-1)*R and P^(-1)*E
  E_j <- as.matrix(E_j)
  Pi.sq.mat <- matrix(
    data = rep(Pi.s.inv * gdp_vector_agri_s, (n_s)),
    nrow = (n_s), ncol = (n_s), byrow = TRUE
  )
  P.sq.mat <- matrix(
    data = rep(P.s.inv * E_j, (n_s)),
    nrow = (n_s), ncol = (n_s), byrow = FALSE
  )
  
  ##X_ij=tau*Pi^(-1)P^(-1)*R*E
  Xij.matrix <- Pi.sq.mat * dist_matrix * P.sq.mat
  write.csv(Xij.matrix, file = paste("1-Intermediate_Processed_Data//Xij_matrix_agric_", yr, ".csv", sep = ""), row.names = FALSE, na = "")

#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
###CHECKS for X_ij
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  #CHECK: US consumption from US =US consumption from US from WIOD
  value1 <- sum(Xij.matrix[(1:n_s), (1:n_s)])
  value2 <- wiod_base %>%
    filter(sector == (n_sec + 2), year == yr, importer == "USA") %>%
    dplyr::select(-importer, -sector, -year)
  value2 <- value2$USA
  # checking
  if (abs(value1 / value2 - 1) > epsilon) {
    stop(paste(" US consumption from US != value from WIOD for", yr, " value=", abs(value1 / value2 - 1)))
  }

}
```
