---
title: "6-3-IO-Tables"
output:
  html_document:
    df_print: paged
  pdf_document:
    number_sections: yes
  word_document: default
header-includes:
- \usepackage{amsmath}
- \usepackage{amssymb}
- \usepackage{pifont}
- \usepackage{mathpazo}
- \newcommand{\xmark}{\ding{55}}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
remove(list = ls())
```

```{r warning=FALSE, include=FALSE, message=FALSE, results='hide', cache=FALSE}
# libraries
options( scipen=100, stringsAsFactors=FALSE );
# vector of libraries to be used
libs <- c( 'dplyr', 'tidyr', 'readstata13', 'haven', 'stringr', 'nleqslv', 'gdata', 'Statamarkdown', 'readxl', 'foreign');
# Install libraries in case they are not installed
for( i in length(libs)){ 
  if( !(libs[i] %in% installed.packages()) ) install.packages( libs ) 
};
# Carga de bibliotecas
lapply( libs, require, character.only=TRUE );
```

# General Instructions

This code calculates the input output trade shares for each sector, country and year. Specifically, a share for year, t, country, c, and sector, k, represents the proportion of c's sector k imports from all countries'sector $k^I$, of c's sector k total imports from all countries. The final output of this code, for each year separately, is a (14*38)x(14+2) matrix, where the first two columns indicate the year and country, and the rest the sectors associated with the shares of the country.    

Each share is defined as follows:
$$ s_{k^I,k,c,t}= \frac{ M_{k^I,k,c,t}^{allcountries}}{\sum_{k^I} M_{k^I,k,c,t}^{allcountries}}$$

## Importing datasets to R
```{r}
#WIOD Full
wiot.full.inter <- read.dta13("1-Intermediate_Processed_Data//wiot_full.dta")
#Transforming WIOD sectors to 1-14 CDP sectors
wiot.full.inter <- wiot.full.inter %>%
  mutate(col_country= recode(col_country, "ROM"="ROU")) %>%
  mutate(row_country= recode(row_country, "ROM"="ROU")) %>%
  filter(col_country != "LVA" & col_country != "LUX" & col_country != "MLT") %>%
  filter(row_country != "LVA" & row_country != "LUX" & row_country != "MLT") %>%
  mutate(row_item= recode(row_item, '1'=14L, '2'=4L, '3'=1L, '4'=2L, '5'=2L, '6'=3L, '7'=3L, '8'=4L, '9'=5L, '10'=6L, '11'=7L, '12'=8L, '13'=9L, '14'=10L, '15'=11L, '16'=12L, '17'=0L, '18'=13L, '19'=13L, '20'=13L, '21'=13L, '22'=13L, '23'=13L, '24'=13L, '25'=13L, '26'=13L, '27'=13L, '28'=13L, '29'=13L, '30'=13L, '31'=0L, '32'=13L, '33'=13L, '34'=13L, '35'=0L, '36'=0L, '37'=0L, '38'=0L, '39'=0L, '40'=0L, '41'=0L, '42'=0L )) %>%
  mutate(col_item= recode(col_item, '1'=14L, '2'=4L, '3'=1L, '4'=2L, '5'=2L, '6'=3L, '7'=3L, '8'=4L, '9'=5L, '10'=6L, '11'=7L, '12'=8L, '13'=9L, '14'=10L, '15'=11L, '16'=12L, '17'=0L, '18'=13L, '19'=13L, '20'=13L, '21'=13L, '22'=13L, '23'=13L, '24'=13L, '25'=13L, '26'=13L, '27'=13L, '28'=13L, '29'=13L, '30'=13L, '31'=0L, '32'=13L, '33'=13L, '34'=13L, '35'=0L, '36'=0L, '37'=0L, '38'=0L, '39'=0L, '40'=0L, '41'=0L, '42'=0L )) %>%
  filter(col_item!=0 & row_item!=0) %>%
  arrange(year, row_country, col_country, row_item, col_item) %>%
  group_by(year, row_country, col_country, row_item, col_item) %>% #aggregating repeated r-c sequence
  mutate(value1= sum(value)) %>%
  ungroup() %>%
  distinct(year, row_country, col_country, row_item, col_item, .keep_all = TRUE) %>%
  select(-value) %>%
  dplyr::rename(value=value1)

```

#Calculating input-output matrices
```{r}
num <- wiot.full.inter %>%
  group_by(year, col_country, col_item, row_item) %>%
  mutate(val= sum(value)) %>%
  ungroup() %>%
  distinct(year, col_country, col_item, row_item, .keep_all = TRUE) %>%
  select(-row_country, -value) %>%
  dplyr::rename(region=col_country, sector_o=row_item, sector_d=col_item)

num <-spread(num, sector_o, val, fill = NA, convert = TRUE,  drop = TRUE, sep = NULL)

io.shares <- num[, 4:dim(num)[2]] / rowSums(num[, 4:dim(num)[2]])
io.shares <- cbind(num[, 1:3], io.shares)

for (yr in 2000:2007) {
  final <- io.shares %>%
    filter(year==yr)
  write.table(final, file = paste("1-Intermediate_Processed_Data//io_shares", yr, ".csv", sep=""), sep = ",", row.names = FALSE)
}


```

