---
title: "7-Employment_migration"
output:
  html_notebook:
    toc: yes
    df_print: paged
    number_sections: yes    
  pdf_document:
    number_sections: yes
header-includes:
- \usepackage{amsmath}
- \usepackage{amssymb}
- \usepackage{pifont}
- \usepackage{mathpazo}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
remove(list = ls())
```

```{r warning=FALSE, include=FALSE, message=FALSE, results='hide', cache=FALSE}
# libraries
options( scipen=100, stringsAsFactors=FALSE );
# vector of libraries to be used
libs <- c( 'dplyr', 'tidyr', 'readstata13', 'haven', 'stringr', 'nleqslv', 'gdata', 'Statamarkdown', 'formattable', 'ggplot2', 'viridis', 'RColorBrewer', 'reshape2', 'foreign', 'readxl', 'writexl', 'rjson', 'read_excel', 'Matrix', 'geodist');
# Install libraries in case they are not installed
#for( i in length(libs)){ 
  #if( !(libs[i] %in% installed.packages()) ) install.packages( libs ) 
#};
# Carga de bibliotecas
lapply( libs, require, character.only=TRUE );
```

## General Instructions

1) This code computes that employment and unemployment level for each state and sector in year 2000.
2) This code computes the mobility matrices for each state and sector for years 1999-2007. 

## Input files

1. `0-Raw_Data/CENSUS_2000/PUMS5_.txt`
2. `0-Raw_Data/CENSUS_2000/employment_2000.xls`
3. `0-Raw_Data/sectors.csv`
4. `0-Raw_Data/FIPS/states_fips_num.xlsx`
5. `0-Raw_Data/ACS/`yr`/ss`yr``state`.csv`
6. `0-Raw_Data/CPS/`month``yr`pub_.txt`

## Output files

1. `1-Intermediate_Processed_Data//L_2000.csv`
2. `1-Intermediate_Processed_Data//mu_`yr`.csv`


## Importing general datasets to R
```{r}
#General
`%notin%` <- Negate(`%in%`)

regions <- read.csv("0-Raw_Data/regions.csv")
c_names <- regions %>% filter(status == "country")
c_names <- c_names$region

s_names <- regions %>% filter(status == "state") 
s_names <- s_names$region

n_s <- length(s_names)
n_c <- length(c_names) -1

#sector reclassification table
reclass_sectors <- read.csv(paste("0-Raw_Data/sectors.csv", sep=""), header = TRUE, sep = ",", dec = ".")
reclass_sectors <- reclass_sectors %>% 
  select(final_sector, naics) %>%
  mutate(naics = as.character(naics))

#state codes
states_fips <- read_excel("0-Raw_Data/Fips/states_fips_num.xlsx", range = "A1:D53")
states_fips <- states_fips %>% select(st_fips, st_num, st_name)

```


# Task 1

```{r}
#employment statistics 2000, by state
employment_pop <- read_excel("0-Raw_Data/CENSUS_2000/employment_2000.xls", skip = 12)
employment_pop <- employment_pop %>%
  dplyr::rename(state = ...1, employed = Employed, unemployed = Unemployed) %>%
  filter(is.na(state) == FALSE) %>%
  mutate(state = gsub(" ", "", state)) %>% #taking out spaces
  filter(state %in% s_names) %>%
  select(state, employed, unemployed)


#loading Census PUMS
list <- c("01", "02", "04", "05", "06", "08", "09", setdiff(10:56, c(11,14,43,52)))
census <- c()
c <- 1
for (i in list){
CENSUS_PUMS <- read.csv(paste("0-Raw_Data/CENSUS_2000/PUMS5_", i,".txt", sep=""), header=FALSE, sep=";")
CENSUS_PUMS$state <- s_names[c]
census <- rbind(census, CENSUS_PUMS)
print(s_names[c])
c <- c + 1
}
rm(CENSUS_PUMS)

#creating variables
census <- census %>%
  mutate(len = nchar(V1)) %>% #only two len: Households and Persons
  mutate(type = substr(V1, start=1, stop=1)) %>% 
  filter(type=="P") %>% #keeping only persons and NOT households
  mutate(age = as.numeric(substr(V1, start=25, stop=26))) %>%
  filter(age >=25 & age <=65) %>%
  mutate(employ_status = as.numeric(substr(V1, start=154, stop=154))) %>%
  filter(is.na(employ_status) == FALSE & (employ_status>=1 & employ_status<=3)) %>%
  mutate(employ_status = as.numeric(recode( as.character(employ_status), '3' ="0", '2' = "1", '1' = "1"))) %>%
  mutate(naics = substr(V1, start=215, stop=217)) %>%
  mutate(naics = ifelse(employ_status == 0, NA, naics)) %>%
  filter((employ_status == 0 & is.na(naics) == TRUE)|(employ_status == 1 & is.na(naics) == FALSE)) %>%
  mutate(naics = ifelse(naics == "23 ", "230", naics)) %>%
  mutate(naics = ifelse(naics == "31M", "313", naics)) %>%
  mutate(naics = ifelse(naics == "4MS", "423", naics)) %>%
  left_join(reclass_sectors, by=c('naics'='naics')) %>% #NAICS to final sector
  dplyr::rename(sector = final_sector) %>%
  filter(is.na(sector) == FALSE) #naics that are not in final sectors

#computing employment and unemployment, by sector and state
census <- census %>%
  group_by(state, sector) %>%
  mutate(num_employ = n()) %>%
  ungroup() %>%
  distinct(state, sector, .keep_all = TRUE) %>%
  select(state, sector, num_employ) %>%
  arrange(state, sector) %>%
  mutate(num_employ = ifelse(sector == 0, 0, num_employ)) %>% #unemployed total (not sample) will be added directly
  group_by(state) %>%
  mutate(tot = sum(num_employ)) %>%
  ungroup() %>%
  mutate(rel = num_employ/tot) %>%
  left_join(employment_pop, by=c('state'='state')) %>% #total employed people
  mutate(employ = rel*employed) %>%
  mutate(employ = ifelse(sector == 0, unemployed, employ)) %>%
  select(state, sector, employ)

final <- spread(census, sector, employ, fill = NA, convert = TRUE,  drop = TRUE, sep = NULL)

#saving
write.table(final, file = paste("1-Intermediate_Processed_Data//L_2000.csv", sep=""), sep = ",", row.names = FALSE)
  
rm(census, employment_pop, final)

```


#Task 2

```{r}

#Mobility matrices (mu)

#ACS: state to state-sector movements
yr_list <- c("00", "01", "02", "03", "04", "05", "06", "07" )
st_list <- c("al", "ak", "az", "ar", "ca", "co", "ct", "de", "fl", "ga", "hi", "id", "il", "in", "ia", "ks", "ky", "la", "me", "md", "ma", "mi", "mn", "ms", "mo", "mt", "ne", "nv", "nh", "nj", "nm", "ny", "nc", "nd", "oh", "ok", "or", "pa", "ri", "sc", "sd", "tn", "tx", "ut", "vt", "va", "wa", "wv", "wi", "wy")


#loading PUMS ACS
acs <- c() 
yr_c <- 2000
for (yr in yr_list) {
print(yr)
c <- 1
for (st in st_list) {
if (yr_c == 2000){ACS_PUMS <- read.csv(paste("0-Raw_Data/ACS/20", yr,"/c2ssp", st,".csv", sep=""), sep=",")}  
if (yr_c != 2000){ACS_PUMS <- read.csv(paste("0-Raw_Data/ACS/20", yr,"/ss", yr,"p", st,".csv", sep=""), sep=",")}    
ACS_PUMS$year <- yr_c
ACS_PUMS <- ACS_PUMS %>%
  select(year, AGEP, POWSP, MIGSP, ESR, NAICSP) %>%
  filter( is.na(AGEP) == TRUE  | (AGEP >= 25 & AGEP <=65)) %>% # age
  filter( is.na(ESR) == FALSE & (ESR>=1 & ESR<=3)) %>% #employ status
  mutate(ESR = as.numeric( recode( as.character(ESR), '3' ="0", '1' = "1", '2' = "1"))) %>%
  filter( is.na(POWSP) == TRUE | (POWSP != 11 & POWSP <= 56)) %>%
  left_join(states_fips, by=c('POWSP' = 'st_fips')) %>% 
  filter( is.na(st_num) == TRUE  | (st_num == c)) %>% # workplace (state)
  select(-st_name, -POWSP, -st_num) %>%
  mutate(state_dest = s_names[c]) %>% 
  filter( is.na(MIGSP) == TRUE | (MIGSP != 11 & MIGSP <= 56)) %>%
  left_join(states_fips, by=c('MIGSP' = 'st_fips')) %>% #lived 1 yr ago
  select(-MIGSP, -st_num, -AGEP) %>%
  dplyr::rename(state_ori = st_name, employ_status = ESR) %>%
  mutate(state_ori = ifelse(is.na(state_ori) == TRUE, s_names[c], state_ori)) %>% # no MIG
  dplyr::rename(naics = NAICSP) %>%
  mutate(naics = substr(naics, start=1, stop=3)) %>%
  mutate(naics = ifelse(employ_status == 0, NA, naics)) %>%
  filter((employ_status == 0 & is.na(naics) == TRUE)|(employ_status == 1 & is.na(naics) == FALSE)) %>%
  mutate(naics = ifelse(naics == "23", "230", naics)) %>%
  mutate(naics = ifelse(naics == "31M", "313", naics)) %>%
  mutate(naics = ifelse(naics == "4MS", "423", naics)) %>%
  left_join(reclass_sectors, by=c('naics'='naics')) %>% #NAICS to final sector
  dplyr::rename(sector_dest = final_sector) %>%
  filter(is.na(sector_dest) == FALSE) %>%
  select(-naics, -employ_status)

#counting by state and sector, origin and dest
ACS_PUMS <- ACS_PUMS %>%
  group_by(year, sector_dest, state_dest, state_ori) %>%
  mutate(n = n()) %>%
  ungroup() %>%
  distinct(year, sector_dest, state_dest, state_ori, .keep_all = TRUE)

#adding missing combinations state_sector_dest-state_ori
add <- c()
for (sec in 0:14) {
temp <- data.frame(s_names)
temp <- temp %>%
  dplyr::rename(state_ori = s_names) %>%
  mutate(state_dest = s_names[c], year = yr_c, sector_dest = sec, n = 0)  
add <- rbind(add, temp)
}
ACS_PUMS <- rbind(ACS_PUMS, add)
rm(add, temp)
ACS_PUMS  <- ACS_PUMS %>%
  group_by(year, sector_dest, state_dest, state_ori) %>%
  mutate(n_acs = max(n)) %>%
  ungroup() %>%
  select(-n) %>%
  distinct(year, sector_dest, state_dest, state_ori, .keep_all = TRUE) %>%
  arrange(year, sector_dest, state_dest, state_ori)

acs <- rbind(acs, ACS_PUMS)
 
c <- c + 1
}
yr_c <- yr_c + 1  
}
rm(ACS_PUMS)

```

```{r}

#Mobility matrices (mu)

#CPS: sector-sector within state movements
yr_list <- c("99", "00", "01", "02", "03", "04", "05", "06", "07" )
mn_list <- c("jan", "feb", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec" )

#loading PUMS CPS
cps <- c()
for (yr in yr_list){
print(yr)
for (mn in mn_list){
CPS_PUMS <- read.csv(paste("0-Raw_Data/CPS/", mn, yr,"pub_.txt", sep=""), header=FALSE, sep=";")
CPS_PUMS$year <- yr
CPS_PUMS$month <- mn
cps <- rbind(cps, CPS_PUMS)
}  
}
rm(CPS_PUMS)  

#creating variables
cps <- cps %>%
  mutate(year = as.numeric(year)) %>%
  mutate(year = ifelse(year != 99, 2000 + year, 1999)) %>%
  mutate(len = nchar(V1)) %>%
  mutate(H_ID = substr(V1, start=1, stop=15)) %>% #household ID 
  mutate(age = as.numeric(substr(V1, start=122, stop=123))) %>% #age
  filter(age >=25 & age <=65) %>%
  arrange(year, month, H_ID, -age) %>%
  group_by(year, month, H_ID) %>%
  mutate(P_ID = 1:n()) %>%
  ungroup() %>%
  mutate(employ_status = as.numeric(substr(V1, start=180, stop=181))) %>% #employ status
  filter( is.na(employ_status) == FALSE & (employ_status>=1 & employ_status<=4)) %>%
  mutate(employ_status = as.numeric( recode( as.character(employ_status), '4' ="0", '3' ="0", '2' = "1", '1' = "1"))) %>%
  mutate(st_fips = as.numeric(substr(V1, start=93, stop=94))) %>% #state
  filter( is.na(st_fips) == FALSE | (st_fips != 11 & st_fips <= 56)) %>%
  left_join(states_fips, by=c('st_fips' = 'st_fips')) %>% 
  select(-st_fips) %>%
  mutate(naics = (substr(V1, start=436, stop=438))) %>% #naics
  mutate(naics = ifelse(employ_status == 0, NA, naics)) %>%
  filter((employ_status == 0 & is.na(naics) == TRUE)|(employ_status == 1 & is.na(naics) == FALSE)) %>%
  left_join(reclass_sectors, by=c('naics'='naics')) %>% #NAICS to final sector
  dplyr::rename(sector = final_sector) %>%
  filter(is.na(sector) == FALSE) %>%
  select(-V1, -naics, -len, -employ_status, -st_num)

#matching persons via same month in consecutive years
base <- c()
for (yr in 2000:2007) {
for (mn in mn_list) {
temp_cps <- cps %>% 
  filter(month == mn & (year == yr | year == yr-1 )) %>%
  group_by(H_ID, P_ID) %>%
  filter( abs(max(age)-min(age)) <= 1 ) %>%
  ungroup()
info_present <- temp_cps %>%
  filter(year == yr) %>%
  select(-age) %>%
  dplyr::rename(sector_dest = sector, state_dest = st_name)
info_past <- temp_cps %>%
  filter(year == yr-1) %>%
  select(-year, -month) %>%
  dplyr::rename(sector_ori = sector, state_ori = st_name, age_ori = age)
temp <- full_join(info_present, info_past, by=c("H_ID", "P_ID"))
temp <- temp %>% 
  filter(is.na(state_ori) == FALSE & is.na(state_dest) == FALSE) #only matches
base <- rbind(base, temp)
}
}
#unifying months if they contain different matches
base <- base %>%
  distinct(year, H_ID, P_ID, age_ori, .keep_all = TRUE) %>%
  select(-age_ori, -month, -H_ID, -P_ID)
#counting sector-sector within state movements
base <- base %>%
  group_by(year, state_ori, sector_ori, state_dest, sector_dest) %>%
  mutate(n_cps = n()) %>%
  ungroup() %>%
  distinct(year, state_ori, sector_ori, state_dest, sector_dest, .keep_all = TRUE)

#adding missing combinations
add <- c()
for (sec_ori in 0:14) {
print(sec_ori)
temp <- acs %>%
  dplyr::rename(n_cps = n_acs) %>%
  mutate(sector_ori = sec_ori, n_cps = 0)  
add <- rbind(add, temp)
}  
base <- rbind(base, add)
rm(add, temp, info_past, info_present, temp_cps)
cps  <- base %>%
  group_by(year, state_ori, sector_ori, state_dest, sector_dest) %>%
  mutate(n_cps = max(n_cps)) %>%
  ungroup() %>%
  distinct(year, state_ori, sector_ori, state_dest, sector_dest, .keep_all = TRUE)
rm(base)  

#filling missing matrix sections as CDP
cps <- cps %>%
  mutate(n_master = ifelse(state_ori == state_dest, n_cps, 0)) %>%
  select(-n_cps) %>%
  group_by(year, state_dest, sector_dest, sector_ori) %>%
  mutate(n_cps = max(n_master)) %>%
  ungroup() %>%
  select(-n_master) %>%
  arrange(year, state_dest, sector_dest, state_ori, sector_ori)

#national CPS
cps <- cps %>%
  group_by(year, sector_ori, sector_dest) %>%
  mutate(n_national = sum(n_cps)) %>%
  ungroup()

```

```{r}

#Mobility matrices

#disaggregating ACS data using CPS sector_ori info and computation of mobility share
#origin in rows and destination in columns
#row shares sum to 1
mobility <- cps %>%
  left_join(acs, by=c('year' = 'year', 'state_dest' = 'state_dest', 'state_ori' = 'state_ori', 'sector_dest' = 'sector_dest')) %>% 
  group_by(year, state_dest, state_ori, sector_dest) %>% 
  mutate(tot_n_cps_accros_sec_ori = sum(n_cps)) %>% 
  ungroup() %>% 
  mutate(n_cps = ifelse(tot_n_cps_accros_sec_ori == 0, n_national, n_cps)) %>%
  group_by(year, state_dest, state_ori, sector_dest) %>% 
  mutate(tot_n_cps_accros_sec_ori = sum(n_cps)) %>% 
  ungroup() %>% 
  mutate(n_cps = ifelse(tot_n_cps_accros_sec_ori==0 & sector_ori==sector_dest, 1, n_cps)) %>%
  mutate(n_cps = ifelse(tot_n_cps_accros_sec_ori==0 & sector_ori!=sector_dest, 0, n_cps)) %>%
  group_by(year, state_dest, state_ori, sector_dest) %>% 
  mutate(tot_n_cps_accros_sec_ori = sum(n_cps)) %>% 
  ungroup() %>% 
  mutate(n_cps_adj = n_cps*n_acs/tot_n_cps_accros_sec_ori) %>% 
  group_by(year, state_ori, sector_ori) %>% 
  mutate(tot_n_cps_accros_dest = sum(n_cps_adj)) %>% 
  ungroup() %>%
  mutate(rel = n_cps_adj/tot_n_cps_accros_dest) %>% 
  mutate(rel = ifelse(is.na(rel)==TRUE & state_dest==state_ori & sector_dest== sector_ori, 1, rel)) %>%
  mutate(rel = ifelse(is.na(rel)==TRUE, 0, rel)) %>%
  distinct(year, state_dest, state_ori, sector_dest, sector_ori, .keep_all = TRUE) %>% 
  select(year, state_dest, state_ori, sector_dest, sector_ori, rel) %>%
  arrange(year, state_ori, sector_ori, state_dest, sector_dest)

problem <- mobility %>%
  filter(state_dest == state_ori & sector_dest == sector_ori) %>%
  filter(rel == 0)

write.table(mobility, file = paste("1-Intermediate_Processed_Data//mu_all.csv", sep=""), sep = ",", row.names = FALSE)

mobility <- mobility %>%
  mutate(sector_dest = recode(as.character(sector_dest), '0' ="00", '1' ="01", '2' ="02", '3' ="03", '4' ="04", '5' ="05", '6' ="06", '7' ="07", '8' ="08", '9' ="09" )) %>%
  mutate(state_dest_sector_dest = paste(state_dest, sector_dest, sep="_")) %>%
  select(-state_dest, -sector_dest)

epsilon <- 0.0001
for (yr in 2000:2007) {
final <- mobility %>%
  filter(year == yr)
final <- spread(final, state_dest_sector_dest, rel, fill = NA, convert = TRUE,  drop = TRUE, sep = NULL)
#check
check <- final %>% select(-year, -state_ori, -sector_ori)
if(sum(abs(as.matrix(rowSums(check)) - matrix(1, dim(check)[1], 1))) > epsilon) 
  stop(paste("mobility shares do not add to 1,", yr))
write.table(final, file = paste("1-Intermediate_Processed_Data//mu_", yr, ".csv", sep=""), sep = ",", row.names = FALSE)
}




```
