# General Instructions

This script does the following:

1) Computes the employment level for each state and sector in year 2000 using CBS data.

2) Computes the employment level for each state and sector in year 2000 and year 1999 using BLS data.

3) Computes the employment level for each country and sector in year 1999 and year 2000 using ILO and SEA data

4) Combines the previous outputs, and applies proportionality to ensure consistency with WIOD SEA

## Input files

1. `0-Raw_Data/sectors.csv`
2. `0-Raw_Data/regions.csv`
3. `0-Raw_Data/CENSUS_2000/REVISEDPUMS5/REVISEDPUMS5_.txt`
4. `0-Raw_Data/CENSUS_2000/employment_2000.xls`
5. `0-Raw_Data\L_1999\Sectors\sector_\1999.annual_.csv`
6. `0-Raw_Data\L_2000\Sectors\sector_\2000.annual_.csv`
7. `0-Raw_Data\staadata.xlsx`
8. `0-Raw_Data/CENSUS_2000/L_1999_2000_countries.csv`
9. `0-Raw_Data/CENSUS_2000/EAP_2WAP_SEX_AGE_RT_A-filtered-2023-11-02.csv`
10. `0-Raw_Data/CENSUS_2000/P_World unemployment rates.xlsx`

## Output files

1. `2-Final_Data//L_`2000CENSUS`.csv`   
2. `2-Final_Data//L_`1999BLS`.csv`
3. `2-Final_Data//L_`2000BLS`.csv`

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
remove(list = ls())
```

```{r message=FALSE, warning=FALSE, cache=FALSE, include=FALSE, results='hide'}
# libraries
options( scipen=100, digits = 15, stringsAsFactors=FALSE ); # no scientific notation, up to 15 digits 

# vector of libraries to be used
libs <- c( 'dplyr', 'tidyr', 'readxl', 'writexl', 'readr', 'data.table', "openxlsx", "readxl");
# Install libraries in case they are not installed
for( i in seq_along(libs)) { 
     if( !(libs[i] %in% installed.packages()) ) install.packages( libs[i] ) 
};
# Loading libraries
lapply( libs, require, character.only=TRUE );
```

# Sector-State employment for 2000 from Census

1. Construct employment levels for each US state from the 2000 Census, using person weights. 
2. Take CBS total non-participants for each state directly as the sector 0 value.

```{r message=FALSE, warning=FALSE}

# Importing general data sets to R

#US states
regions <- read.csv("0-Raw_Data/regions.csv")  
s_names <- regions %>% filter(status == "state") 
s_names <- s_names$region
n_s <- length(s_names)

# sector reclassification table (NAICS to final)
sectors_naics_final <- read.csv(paste("0-Raw_Data/sectors.csv", sep=""), header = TRUE, sep = ",", dec = ".")
sectors_naics_final <- sectors_naics_final %>% 
  dplyr::select(final_sector, naics) %>%
  mutate(naics = as.character(naics)) %>%
  distinct(final_sector, naics)

# employment statistics in year 2000, by state
employment_pop <- read_excel("0-Raw_Data/CENSUS_2000/employment_2000.xls", skip = 12)
employment_pop <- employment_pop %>%
  dplyr::rename(state = ...1, Not_LF = ...8, employed = Employed, unemployed = Unemployed) %>%
  filter(is.na(state) == FALSE) %>%
  mutate(state = gsub(" ", "", state)) %>% #taking out spaces
  filter(state %in% s_names) %>%
  dplyr::select(state, employed, Not_LF) %>%
  mutate(Not_LF = Not_LF/1000, employed = employed/1000) #units in thousands (as WIOD)

# loading Census PUMS 2000
list <- c("01", "02", "04", "05", "06", "08", "09", setdiff(10:56, c(11,14,43,52)))
c <- 1
census_list <- vector("list", length = n_s)

for (i in list){
CENSUS_PUMS <- read_csv(paste("0-Raw_Data/CENSUS_2000/REVISEDPUMS5/REVISEDPUMS5_", i,".txt", sep=""), col_names = FALSE)
CENSUS_PUMS$state <- s_names[c]

# creating variables: age, employment status, person weight, and industry info for each US state
CENSUS_PUMS <- CENSUS_PUMS %>%
  mutate(len = nchar(X1)) %>% #only two len: Households and Persons; len matches with dictionary
  mutate(type = substr(X1, start=1, stop=1)) %>%  #identify Households and Persons
  filter(type=="P") %>% #keeping only Persons and NOT Households
  mutate(personweight = substr(X1, start=13, stop=16)) %>%
  mutate(age = as.numeric(substr(X1, start=25, stop=26))) %>% 
  filter(age >=25 & age <=65) %>%
  mutate(employ_status = as.numeric(substr(X1, start=154, stop=154))) %>%
  filter(is.na(employ_status) == FALSE & ((employ_status>=1 & employ_status<=3)|employ_status==6)) %>%
  mutate(employ_status = as.numeric(recode( as.character(employ_status), '6' ="0", '3' ="0", '2' = "1", '1' = "1"))) %>% 
  mutate(naics = substr(X1, start=215, stop=217)) %>% #industry information
  mutate(naics = ifelse(employ_status == 0, NA, naics)) %>%
  filter((employ_status == 0 & is.na(naics) == TRUE)|(employ_status == 1 & is.na(naics) == FALSE)) %>%
  mutate(naics = ifelse(naics == "23 ", "230", naics)) %>%
  mutate(naics = ifelse(naics == "31M", "313", naics)) %>%
  mutate(naics = ifelse(naics == "4MS", "423", naics)) %>%
  left_join(sectors_naics_final, by=c('naics'='naics')) %>% #NAICS to final sector
  dplyr::rename(sector = final_sector) %>%
  filter(is.na(sector) == FALSE) %>% #NAICS that are not in final sectors
  dplyr::select(-X1)
census_list[[c]] <- CENSUS_PUMS #each list in census_list contains all the micro-info for each US state
rm(CENSUS_PUMS)
print(s_names[c])
c <- c + 1
}
census <- rbindlist(census_list)
census$personweight <- as.numeric(census$personweight)
rm(census_list)

# computing employment, by sector and state
census <- census %>%
  group_by(state, sector) %>%
  mutate(num_employ = sum(personweight)) %>%
  ungroup() %>%
  distinct(state, sector, .keep_all = TRUE) %>%
  dplyr::select(state, sector, num_employ) %>%
  arrange(state, sector) %>%
  mutate(num_employ = ifelse(sector == 0, 0, num_employ)) %>% #non-participants total (not sample) will be added directly
  group_by(state) %>%
  mutate(tot = sum(num_employ)) %>%
  ungroup() %>%
  mutate(rel = num_employ/tot) %>% #employment share for each sector
  left_join(employment_pop, by=c('state'='state')) %>% #total employed people
  mutate(employ = rel*employed) %>% #total employed people by sector (not sample)
  mutate(employ = ifelse(sector == 0, Not_LF, employ)) %>% #total non-participants
  dplyr::select(state, sector, employ)
```

# Sector-State employment for 1999 and 2000 from BLS data

1. Calculate employment by sector and state directly from BLS, for sectors 1-14.
2. Take BLS total non-participants for each state directly as the sector 0 value.

```{r message=FALSE, warning=FALSE}
################################
### US-sectors 1-14 for 1999 ###
################################
# Main directory
mainpath <- "0-Raw_Data/L_1999/Sectors"

#Load every NAICS sub-sector and append them according to our sectors.
for (i in 1:14) {
  current_sector <- paste0("sector", i)
  current_sector_file <- file.path(mainpath, current_sector)
  files <- list.files(path = current_sector_file, pattern = "\\.csv", full.names = TRUE)
  concat <- data.frame()
  for (file in files) {
    data <- read.csv(file, header = TRUE, stringsAsFactors = FALSE)
    concat <- rbind(concat, data)
  }
  assign(paste0("sector_", i), concat)
}
#Select states only (no counties).
for (i in 1:14) {
  sector <- paste0("sector_", i)
  assign(sector, get(sector)[substr(get(sector)$area_title, nchar(get(sector)$area_title) - 8, nchar(get(sector)$area_title)) == "Statewide", , drop = FALSE])
}
#Sum over NAICS sub-sectors for each state to get our sectors.
for (i in 1:14) {
  sector <- paste0("sector_", i)
  assign(sector, get(sector) %>%
           group_by(area_title) %>%
           mutate(employment = sum(annual_avg_emplvl, na.rm = TRUE)) %>%
           ungroup()
  )
}
#Leave only one observation for each state for each of our sectors.
for (i in 1:14) {
  sector <- paste0("sector_", i)
  assign(sector, dplyr::select(get(sector), employment, area_title) %>%
           dplyr::distinct()
  )
}
#Leave only each state's name.
for (i in 1:14) {
  sector <- paste0("sector_", i)
  assign(sector, get(sector) %>%
           mutate(state = substr(area_title, 1, nchar(area_title) - 13))
  )
}
#Drop Puerto Rico and Virgin Islands.
for (i in 1:14) {
  sector <- paste0("sector_", i)
  assign(sector, get(sector) %>%
           filter(state != "Puerto Rico" & state != "Virgin Islands") %>%
           dplyr::select(-area_title)
  )
}
#Add columns to indicate year and sector.
for (i in 1:14) {
  namesector <- paste0("sector_", i)
  assign(namesector, get(namesector) %>%
           mutate(sector = ifelse(i < 10, sprintf("%02d", i), as.character(i))) %>%
           mutate(year = 1999)
  )
}
#Add missing states.
sector_4 <- rbind(sector_4, data.frame(state = c("Vermont", "Rhode Island"),
                                       year = 1999,
                                       employment = c(0, 0),
                                       sector = rep("04", 2)))
sectorsmissingstates <- 5:8
for (i in sectorsmissingstates) {
  sector <- paste0("sector_", i)
  assign(sector, rbind(get(sector), data.frame(state = "Alaska", year = 1999, employment = 0, sector = sprintf("%02d", i))))
}
sector_10 <- rbind(sector_10, data.frame(state = c("Alaska", "Hawaii"),
                                         year = 1999,
                                         employment = c(0, 0),
                                         sector = rep("10", 2)))
#Put together all the sectors.
states_1999BLS <- bind_rows(mget(paste0("sector_", 1:14)))

################################
### US-sectors 1-14 for 2000 ###
################################
# Main directory
mainpath <- "0-Raw_Data/L_2000/Sectors"
#Load every NAICS sub-sector and append them according to our sectors.
for (i in 1:14) {
  current_sector <- paste0("sector", i)
  current_sector_file <- file.path(mainpath, current_sector)
  files <- list.files(path = current_sector_file, pattern = "\\.csv", full.names = TRUE)
  concat <- data.frame()
  for (file in files) {
    data <- read.csv(file, header = TRUE, stringsAsFactors = FALSE)
    concat <- rbind(concat, data)
  }
  assign(paste0("sector_", i), concat)
}
#Select states only (no counties).
for (i in 1:14) {
  sector <- paste0("sector_", i)
  assign(sector, get(sector)[substr(get(sector)$area_title, nchar(get(sector)$area_title) - 8, nchar(get(sector)$area_title)) == "Statewide", , drop = FALSE])
}
#Sum over NAICS sub-sectors for each state to get our sectors.
for (i in 1:14) {
  sector <- paste0("sector_", i)
  assign(sector, get(sector) %>%
           group_by(area_title) %>%
           mutate(employment = sum(annual_avg_emplvl, na.rm = TRUE)) %>%
           ungroup()
  )
}
#Leave only one observation for each state for each of our sectors.
for (i in 1:14) {
  sector <- paste0("sector_", i)
  assign(sector, dplyr::select(get(sector), employment, area_title) %>%
           distinct()
  )
}
#Leave only each state's name.
for (i in 1:14) {
  sector <- paste0("sector_", i)
  assign(sector, get(sector) %>%
           mutate(state = substr(area_title, 1, nchar(area_title) - 13))
  )
}
#Drop Puerto Rico and Virgin Islands.
for (i in 1:14) {
  sector <- paste0("sector_", i)
  assign(sector, get(sector) %>%
           filter(state != "Puerto Rico" & state != "Virgin Islands") %>%
           dplyr::select(-area_title)
  )
}
#Add columns to indicate year and sector.
for (i in 1:14) {
  namesector <- paste0("sector_", i)
  assign(namesector, get(namesector) %>%
           mutate(sector = ifelse(i < 10, sprintf("%02d", i), as.character(i))) %>%
           mutate(year = 2000)
  )
}
#Add missing states.
sector_4 <- rbind(sector_4, data.frame(state = c("Vermont", "Rhode Island"),
                                       year = 2000,
                                       employment = c(0, 0),
                                       sector = rep("04", 2)))
sectorsmissingstates <- 5:8
for (i in sectorsmissingstates) {
  sector <- paste0("sector_", i)
  assign(sector, rbind(get(sector), data.frame(state = "Alaska", year = 2000, employment = 0, sector = sprintf("%02d", i))))
}
sector_10 <- rbind(sector_10, data.frame(state = c("Alaska", "Hawaii"),
                                         year = 2000,
                                         employment = c(0, 0),
                                         sector = rep("10", 2)))
states_2000BLS <- bind_rows(mget(paste0("sector_", 1:14)))


####################
### US-sector 0 ####
####################

#Prepare BLS labor force participation data
PARTUS <- read_excel("0-Raw_Data/staadata.xlsx", range = "B1228:E1333",col_names = FALSE)
colnames(PARTUS) <- c("state", "year", "pop", "part")
PARTUS <- PARTUS %>%
  mutate(employment = pop - part)
PARTUS <- PARTUS %>%
  filter(!(state %in% c("Los Angeles County", "District of Columbia", "New York city")))
#Create 1999's sector 0 (non-participant).
sector_0 <- subset(PARTUS, year == 1999)
sector_0 <- sector_0 %>%
  dplyr::select(year, employment, state)
sector_0 <- sector_0  %>%
  mutate(sector="00")
states_1999BLS <- rbind(states_1999BLS,sector_0)
#Create 2000's sector 0 (non-participant).
sector_0 <- subset(PARTUS, year == 2000)
sector_0 <- sector_0 %>%
  dplyr::select(year, employment, state)
sector_0 <- sector_0  %>%
  mutate(sector="00")
states_2000BLS <- rbind(states_2000BLS,sector_0)
```

# Sector-Country employment for 1999 and 2000

1. Calculate the number of non-participants for each country. 
2. Use WIOD-SEA for all employment sectors in each country.

For each country $c$ (excluding USA), let $L^c_{tot}$ represent working age population, $L^c_{work}$ represent employment, $unemp^c$ represent unemployment, and $L^c_{i}$ represent "employment" in sector $i$ $\forall i=0,...,14$. If we define $L^c_0$ as non-participants in the labor force, and if we assume that the labor market only has 14 sectors, then we know that:
$$
L^c_{tot}	=L^c_{0}+unemp^c+L^c_{work}
$$
$$
L^c_{work}	=\sum_{i=1}^{14}L^c_{i}
$$
To ease notation, here we ignore from which year we are taking each variable. However, in the exposition of step 4 we do include a time superscript to indicate the year; the rest of the notation remains the same.

From ILO we know the value of the labor force participation rate ($lfp^c$) for each country; which using our assumptions satisfies
$$
1-lfp^c	=\dfrac{L_{0}^{c,ILO}}{L_{tot}^{c,ILO}}
$$
The superscripts indicate where the data is taken from; we adopt the same convention in what follows. Also, from ILO we know the value of the unemployment rate ($u^c$) for each country, which is:

$$
u^c	=\dfrac{unemp^{c,ILO}}{unemp^{c,ILO}+L_{work}^{c,ILO}}
$$
For the SEA values of each of our 14 (employment) sectors to be consistent with $u^c$ and $lfp^c$, we must have
$$
1-lfp^c = \dfrac{L_{0}^{c,SEA}}{L_{0}^{c,SEA}+unemp^{c,SEA}+L_{work}^{c,SEA}} \; \; \; (1)
$$
$$
unemp^{c,SEA}	=u^c\times(L_{work}^{c,SEA}+unemp^{c,SEA}) \; \; \; (2)
$$
Equation $(2)$ implies
$$
unemp^{c,SEA}(1-u^c)	=u^c\times L_{work}^{c,SEA}
\Leftrightarrow unemp^{c,SEA}	=\frac{u^c}{1-u^c}L_{work}^{c,SEA}
$$
Hence, we can substitute for $unemp^{c,SEA}$ in equation $(1)$:
$$
1-lfp^c=\dfrac{L_{0}^{c,SEA}}{L_{0}^{c,SEA}+\left[\frac{1}{1-u^c}\right]L_{work}^{c,SEA}}
$$
Then,
$$
(1-lfp^c)\times L_{0}^{c,SEA}+(1-lfp^c)\left[\frac{1}{1-u^c}\right]L_{work}^{c,SEA}	=L_{0}^{c,SEA} \\
\Rightarrow -lfp^c\times L_{0}^{c,SEA}+(1-lfp^c)\left[\frac{1}{1-u^c}\right]L_{work}^{c,SEA}	= 0 \\
\Rightarrow L_{0}^{c,SEA}=\frac{1-lfp^c}{lfp^c}\left[\frac{1}{1-u^c}\right]L_{work}^{c,SEA}
$$

We use the formula in the previous line to calculate $L^c_{0}$ for each country using the unemployment rate and the labor force participation rate (both taken from ILO), and total employment, $\sum_{i=1}^{14}L^c_{i}$ (taken from SEA).  

```{r message=FALSE, warning=FALSE}
# Prepare lfp for each country.
participation <- fread("0-Raw_Data/CENSUS_2000/EAP_2WAP_SEX_AGE_RT_A-filtered-2023-11-02.csv", encoding = "UTF-8")
participation <- subset(participation, select = c(time, ref_area, obs_value))
participation <- participation %>%
  mutate(obs_value = obs_value/100)
participation <- participation %>%
  rename(
    year = time,
    lfp_rate = obs_value,
    region = ref_area
  )

#Prepare u for each country.
unemployment<- read_excel("0-Raw_Data/CENSUS_2000/P_World unemployment rates.xlsx")
unemployment_1999 <- unemployment %>%
  dplyr::select(2, 5) %>%
  rename(region = colnames(.)[1], u_rate = colnames(.)[2])
unemployment_1999$year <- 1999
unemployment_2000 <- unemployment %>%
  dplyr::select(2, 6) %>%
  rename(region = colnames(.)[1], u_rate = colnames(.)[2])
unemployment_2000$year <- 2000
unemployment <- rbind(unemployment_1999,unemployment_2000)
unemployment$u_rate <- as.numeric(unemployment$u_rate)
unemployment <- unemployment %>%
  mutate(u_rate = u_rate/100)
#Taiwan's unemployment rate is missing from ILO, so we take it from https://www.worldeconomics.com/Demographics/Unemployment-Rate/Taiwan.aspx
TWNu <- data.frame(
  region = c("TWN", "TWN"),
  year = c(1999, 2000),
  u_rate = c(0.029, 0.03)
)
unemployment <- rbind(unemployment,TWNu)
rates <- merge(unemployment, participation, by = c("year", "region"))
rates <- rates[rates$region != "RWA", ]
#We take RoW's lfp and u rates as the average of the corresponding rates for the other countries in our list.
ROW <- rates %>%
  group_by(year) %>%
  summarize(lfp_rate = mean(lfp_rate, na.rm = TRUE),
            u_rate = mean(u_rate, na.rm = TRUE))
ROW$region <- "RoW"
rates <- rbind(rates,ROW)

#Prepare L_work for each country
countries <- fread("0-Raw_Data/CENSUS_2000/L_1999_2000_countries.csv", encoding = "UTF-8")
tot_employ_countries <- countries %>%
  group_by(region, year) %>%
  summarize(employ_sum = sum(employ))

#Calculate secto_0 for each country.
sector0countries <- merge(rates, tot_employ_countries, by = c("year", "region"))
sector0countries <- sector0countries %>%
  mutate(employ = ((1-lfp_rate)/lfp_rate)*(1/(1-u_rate))*employ_sum)
sector0countries <- subset(sector0countries, select = c(employ, year, region))
sector0countries$sector <- 0
countries <- rbind(countries,sector0countries)
countries$sector <- sprintf("sector_%02d", countries$sector)
```

# Combining previous outputs and doing final checks

1. Employment for US-states for year 2000 must sum up to US' total employment for year 2000 according to SEA. 

Specifically, if $L_{i,s}^{2000}$ represents the employment for year 2000 of sector $s=0,...,14$ and state $i=1,...,50$ that we calculated in step 1 (for CBS) and step 2 (for BLS), for every $s\neq0$ we do the following normalization:
$$
 L_{i,s}^{*CENSUS,2000} = L_{i,s}^{CENSUS,2000} \times\frac{\sum_{s=1}^{14}L_{s}^{USA,SEA,2000}}{\sum_{i=1}^{50}\sum_{s=1}^{14}L_{i,s}^{CENSUS,2000}} \\
 L_{i,s}^{*BLS,2000} = L_{i,s}^{BLS,2000} \times\frac{\sum_{s=1}^{14}L_{i}^{USA,SEA,2000}}{\sum_{i=1}^{50}\sum_{s=1}^{14}L_{i,s}^{BLS,2000}} 
$$
where $L_{i,s}^{*2000}$ is our "official" value for the employment for year 2000 of sector $s$ and state $i$.

For units of sector 0 to be consistent with this normalization, we apply the same proportionality constant to them:
$$
L_{i,0}^{*CENSUS,2000} = L_{i,0}^{CENSUS,2000} \times\frac{\sum_{s=1}^{14}L_{s}^{USA,SEA,2000}}{\sum_{i=1}^{50}\sum_{s=1}^{14}L_{i,s}^{CENSUS,2000}} \\
L_{i,0}^{*BLS,2000} = L_{i,0}^{BLS,2000} \times\frac{\sum_{s=1}^{14}L_{i}^{USA,SEA,2000}}{\sum_{i=1}^{50}\sum_{s=1}^{14}L_{i,s}^{BLS,2000}}
$$
2. Employment in year 1999 and year 2000 must be consistent with zero population growth for each country (including USA in the case of BLS data).

Specifically, using the values of $L_{s}^{c,SEA,1999}$ and $L_{s}^{c,SEA,2000}$ that we calculated in step 3 for each country $c$ (excluding USA) and each sector $s$, we do the following normalization:

$$
L_{s}^{*c,SEA,1999}  =  L_{s}^{c,SEA,1999} \times\frac{L_{0}^{c,SEA,2000}+\sum_{s=1}^{14}L_{s}^{c,SEA,2000}}{L_{0}^{c,SEA,1999}+\sum_{s=1}^{14}L_{s}^{c,SEA,1999}}
$$
where $L_{s}^{*c,1999}$ is our "official" value for the employment for year 1999 of sector $s$ and country $c$.

Similarly, for USA, letting $L_{i,s}^{*2000}$ represent our "official" value for the employment for year 1999 of sector $s$ and state $i$, we do the following normalization:
$$
 L_{i,s}^{*,BLS,1999} = L_{i,s}^{BLS,1999} \times\frac{\sum_{i=1}^{50}\sum_{s=0}^{14}L_{i,s}^{*,BLS,2000}}{\sum_{i=1}^{50}\sum_{s=0}^{14}L_{i,s}^{BLS,1999}}
$$
where $L_{i,s}^{1999}$ represents the employment for year 2000 of sector $s=0,...,14$ and state $i=1,...,50$ that we calculated in step 2 for BLS.

3. Reshape employment for states (CBS and BLS) and countries, and bind them to get a labor matrix. 

```{r message=FALSE, warning=FALSE}
###############################################################################
#US' total employment for year 2000 according to SEA.
USA2000 <- subset(countries, year == 2000 & region == "USA" & sector != "sector_00")
total_USA2000 <- sum(USA2000$employ)

#Normalization for BLS employment (year 2000).
employment2000BLS <- subset(states_2000BLS, sector != "00")
total_2000BLS <- sum(employment2000BLS$employment)
states_2000BLS$employment <- states_2000BLS$employment * total_USA2000/total_2000BLS

#Normalization for BLS employment (year 1999).
total_2000BLS <- sum(states_2000BLS$employment)
total_1999BLS <- sum(states_1999BLS$employment)
states_1999BLS$employment <- states_1999BLS$employment * total_2000BLS/total_1999BLS

#Normalization for CBS
states_2000CENSUS <- census
colnames(states_2000CENSUS)[colnames(states_2000CENSUS) == "employ"] <- "employment"
employstates_2000CENSUS <- subset(census, sector != 0)
total_2000CENSUS <- sum(employstates_2000CENSUS$employ)
states_2000CENSUS$employment <- states_2000CENSUS$employment*total_USA2000/total_2000CENSUS

#Change labels, and order to prepare for reshape.
colnames(states_2000CENSUS)[colnames(states_2000CENSUS) == "state"] <- "region"
states_2000CENSUS$sector <- sprintf("sector_%02d", states_2000CENSUS$sector)
states_1999BLS$sector <- paste0("sector_", states_1999BLS$sector)
states_1999BLS <- states_1999BLS %>% 
  arrange(sector)
colnames(states_1999BLS)[colnames(states_1999BLS) == "state"] <- "region"
states_2000BLS$sector <- paste0("sector_", states_2000BLS$sector)
states_2000BLS <- states_2000BLS %>% 
  arrange(sector)
colnames(states_2000BLS)[colnames(states_2000BLS) == "state"] <- "region"
states_1999BLS$year <- NULL
states_2000BLS$year <- NULL

#Separate countries (1999 and 2000) and drop USA because now we have states.
#And make country employment in 1999 sum up to country employment in 2000.
countries <- subset(countries,region != "USA")
countries <- countries %>%
  group_by(year, region) %>%
  mutate(tot_employ = sum(employ))
countries_1999 <- countries[countries$year == 1999, ]
countries_1999$year <- NULL
countries_2000 <- countries[countries$year == 2000, ]
countries_2000$year <- NULL
countries_1999$sum2000 <- countries_2000$tot_employ
countries_1999$employ <- countries_1999$employ*countries_1999$sum2000 / countries_1999$tot_employ
countries_1999 <- countries_1999 %>%
  dplyr::select(-sum2000, -tot_employ)
countries_2000$tot_employ <- NULL

#Reshape and append to get the complete datasets
states_1999BLS <- pivot_wider(states_1999BLS, 
                           names_from = sector, 
                           values_from = employment, 
                           values_fill = 0)
states_2000BLS <- pivot_wider(states_2000BLS, 
                           names_from = sector, 
                           values_from = employment, 
                           values_fill = 0)
states_2000CENSUS <- pivot_wider(states_2000CENSUS, 
                           names_from = sector, 
                           values_from = employment, 
                           values_fill = 0)
countries_1999 <- pivot_wider(countries_1999, 
                              names_from = sector, 
                              values_from = employ, 
                              values_fill = 0)
countries_1999 <- countries_1999 %>%
  mutate(region = ifelse(region == "RoW", "ROW", region)) %>%
  arrange(region) %>%
  mutate(region = ifelse(region == "ROW", "RoW", region))
countries_2000 <- pivot_wider(countries_2000, 
                              names_from = sector, 
                              values_from = employ, 
                              values_fill = 0)
countries_2000 <- countries_2000 %>%
  mutate(region = ifelse(region == "RoW", "ROW", region)) %>%
  arrange(region) %>%
  mutate(region = ifelse(region == "ROW", "RoW", region))
L_1999BLS <- rbind(states_1999BLS,countries_1999)
L_2000BLS <- rbind(states_2000BLS,countries_2000)
L_2000CENSUS <- rbind(states_2000CENSUS,countries_2000)

#Save data sets

wb <- loadWorkbook("2-Final_Data/Output.xlsx")
addWorksheet(wb, sheetName = "L_2000CENSUS")
writeData(wb, sheet = "L_2000CENSUS", x = L_2000CENSUS)
saveWorkbook(wb, file = "2-Final_Data/Output.xlsx", overwrite = TRUE)
addWorksheet(wb, sheetName = "L_1999BLS")
writeData(wb, sheet = "L_1999BLS", x = L_1999BLS)
saveWorkbook(wb, file = "2-Final_Data/Output.xlsx", overwrite = TRUE)
addWorksheet(wb, sheetName = "L_2000BLS")
writeData(wb, sheet = "L_2000BLS", x = L_2000BLS)
saveWorkbook(wb, file = "2-Final_Data/Output.xlsx", overwrite = TRUE)
```