# General Instructions

This script does the following:

1) Computes ACS total state to state migration movements (knowing the sector of destination, but not the sector of origin), by year for 1999-2001.That is, we calculate $L_{ACS}^{n?,iq}$: migration from state $n$ to state $i$ and sector $q$, according to ACS.

2) Computes IRS total state to state migration movements (not knowing the sector of origin, nor the sector of destination), for 1999. That is, we calculate $L_{IRS}^{n,i}$: migration from state $n$ to state $i$, according to IRS.

3) Computes CPS total state to state migration movements (knowing both the sector of origin and the sector of destination), by year for 1999-2001. That is, we calculate $L_{CPS}^{{nj,ik}}$: migration from state $n$ and sector $j$ to state $i$ and sector $k$, according to CPS.

4) We combine the data from the previous three steps to compute our final $L^{{nj,ik}}$'s. Then we calculate the corresponding mobility shares: $\mu_{nj,ik}$'s.

5) We calculate $\mu_{nj,ik}$'s for a non-migration case (that is, not allowing state to state movements, just allowing sector reallocation).

## Input files

1. `0-Raw_Data/regions.csv`
2. `0-Raw_Data/sectors.csv`
3. `0-Raw_Data/Fips/states_fips_num.xlsx`
4. `0-Raw_Data/ACS/20*/c2ssp**.csv`
5. `0-Raw_Data/IRS/1999to2000CountyMigration/1999to2000flow.dta`
6. `0-Raw_Data/CPS/CPS_NBER/Inputs/cps_panel.dta`

## Output files

1. `1-Intermediate_Processed_Data//acs_temp.csv`
2. `1-Intermediate_Processed_Data/acs.csv`
3. `1-Intermediate_Processed_Data/irs.csv`
4. `1-Intermediate_Processed_Data/cps_nber_2_yearly.csv`
5. `2-Final_Data/mu_1999.xlsx`
6. `2-Final_Data/mu_1999_no_migration.xlsx`

```{r Clean environment, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
remove(list = ls())  
```

```{r Set environment, message=FALSE, warning=FALSE, cache=FALSE, include=FALSE, results='hide'}
# Define function
`%notin%` <- Negate(`%in%`)
# libraries
options( scipen=100, digits = 15, stringsAsFactors=FALSE ); # no scientific notation, up to 15 digits 
# vector of libraries to be used
libs <- c( 'dplyr', 'tidyr', 'haven', 'stringr', 'nleqslv', 'gdata', 'formattable', 'ggplot2', 'viridis', 'RColorBrewer', 'reshape2', 'foreign', 'readxl', 'writexl', 'rjson', 'read_excel', 'Matrix', 'geodist', 'readr', 'data.table', 'stringr', 'haven');
# Install libraries in case they are not installed
for( i in seq_along(libs)){ 
    if( !(libs[i] %in% installed.packages()) ) install.packages( libs[i] ) 
};
# Loading libraries
lapply( libs, require, character.only=TRUE );

```


# Importing preliminary data
```{r Input data}
#States and countries.
regions <- read.csv("0-Raw_Data/regions.csv")  
c_names <- regions %>% filter(status == "country")
c_names <- c_names$region
s_names <- regions %>% filter(status == "state") 
s_names <- s_names$region

#Fips state codes
states_fips <- read_excel("0-Raw_Data/Fips/states_fips_num.xlsx", range = "A1:D53")
states_fips_mod_ACS <- states_fips %>% 
  select(st_num, st_name) %>%
  #make the codes consistent with ACS
  mutate(st_num = ifelse(st_num > 8, st_num + 1, st_num)) %>%
  distinct(st_name, .keep_all = TRUE) %>%
  dplyr::rename(st_name1 = st_name)
states_fips <- states_fips %>% dplyr::select(st_fips, st_num, st_name)

#Sector reclassification table (NAICS to final)
sectors_naics_final <- read.csv(paste("0-Raw_Data/sectors.csv", sep=""), header = TRUE, sep = ",", dec = ".")
sectors_naics_final <- sectors_naics_final %>% 
  select(final_sector, naics) %>%
  mutate(naics = as.character(naics)) %>%
  distinct(final_sector, naics)

```



# ACS migration movements

Compute $L_{ACS}^{n?,iq}$.
```{r Clean ACS, message=FALSE, warning=FALSE}
#list of years and states
yr_list <- c("00", "01", "02")
st_list <- c("al", "ak", "az", "ar", "ca", "co", "ct", "de", "fl", "ga", "hi", "id", "il", "in", "ia", "ks", "ky", "la", "me", "md", "ma", "mi", "mn", "ms", "mo", "mt", "ne", "nv", "nh", "nj", "nm", "ny", "nc", "nd", "oh", "ok", "or", "pa", "ri", "sc", "sd", "tn", "tx", "ut", "vt", "va", "wa", "wv", "wi", "wy")

#loading PUMS ACS
acs <- c() 
yr_c <- 2000
for (yr in yr_list) {
print(yr)
c <- 1
for (st in st_list) {
if (yr_c == 2000){ACS_PUMS <- read_csv(paste("0-Raw_Data/ACS/20", yr,"/c2ssp", st,".csv", sep=""))}  
if (yr_c != 2000){ACS_PUMS <- read_csv(paste("0-Raw_Data/ACS/20", yr,"/ss", yr,"p", st,".csv", sep=""))}    
ACS_PUMS$year <- yr_c
ACS_PUMS <- ACS_PUMS %>%  
  #ACS: inter- and intra-state flows but without knowing the sector of origin
  select(year, AGEP, POWSP, MIGSP, ESR, NAICSP, PWGTP) %>%
  mutate(AGEP = as.numeric(AGEP),
         POWSP = as.numeric(POWSP),
         MIGSP = as.numeric(MIGSP),
         ESR = as.numeric(ESR), 
         PWGTP = as.numeric(PWGTP)) %>%
  #select range of age
  filter( is.na(AGEP) == TRUE  | (AGEP >= 25 & AGEP <=65)) %>%
  #select (and relabel) employment status of interest
  filter( is.na(ESR) == FALSE & (ESR==1 |ESR==2 | ESR==3 | ESR==6)) %>%
  mutate(ESR = as.numeric( recode( as.character(ESR), '6' ="0", '3' ="0", '1' = "1", '2' = "1"))) %>%
  #Filter and merge Fips codes.
  filter( is.na(POWSP) == TRUE | (is.na(POWSP) == FALSE & POWSP != 11 & POWSP <= 56 & year>=2003) | (is.na(POWSP) == FALSE & POWSP != 9 & POWSP <= 51 & year<=2002)) %>%
  left_join(states_fips, by=c('POWSP' = 'st_fips')) %>% 
  mutate(st_num = ifelse(year <= 2002, POWSP, st_num)) %>%
  mutate(st_num = ifelse(year <= 2002 & POWSP > 8, st_num - 1, st_num)) %>%        #state of workplace
  filter( is.na(st_num) == TRUE  | (is.na(st_num) == FALSE & st_num == c)) %>% # 
  select(-st_name, -POWSP,-st_num) %>% 
  mutate(state_dest = s_names[c]) %>% 
  #migration (indicate state of origin from 1 year ago)
  filter( is.na(MIGSP) == TRUE | (is.na(MIGSP) == FALSE & MIGSP != 11 & MIGSP <= 56 & year>=2003) | (is.na(MIGSP) == FALSE & MIGSP != 9 & MIGSP <= 51 & year<=2002)) %>%
  left_join(states_fips, by=c('MIGSP' = 'st_fips')) %>%
  left_join(states_fips_mod_ACS, by=c('MIGSP' = 'st_num')) %>%
  mutate(st_name = ifelse(year <= 2002, st_name1, st_name)) %>%
  select(-MIGSP, -st_num, -AGEP, -st_name1) %>%
  dplyr::rename(state_ori = st_name, employ_status = ESR) %>%
  mutate(state_ori = ifelse(is.na(state_ori) == TRUE, s_names[c], state_ori)) %>%   #Map NAICS to final sector (and label as sector of destination)
  dplyr::rename(naics = NAICSP) %>%
  mutate(naics = substr(naics, start=1, stop=3)) %>%
  mutate(naics = ifelse(employ_status == 0, NA, naics)) %>%
  filter((employ_status == 0 & is.na(naics) == TRUE)|(employ_status == 1 & is.na(naics) == FALSE)) %>%
  mutate(naics = ifelse(naics == "23", "230", naics)) %>%
  mutate(naics = ifelse(naics == "31M", "313", naics)) %>%
  mutate(naics = ifelse(naics == "4MS", "423", naics)) %>%
  left_join(sectors_naics_final, by=c('naics'='naics')) %>%
  dplyr::rename(sector_dest = final_sector) %>%
  filter(is.na(sector_dest) == FALSE) %>%
  select(-naics, -employ_status)  

#At this point we have the ACS PUMS person weights (PWGTP, at surveyed representative individual level) for state of origin, state of destination and sector destination by year.

#Counting total by (state of origin)-(state of destination-sector of destination)
ACS_PUMS <- ACS_PUMS %>%
  group_by(year, sector_dest, state_dest, state_ori) %>%
  mutate(n = sum(PWGTP)) %>% #using weights
  ungroup() %>%
  distinct(year, sector_dest, state_dest, state_ori, .keep_all = TRUE) %>%  #keep one migration combination
  select(-PWGTP)
#adding missing combinations
add <- c()
for (sec in 0:14) {
temp <- data.frame(s_names)
temp <- temp %>%
  dplyr::rename(state_ori = s_names) %>%
  mutate(state_dest = s_names[c], year = yr_c, sector_dest = sec, n = 0)
add <- rbind(add, temp)
}
ACS_PUMS <- rbind(ACS_PUMS, add)
rm(add, temp)
ACS_PUMS  <- ACS_PUMS %>%
  group_by(year, sector_dest, state_dest, state_ori) %>%
  mutate(n_acs = max(n)) %>% #max=0 if missing, otherwise max=migration flow
  ungroup() %>%
  select(-n) %>%
  distinct(year, sector_dest, state_dest, state_ori, .keep_all = TRUE) %>%
  arrange(year, sector_dest, state_dest, state_ori)
acs <- rbind(acs, ACS_PUMS)
 
c <- c + 1
}
yr_c <- yr_c + 1  
}

#Change to the year that migration happened
acs <- acs %>% 
  mutate(year = year - 1)  
rm(ACS_PUMS)

write.table(acs, file = paste("1-Intermediate_Processed_Data//acs_temp.csv", sep=""), sep = ",", row.names = FALSE)

acs <- read.csv(paste("1-Intermediate_Processed_Data//acs_temp.csv", sep=""), header = TRUE, sep = ",", dec = ".")

#rolling basis: three-year window (to reduce noise due to the fact that each year has relatively few observations)
final_base <- c()
for (yr in 1999:2006) { # the year is changed from 2000-2007 to 1999-2006
temp <- acs %>%
  filter(year >= yr & year <= yr + 2) %>%
  mutate(year = yr)  #now under yr, it actually contains three years obs
final_base <- rbind(final_base, temp) 
}
acs <- final_base
#counting by state and sector, origin and destination on three-year rolling basis
acs <- acs %>%
  group_by(year, sector_dest, state_dest, state_ori) %>%
  mutate(n_acs = sum(n_acs)) %>%  #from 1999 to 2004, n_acs is the sum of consecutive 3 years; for 2005, consecutive 2 years; and for 2006, 1 year.
  ungroup() %>%
  distinct(year, sector_dest, state_dest, state_ori, .keep_all = TRUE)

#Some checks and adjustment (all based on three-year window)
##a. use average total migration flow across years to fill in the missing within state migration
acs <- acs %>%
  mutate(diag = ifelse(state_ori == state_dest, 1, 0)) %>%
  group_by(state_ori, state_dest, sector_dest) %>%
  mutate(n_acs_mean = mean(n_acs)) %>%
  ungroup() %>%
  mutate(n_acs = ifelse(diag == 1 & n_acs == 0, n_acs_mean, n_acs)) %>%
  select(-n_acs_mean, -diag)
##b. given (year, state_ori, sector_dest), the within state_ori migration to that destination sector should be greater than any across state migration from that state_ori to that sector_dest; otherwise, replace it with average across destination states
acs <- acs %>%
  mutate(n_acs_temp = ifelse(state_ori==state_dest, n_acs, 0)) %>%
  group_by(year, state_ori, sector_dest) %>%
  mutate(n_acs_temp = max(n_acs_temp), mean_n_acs = mean(n_acs)) %>%
  ungroup() %>%
  mutate(n_acs = ifelse(n_acs_temp<n_acs, mean_n_acs, n_acs)) %>%
  select(-n_acs_temp, -mean_n_acs)
##c. given (year, state_dest, sector_dest), the within state_dest migration to that destination sector should be greater than any across state migration to that state_dest and to that sector_dest
acs <- acs %>%
  mutate(n_acs_temp = ifelse(state_ori==state_dest, n_acs, 0)) %>%
  group_by(year, state_dest, sector_dest) %>%
  mutate(n_acs_temp = max(n_acs_temp),mean_n_acs = mean(n_acs)) %>%
  ungroup() %>%
  mutate(n_acs = ifelse(n_acs_temp<n_acs, mean_n_acs, n_acs)) %>%
  select(-n_acs_temp, -mean_n_acs)

#Save
write.table(acs, file = paste("1-Intermediate_Processed_Data/acs.csv", sep=""), sep = ",", row.names = FALSE)

```

#IRS migration movements
Compue $L_{IRS}^{n,i}$.
```{r Clean IRS, message=FALSE, warning=FALSE}
# Load IRS data (1999-2000) (to get 1999to2000flow.dta, one must run 0-Raw_Data/IRS/1999to2000CountyMigration/process_raw_1999to2000CountyMigration.dofirst).
irs_temp  <- read_dta("0-Raw_Data/IRS/1999to2000CountyMigration/1999to2000flow.dta")

#Create all state-state combinations.
state_name  <- c("Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut", "Delaware", "Florida", "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky", "Louisiana", "Maine", "Maryland", "Massachusetts", "Michigan", "Minnesota", "Mississippi", "Missouri", "Montana", "Nebraska", "Nevada", "NewHampshire", "NewJersey", "NewMexico", "NewYork", "NorthCarolina", "NorthDakota", "Ohio", "Oklahoma", "Oregon", "Pennsylvania", "RhodeIsland", "SouthCarolina", "SouthDakota", "Tennessee", "Texas", "Utah", "Vermont", "Virginia", "Washington", "WestVirginia", "Wisconsin", "Wyoming")
all_combo <- c()
for (state in state_name) {
temp <- data.frame(state_name)
temp <- temp %>%
  dplyr::rename(state_ori = state_name) %>%
  mutate(state_dest = state)
all_combo <- bind_rows(all_combo, temp)
}
#Replace missing combinations flows with zero, and keep year 1999 only.
irs  <- all_combo %>% 
  left_join(irs_temp, by = c('state_dest' = 'state_dest', 'state_ori' = 'state_ori'))  %>% 
  mutate(year = 1999)  %>% 
  mutate(exemption = ifelse(is.na(exemption), 0, exemption))

#Save.
write.table(irs, file = paste("1-Intermediate_Processed_Data/irs.csv", sep=""), sep = ",", row.names = FALSE) 

```

#CPS migration movements.

Compue $L_{CPS}^{{nj,ik}}$.
```{r Generate cps_nber yearly, message=FALSE, warning=FALSE}
# Load the CPS data. 
cps_panel <-read_dta("0-Raw_Data/CPS/CPS_NBER/Inputs/cps_panel.dta")
# Create a unique identifier for each individual
cps_panel <- cps_panel %>%
  group_by(cpsidp) %>%
  mutate(id = cur_group_id()) %>%
  ungroup()
# Generating 'quarter' within each 'id'
cps_panel <- cps_panel%>%
  arrange(id, month, mish) %>%
  group_by(id) %>%
  mutate(quarter = row_number()) %>%
  ungroup()
#Drop irrelevant variables
cps_panel0<-cps_panel %>%
  select(id, month, year, mish, statefip, ind, wtfinl, quarter, sex, race, bpl, nativity, popstat, age, labforce)
#Keep people that we can observe for at least 4 quarters. 
cps_panel0  <- cps_panel0%>%
  group_by(id) %>%
  mutate(num = n()) %>%
  filter(num >= 4) %>%
  ungroup() %>%
  select(-num)

# Generating 'quarter' column and dropping rows where 'quarter' > 8
cps_panel0  <- cps_panel0 %>%
  filter(quarter <= 8)

#Checking consistency.
# First consistency: each observation's sex constant over the time
cps_panel0 <- cps_panel0 %>%
  arrange(id, sex) %>%
  group_by(id) %>%
  mutate(change = 1000 * sex[1] + 100 * sex[2] + 10 * sex[3] + sex[4]) %>%
  ungroup()
cps_panel0 <-cps_panel0 %>%
  filter(change == 2222 | change == 1111) %>% select(-change, -sex)

# Second consistency: each observation's race constant over time
cps_panel0 <- cps_panel0 %>%
  group_by(id) %>%
   mutate(change2 = ifelse(n_distinct(race) == 1, 1, 0)) %>%
  ungroup() %>%  filter(change2 == 1 )  %>% select(-change2, -race)

# Third consistency: each observation's Nativity (condition of foreign birth place or parentage) constant over time
cps_panel0 <- cps_panel0 %>%
  group_by(id) %>%
   mutate(change3 = ifelse(n_distinct(nativity) == 1, 1, 0)) %>%
  ungroup() %>%  filter(change3 == 1 ) %>% select(-change3, -nativity)

#Fourth consistency: population status ("civilian" or "armed forces") constant over time
cps_panel0 <- cps_panel0 %>%
  group_by(id) %>%
   mutate(change4 = ifelse(n_distinct(popstat) == 1, 1, 0)) %>%
  ungroup() %>%  filter(change4 == 1 )  %>% select(-change4, -popstat)

#Fifth consistency: each year, nobody's age should have change more than one year.
cps_panel0<-cps_panel0[order(cps_panel0$id, cps_panel0$age), ]
cps_panel0 <-cps_panel0 %>%
  arrange(id, age) %>%
   group_by(id) %>%
  mutate(age_difference = max(age) - min(age)) %>%  ungroup() 
cps_panel0  <- cps_panel0 %>% 
    arrange(id, age) %>%
   filter(age_difference <= 3) 

#Sixth consistency: Place of birth constant over time
cps_panel0 <- cps_panel0 %>%
  group_by(id) %>%
   mutate(change6 = ifelse(n_distinct(bpl) == 1, 1, 0))  %>%
  ungroup() %>%  filter(change6 == 1 )

#Mapping sectors (currently, column ind for "industry") to our final classification
cps_panel0$sector <- NA
cps_panel0$sector <- ifelse(cps_panel0$labforce == 1, 0, cps_panel0$sector) #sector 0 (non-participant)
clean_data <- cps_panel0 %>%select(id, month, year, mish, statefip, ind, sector, wtfinl, age) 
clean_data$sector <- ifelse(clean_data$ind %in% c(10, 11, 12, 20, 30, 31, 32), 14, clean_data$sector)
clean_data <- clean_data %>%
  mutate(sector = case_when(
    ind %in% c(100, 101, 102, 110, 111, 112, 120, 121, 122, 130) ~ 1,
    ind %in% c(132, 140, 141, 142, 150, 151, 152) ~ 2,
    ind %in% c(160, 161, 162, 171, 172) ~ 3,
    ind %in% c(200, 201, 40, 41, 42, 50) ~ 4,
    ind %in% c(180, 181, 182, 190, 191, 192) ~ 5,
    ind %in% c(210, 211, 212) ~ 6,
    ind %in% c(250, 251, 252, 261, 262) ~ 7,
    ind %in% c(270, 271, 272, 280, 281, 282, 290, 291 ) ~ 8,
    ind %in% c(300, 301, 310, 311, 312, 320) ~ 9,
    ind %in% c(321, 322, 331, 332, 340, 341, 342, 350) ~ 10,
    ind %in% c(352, 360, 361, 362, 370) ~ 11,
    ind %in% c(372, 380, 381, 390, 391, 392) ~ 12,
    ind %in% c(60, 351, 242, 501, 502, 510, 511, 512, 521, 522, 530, 531, 532, 540, 541, 542, 550, 551, 552, 560, 561, 562, 571, 580, 581, 582 ) ~ 13,
    TRUE ~ sector  # Keep unchanged if not matching any condition
  ))
clean_data <- clean_data %>%
  filter(ind <= 893)#Dropping military and public services
clean_data <- clean_data %>%
  mutate(sector = case_when(
    ind > 582 ~ 13,
    ind %in% c(400, 401, 402, 410, 411, 420, 421, 422, 432, 440, 441, 442, 450, 451, 452, 470, 471, 472) ~ 13,
    TRUE ~ sector
  )) %>%
  filter(!(ind == 412))
clean_data <- clean_data %>%
  mutate(sector = ifelse(ind == 0, 0, sector))

#Age filter 
clean_data  <- clean_data %>%
  filter(age < 66 & age > 24)

#Keep relevant columns.
clean_data <- clean_data %>%
  select(id, month, year, mish, statefip, ind, sector, 
       wtfinl)

#Drop missing values for sectors
clean_data<-clean_data %>%
  filter(!is.na(sector))

#Final changes of format.

#Rename ID, mish, and st_fips; and create abbreviated month column.
clean_data<-clean_data %>%
    rename(P_ID = id) %>%
    mutate(monthname = month.name[month]) %>% 
    select(-month) %>%
   mutate(month = substr(monthname, 1, 3))
clean_data<-clean_data %>%
  select(-monthname) %>%
  rename(inter_num=mish) %>%
  select(-ind)  %>%
  rename(st_fips=statefip)
#Drop Washington DC.
clean_data_new<- left_join(clean_data, states_fips , by="st_fips")
clean_data_new <- clean_data_new %>%
  select(-st_num) %>%
  filter(st_fips != 11)

#Choosing months in sample of interest.
cps_nber <- clean_data_new %>%
  group_by(P_ID) %>%
  filter( min(inter_num) <=4 & max(inter_num) >=5 ) %>%
  ungroup()

## CREATE SECTOR-STATE ORIGINS AND DESTINATIONS.

#list of months
mn_list  <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
# Choose the number of observation for each household
num_obs_present  <- c(5, 6, 7, 8)   # up to: 1 observation c(5); 2 observations c(5,6); 4 observations c(5,6,7,8)
num_obs_past  <- c(1, 2, 3, 4)       # up to: 1 observation c(1); 2 observations c(1,2); 4 observations c(1,2,3,4)
# Choose the restriction on the number of respondents (i.e sample size) within (year, state_ori, sector_ori)
num_sample  <- 0

# Loop to take state-sector origin as state-sector in period t, and state-sector destination as state-sector in period t+1.
base <- c()
for (yr in 1997:2002) {
for (mn in mn_list) {
temp_cps_nber  <- cps_nber  %>% 
  filter(month == mn & (  (year == yr & inter_num  %in% num_obs_present ) | (year == yr-1 & inter_num  %in% num_obs_past ) ))
info_present <- temp_cps_nber %>% 
  filter(year == yr) %>%
  dplyr::rename(sector_dest = sector, state_dest = st_name)
info_past <- temp_cps_nber %>%
  filter(year == yr-1) %>%
  dplyr::select(-year, -month, -wtfinl) %>% 
  dplyr::rename(sector_ori = sector, state_ori = st_name)
temp <- full_join(info_present, info_past, by=c("P_ID"))  %>%
        filter(inter_num.x - inter_num.y == 4) 
temp <- temp %>%
  filter(is.na(state_ori) == FALSE & is.na(state_dest) == FALSE) %>% # only matches (origins consistent with destinations)
  distinct( P_ID, .keep_all = TRUE)
base <- rbind(base, temp)
}
}

# rolling basis: five-year window (to reduce noise due to the fact that each year has relatively few observations)
final_base <- c()
for (yr in 1997:2002) {
temp <- base %>%
  filter(year >= yr - 2 & year <= yr + 2) %>% 
  mutate(year = yr)
final_base <- rbind(final_base, temp)
}
base <- final_base %>% 
  mutate(year = year - 1)   # the original year is the year of the survey at the present; minus one gives the year of survey at the start

# counting sector-sector within state movements (using person weights)
base <- base %>%
  dplyr::select(-month, -P_ID, -inter_num.x, -inter_num.y, -st_fips.x, -st_fips.y) %>%
  group_by(year, state_ori, sector_ori, state_dest, sector_dest) %>%
  mutate(n_cps = sum(wtfinl)) %>%
  ungroup() %>%
  group_by(year, state_ori, sector_ori) %>%
  mutate(n_sample = n()) %>%
  ungroup() %>%
  distinct(year, state_ori, sector_ori, state_dest, sector_dest, .keep_all = TRUE) %>%
  dplyr::select(year, state_ori, sector_ori, state_dest, sector_dest, n_cps, n_sample)

# adding missing combinations
temp <- data.frame(s_names)
sec  <- c(0:14)
temp_sec  <- data.frame(sec)
yr  <- c(1998:2001)
temp_yr  <- data.frame(yr)
combo_st_sec  <- expand.grid(year = temp_yr$yr,state_ori = temp$s_names, sector_ori = temp_sec$sec, state_dest = temp$s_names, sector_dest = temp_sec$sec)
cps_nber  <- combo_st_sec  %>% 
  left_join(base, by = c('year', 'state_dest', 'state_ori','sector_dest','sector_ori'))  %>% 
  mutate(n_cps = ifelse(is.na(n_cps), 0, n_cps))

## Modification: proportionality assumption on diagonals
# We want to set maximum consistent stock in the diagonal (stay at the same state and sector).
temp <- cps_nber  %>% 
  #Binary variable: the observation is on the diagonal: yes or no?
  mutate(diag = ifelse(sector_ori == sector_dest & state_ori == state_dest, 1, 0))  %>% 
  #Sum of people by state-sector origin.
  group_by(year, state_ori, sector_ori)  %>% 
  mutate(stock_ori = sum(n_cps))  %>%  
  #Share of outflow (with respect of total at the origin)
  mutate(outflow_share = n_cps/stock_ori)  %>% 
  #Replace missing values of shares with zeros.
  mutate(outflow_share = ifelse(is.na(outflow_share), 0, outflow_share))  %>% 
  #Shares on the diagonal
  mutate(stay_in_share = ifelse(diag == 1, outflow_share, 0))  %>%
  mutate(stay_in_share = max(stay_in_share))  %>% # no NA; minimum 0.44
  ungroup() %>%
  #Median of shares in the diagonal by year and sector of origin
  group_by(year, sector_ori)  %>% 
  mutate(stay_in_share_median = median(stay_in_share))  %>% 
  ungroup()  %>% 
  #Sum of people by state-sector destination.
  group_by(year, state_dest, sector_dest)  %>% 
  mutate(stock_dest = sum(n_cps))  %>%
  #Leave only sums of people on the diagonal.
  mutate(stock_dest_diag = ifelse(diag == 1, stock_dest, 0))  %>% 
  ungroup()  %>% 
  group_by(year, state_ori, sector_ori)  %>% 
  mutate(stock_dest_diag = max(stock_dest_diag))  %>%
  ungroup()  %>% 
  group_by(year, sector_ori)  %>% 
  #Minimum (positive) stock in the diagonal.
  mutate(stock_dest_diag_min = min(stock_dest_diag[stock_dest_diag > 0]))  %>%   
  ungroup()  %>% 
  #When the "probability" of staying at the same sector-state is 1, assign the corresponding combination the maximum between the minimum stock at the diagonal (according to destination), the stock at the origin, and the stock at the destination. Otherwise, set the maximum between the stock at the origin, the stock at the destination, and a proportionality rule that uses the median of the "probability" of staying at the same sector-state. We apply this replacement for the combinations of the diagonal that do not registered people (n_cps=0).
  mutate(diag_new = ifelse(stay_in_share_median == 1, pmax(stock_dest_diag_min, stock_ori, stock_dest_diag),  pmax(stock_dest_diag_min, stock_ori, stock_dest_diag)/(1-stay_in_share_median) * stay_in_share_median)) %>%
  mutate(n_cps = ifelse(diag == 1 & n_cps == 0, diag_new, n_cps)) %>% 
  select(year, state_ori, sector_ori, state_dest, sector_dest, n_cps) 
cps_nber  <- temp

#filling missing matrix sections
cps_nber <- cps_nber %>%
  mutate(n_master = ifelse(state_ori == state_dest, n_cps, 0)) %>%
  select(-n_cps) %>%
  group_by(year, state_dest, sector_dest, sector_ori) %>% #based on destination state
  mutate(n_cps = max(n_master)) %>%  #n_cps:  each entry takes the value of within destination state migration for that sector-sector movement, therefore same value for same (year, state_dest, sector_dest, sector_ori) across original states
  ungroup() %>%
  select(-n_master) %>%
  arrange(year, state_dest, sector_dest, state_ori, sector_ori)

#Save
write.table(cps_nber, file = paste("1-Intermediate_Processed_Data/cps_nber_2_yearly.csv", sep=""), sep = ",", row.names = FALSE) 

```

# Calculate mobility shares with migration.
For movements between sectors within the same state we use the following proportionality rule:
$$
L^{n j, n k}=L_{I R S}^{n, n} \times \frac{L_{C P S}^{n j, n k}}{\sum_q \sum_h L_{C P S}^{n h, n q}} \quad \forall n \in U S A, \forall j, k
$$
For all the other cases, we use the following proportionality rule:
$$
L^{nj,ik}=\frac{L_{CPS}^{ij,ik}}{\sum_{h}L_{CPS}^{ih,ik}}\times L_{IRS}^{n,i}\times\frac{\sum_{i}L_{ACS}^{n?,ik}}{\sum_{i}\sum_{q}L_{ACS}^{n?,iq}}
$$
Then we calculate the corresponding migration shares:
$$
\mu_{nj,ik}=\frac{L^{nj,ik}}{\sum_{p=1}^{N}\sum_{q=1}^{J}L^{nj,pq}}
$$

```{r Migration matrix solution 2, message=FALSE, warning=FALSE}
# loading intermediate CPS
cps <- read.csv(paste("1-Intermediate_Processed_Data/cps_nber_2_yearly.csv", sep=""), header = TRUE, sep = ",", dec = ".")

# create the CPS term in the within state proportionality rule
temp_within  <- cps %>%
  filter(state_ori == state_dest & year == 1999)  %>% 
  group_by(year, state_dest, state_ori)  %>% 
  mutate(n_cps_two_states_tot = sum(n_cps))  %>% 
  ungroup()  %>% 
  mutate(cps_share_wti = n_cps/n_cps_two_states_tot) %>% 
  select(year, state_ori, state_dest, sector_ori, sector_dest, cps_share_wti)

# create the CPS term in the proportionality rule for all the other cases
temp_between_cps  <- cps  %>%
  filter(state_ori == state_dest & year == 1999)  %>% 
  group_by(year, state_dest, state_ori, sector_dest)  %>% 
  mutate(n_cps_same_sector_dest = sum(n_cps))  %>% 
  ungroup()  %>% 
  mutate(cps_share_btw = n_cps/n_cps_same_sector_dest)  %>% 
  select(year, state_ori, state_dest, sector_ori, sector_dest, cps_share_btw)

# loading intermediate ACS
acs <- read.csv(paste("1-Intermediate_Processed_Data/acs.csv", sep=""), header = TRUE, sep = ",", dec = ".")

# create the ACS term in the proportionality rule for all the other cases
temp_between_acs  <-  acs  %>%
  filter(year == 1999)  %>% 
  group_by(year, state_ori)  %>% 
  mutate(n_acs_two_tot = sum(n_acs))  %>% 
  ungroup()  %>% 
  group_by(year, state_ori, sector_dest)  %>% 
  mutate(n_acs_ori = sum(n_acs))  %>% 
  ungroup()  %>% 
  mutate(acs_share = n_acs_ori/n_acs_two_tot)  %>% 
  mutate(acs_share = ifelse(is.na(acs_share), 0, acs_share))  %>% 
  select(year, state_ori, state_dest, sector_dest, acs_share)

# IRS terms in the formula are already computed
irs <-  read.csv(paste("1-Intermediate_Processed_Data/irs.csv", sep=""), header = TRUE, sep = ",", dec = ".")

#Calculate the final/adjusted migration flows.
mobility  <- cps  %>%
  filter(year == 1999)  %>% 
  left_join(irs, by=c('year' = 'year', 'state_dest' = 'state_dest', 'state_ori' = 'state_ori')) %>%
  left_join(temp_within, by = c('year' = 'year', 'state_dest' = 'state_dest', 'state_ori' = 'state_ori', 'sector_dest' = 'sector_dest', 'sector_ori' = 'sector_ori'))  %>% 
  left_join(temp_between_acs, by = c('year' = 'year', 'state_dest' = 'state_dest', 'state_ori' = 'state_ori', 'sector_dest' = 'sector_dest'))  %>% 
  left_join(temp_between_cps, by = c('year' = 'year', 'state_dest' = 'state_dest', 'sector_dest' = 'sector_dest', 'sector_ori' = 'sector_ori'))  %>% 
  rename(state_ori = state_ori.x)  %>% 
  mutate(n_cps_adj = exemption * acs_share * cps_share_btw)  %>% 
  mutate(n_cps_adj = ifelse(state_ori == state_dest, exemption*cps_share_wti, n_cps_adj))  %>% 
  select(-state_ori.y)

# Calculate mu_{nj,ik}
mu_1999_prime <- mobility %>%
  group_by(year, state_ori, sector_ori) %>% 
  mutate(tot_n = sum(n_cps_adj)) %>%
  ungroup() %>%
  mutate(rel = n_cps_adj/tot_n) %>%
  select(year, state_dest, state_ori, sector_dest, sector_ori, rel)

#Final changes of format (and reshape to matrix form).
mu_1999_prime <- mu_1999_prime %>%
  mutate(sector_dest = recode(as.character(sector_dest), '0' ="00", '1' ="01", '2' ="02", '3' ="03", '4' ="04", '5' ="05", '6' ="06", '7' ="07", '8' ="08", '9' ="09" )) %>%
  mutate(state_dest_sector_dest = paste(state_dest, sector_dest, sep="_")) %>%
  select(-state_dest, -sector_dest)
mu_1999_prime <- spread(mu_1999_prime, state_dest_sector_dest, rel, fill = NA, convert = TRUE,  drop = TRUE, sep = NULL)

# check sum(prob)==1
epsilon <- 0.01
check <- mu_1999_prime %>% select(-year, -state_ori, -sector_ori)
if(sum(abs(as.matrix(rowSums(check)) - matrix(1, dim(check)[1], 1))) > epsilon) 
  stop("mobility shares do not add to 1")

#Save
write_xlsx(mu_1999_prime, path = paste("2-Final_Data/mu_1999.xlsx", sep=""))

```

## Calculate mobility sharess "without migration".

We take CPS $L^{nj,nk}$'s as our final ones.
$$
L^{nj,nk}=L_{CPS}^{nj,nk}\text{,}\forall n\in USA,\:\forall j,k
$$
Then we define
$$
\mu^{nj,nk} =\frac{L^{nj,nk}}{\sum_{q=0}^{J}L^{nj,nq}} \\
\mu^{nj,ik} = 0\quad\forall n\neq i
$$

```{r Migration matrix, message=FALSE, warning=FALSE}
# loading intermediate CPS
cps <- read.csv(paste("1-Intermediate_Processed_Data/cps_nber_2_yearly.csv", sep=""), header = TRUE, sep = ",", dec = ".")
 
# Calculate mu_{nj,ik}
mobility  <- cps %>% 
  mutate(diag = ifelse(state_ori == state_dest, 1, 0))  %>% 
  mutate(n_cps = ifelse(diag == 1, n_cps, 0))  %>% #set between states flows to zero
  group_by(year, state_ori, diag, sector_ori)  %>% 
  mutate(n_cps_tot = sum(n_cps))  %>% 
  ungroup()  %>% 
  mutate(rel = n_cps/n_cps_tot)  %>% 
  mutate(rel = ifelse(is.na(rel), 0, rel))  %>% 
  filter(year == 1999)  %>% 
  select(year, state_dest, state_ori, sector_dest, sector_ori, rel)

#Final changes of format (and reshape to matrix form).
mu_1999_prime <- mobility %>%
  mutate(sector_dest = recode(as.character(sector_dest), '0' ="00", '1' ="01", '2' ="02", '3' ="03", '4' ="04", '5' ="05", '6' ="06", '7' ="07", '8' ="08", '9' ="09" )) %>%
  mutate(state_dest_sector_dest = paste(state_dest, sector_dest, sep="_")) %>%
  select(-state_dest, -sector_dest)
mu_1999_prime <- spread(mu_1999_prime, state_dest_sector_dest, rel, fill = NA, convert = TRUE,  drop = TRUE, sep = NULL)

# check sum(prob)==1
epsilon <- 0.01
check <- mu_1999_prime %>% select(-year, -state_ori, -sector_ori)
if(sum(abs(as.matrix(rowSums(check)) - matrix(1, dim(check)[1], 1))) > epsilon) 
  stop("mobility shares do not add to 1")

#Save
write_xlsx(mu_1999_prime, path = paste("2-Final_Data/mu_1999_no_migration.xlsx", sep=""))

```