---
title: "1-ICIO_VA_shares_io_shares_distances"
output:
  html_notebook:
    toc: yes
    df_print: paged
    number_sections: yes    
  pdf_document:
    number_sections: yes
  word_document: default
header-includes:
- \usepackage{amsmath}
- \usepackage{amssymb}
- \usepackage{pifont}
- \usepackage{mathpazo}
---

## General information

1) Takes ICIO and maps to the 14 relevant sectors that we consider.

2) This code calculates the sector-country bilateral trade flows $X_{ij,k}$ between countries using data from ICIO.

3) This code calculates the share of value-added (mapped in the model to the labor share) for each country, each sector, and each year using data from ICIO. 

4) This code calculates the input output matrix shares for each sector, country and year using data from ICIO. Each entry is the share of expenditure of each buying sector. For each country-year, the sum across rows of each column sums up to one. 

5) The code in this document produces the Country_Coordinates Base that contains information about the most populated cities in each country, the cities' coordinates and population, and each country's population; the previous information is available for the period 1950-2018 for almost all countries except: Austria, Cyprus, Denmark, Estonia, Hungary, Ireland, Lithuania, Slovakia and Slovenia. The missing information is only for the variable, "city population", for certain years. 

6) Calculates the distances between all regions.

7) This code calculates the distance elasticity and own-dummy coefficients for trade flows in services and agriculture between countries (including the US).

## Input files

1. `0-Raw_Data/regions.csv`
2. `0-Raw_Data/sectors_icio.csv`
3. `0-Raw_Data/OECD/ICIO2021_, i,".xlsx`
4. `1-Intermediate_Processed_Data/ICIOFixAgg, i,".csv`
5. `1-Intermediate_Processed_Data/icio_full.csv`
6. `0-Raw_Data//Population_geography//UN_coordinates.xls`
7. `0-Raw_Data//Population_geography//Full_population.xlsx`
8. `0-Raw_Data/Fips/us_states_coordinates_counties.xlsx`
9. `0-Raw_Data/Fips/county_pop_alldata.csv`
10. `1-Intermediate_Processed_Data/us_states_coordinates_counties_2018.csv`
11. `1-Intermediate_Processed_Data/country_coordinates.csv`
12. `0-Raw_Data/Fips/state_codes.csv`
13. `1-Intermediate_Processed_Data/distances.csv`

## Output files

1. `1-Intermediate_Processed_Data/ICIOFixAgg, i,".csv`
2. `1-Intermediate_Processed_Data/icio_full.csv`
3. `1-Intermediate_Processed_Data//ICIO_countries.csv`
4. `1-Intermediate_Processed_Data/country_country_step_.csv`
5. `1-Intermediate_Processed_Data/labor_shares_countries.csv`
6. `1-Intermediate_Processed_Data//value_added_countries", yr, ".csv`
7. `1-Intermediate_Processed_Data/io_shares.csv`
8. `2-Final_Data/io_allyears.xlsx`
9. `1-Intermediate_Processed_Data/us_states_coordinates_counties_2018.csv`
10. `1-Intermediate_Processed_Data//country_coordinates.csv`
11. `1-Intermediate_Processed_Data/distances.csv`



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
remove(list = ls())
```

```{r packages, warning=FALSE, include=FALSE, message=FALSE, results='hide', cache=FALSE}
# libraries
options( scipen=100, stringsAsFactors=FALSE );
# vector of libraries to be used
libs <- c( 'dplyr', 'tidyr', 'readstata13', 'haven', 'stringr', 'nleqslv', 'gdata', 'Statamarkdown', 'formattable', 'ggplot2', 'viridis', 'RColorBrewer', 'reshape2', 'foreign', 'readxl', 'writexl', 'rjson', 'Matrix', 'geodist', 'devtools');
# Install libraries in case they are not installed
for( i in length(libs)){
 if( !(libs[i] %in% installed.packages()) ) install.packages( libs )
};
 install.packages("https://www.ssc.wisc.edu/~hemken/Stataworkshops/Stata%20and%20R%20Markdown/Statamarkdown_0.3.9.zip", repos=NULL)
 install.packages("https://www.ssc.wisc.edu/~hemken/Stataworkshops/Stata%20and%20R%20Markdown/Statamarkdown_0.5.0.tar.gz", type="source", repos=NULL)
 devtools::install_github("Hemken/Statamarkdown")
# load library
lapply( libs, require, character.only=TRUE );
```


# Task 1

## Importing datasets to R

```{r data}
regions <- read.csv(paste("0-Raw_Data/regions.csv", sep=""), header = TRUE, sep = ",", dec = ".")

#sector reclassification table
reclass_sectors <- read.csv(paste("0-Raw_Data/sectors_icio.csv", sep=""), header = TRUE, sep = ",", dec = ".", fileEncoding="UTF-8-BOM")
reclass_sectors <- reclass_sectors %>% 
  select(final_sector, ICIO_sector) %>%
  mutate(ICIO_sector = gsub("D","",ICIO_sector)) %>%
  mutate(ICIO_sector = ifelse(ICIO_sector=="PABR", "DPABR", ICIO_sector))
```


## Fixing ICIO

```{r}

N <- 71   # Number of countries in ICIO
S <- 45   # Number of sectors in ICIO

list <- c("15", "16", "17", "18")
for (i in list){
print(i)
icio_fix <- as.data.frame(read_excel(paste("0-Raw_Data/OECD/ICIO2021_20", i,".xlsx", sep=""), range = "A1:EHJ3196", col_names = TRUE)) 
#names of rows
names_rows <- as.matrix(icio_fix$...1)
colnames(icio_fix)[1] <- " "
#names of cols
names_cols <- t(as.matrix(colnames(icio_fix)))
icio_fix <- icio_fix %>%
  select(-" ") 
icio_fix <- as.matrix(icio_fix)

# Set zero flows equal to a small number
icio_fix[icio_fix == 0] <- 0.0000001   
# unadjusted data on intermediate input flows
Zinit <- icio_fix[ ,1:3195]  
# unadjusted data on both flows of intermediate and final goods
Xinit <- icio_fix       
# unadjusted total output
Rinit <- rowSums(Xinit)                                                      
# Adjust for negative inventory changes. See Handbook chapter (online appendix, page 15, footnote 1) for details.
# We model production as Rinit = A*Rinit + FINsum (where A is calculated by Zinit= A * diag(Rinit))
# With the adjustment below we will have R = A*R + Fsum (where again Z = A * diag(R) with no negative final consumption)

# Matrix of final demand and inventories adjustment
FIN <- Xinit[, (3196:dim(Xinit)[2])]      
# This is only needed for checks
FINsum <- rowSums(FIN)     
# Positive component of final demand (becomes final demand after correcting for INV<0)
F <- FIN
F[F < 0] <- 0 
# Initial final demand (remains the same)
Fsum <- rowSums(F) 
# Estimate matrix on direct input coefficients (small number is added to avoid NaN for zero-sectors)
A <- Zinit %*% solve(diag(Rinit))  
# Compute a new vector of total output under zero decline in inventories
R <- as.vector(solve((diag(N*S)-A),Fsum))    
# Compute a new matrix of intermediate goods flows
diagR <- diag(R)
A <- Matrix(A, sparse=TRUE)
diagR <- Matrix(diagR, sparse=TRUE)
Z <- A %*% diagR    
# Both flows of intermediate and final goods
X <- cbind(Z,F)
X <- as.matrix(X)
# Save data to excel spreadsheet
final <- cbind(names_rows, X)
final <- as.data.frame(final)
colnames(final) <- names_cols
write.table(final, file = paste("1-Intermediate_Processed_Data/ICIOFixAgg", i,".csv", sep=""), sep = ",", row.names = FALSE, col.names = TRUE)
}

rm(Zinit, Z, Xinit, X, icio_fix, FIN, F, A)
```



## Icio full (long format)

```{r}
extra_countries <- c("CHL", "COL", "CRI", "ISL", "ISR", "LVA", "LUX", "NZL", "NOR", "CHE", "ARG", "BRN", "KHM", "HRV", "HKG", "KAZ", "LAO", "MYS", "MLT", "MAR", "MMR", "PER", "PHL", "SAU", "SGP", "ZAF", "THA", "TUN", "VNM")

list <- c("15", "16", "17", "18")
yr <- 2015
final <- c()
for (i in list){
print(yr)
icio_full <- read.csv(paste("1-Intermediate_Processed_Data/ICIOFixAgg", i,".csv", sep=""), header = TRUE, sep = ",", dec = ".")

colnames(icio_full)[1] <- "row"
icio_full <- melt(icio_full, id.vars=c("row"), variable.name="col", value.name = "value")

icio_full <- icio_full %>% 
  mutate(row_country = substr(x=row, start=1, stop=3)) %>%
  mutate(col_country = substr(x=col, start=1, stop=3)) %>%
  mutate(row_item = substr(x=row, start=5, stop=10)) %>%
  mutate(col_item = substr(x=col, start=5, stop=10)) %>%
  mutate(row_country = ifelse(row_country=="MX1" | row_country=="MX2", "MEX", row_country)) %>%
  mutate(col_country = ifelse(col_country=="MX1" | col_country=="MX2", "MEX", col_country)) %>%
  mutate(row_country = ifelse(row_country=="CN1" | row_country=="CN2", "CHN", row_country)) %>%
  mutate(col_country = ifelse(col_country=="CN1" | col_country=="CN2", "CHN", col_country)) %>%
  mutate(row_country = ifelse(row_country %in% extra_countries, "RoW", row_country)) %>%
  mutate(col_country = ifelse(col_country %in% extra_countries, "RoW", col_country)) %>%
  mutate(row_country = ifelse(row_country =="ROW", "RoW", row_country)) %>%
  mutate(col_country = ifelse(col_country =="ROW", "RoW", col_country)) %>%
  mutate(row_country = ifelse(row_country == "ROM", "ROU", row_country)) %>%
  mutate(col_country = ifelse(col_country == "ROM", "ROU", col_country)) %>%
  select(-row, -col) %>%
  mutate(year = yr)

#icio sectors to final sectors   
icio_full <- icio_full %>%
  left_join(reclass_sectors, by=c('row_item'='ICIO_sector')) %>%
  select(-row_item) %>%
  dplyr::rename(row_item = final_sector) %>%
  left_join(reclass_sectors, by=c('col_item'='ICIO_sector')) %>%
  select(-col_item) %>%
  dplyr::rename(col_item = final_sector) %>%
  filter(is.na(row_item)==FALSE & is.na(col_item)==FALSE) %>%
  group_by(year, row_country, col_country, row_item, col_item) %>%
  mutate(tot_value=sum(value)) %>%
  ungroup() %>%
  distinct(year, row_country, col_country, row_item, col_item, .keep_all = TRUE) %>%
  select(-value) %>%
  dplyr::rename(value = tot_value) %>%
  select(year, row_country, col_country, row_item, col_item, value) %>%
  arrange(year, row_country, col_country, row_item, col_item)
  
final <- rbind(final, icio_full)  
yr <- yr + 1

}

write.table(final, file = paste("1-Intermediate_Processed_Data//icio_full.csv", sep=""), sep = ",", row.names = FALSE)
rm(icio_full, final)

```


## ICIO in matrix form

```{r}

regions <- regions %>% 
  filter(status == "country") %>%
  mutate(number = ifelse(region == "USA", 1, number)) %>%
  select(-status) 

icio_full <- read.csv(paste("1-Intermediate_Processed_Data//icio_full.csv", sep=""), header = TRUE, sep = ",", dec = ".")

#collapsing by sector origin
icio_matrix <- icio_full %>%
  dplyr::rename(sector = row_item) %>%
  group_by(year, sector, row_country, col_country) %>%
  mutate(tot_value = sum(value)) %>%
  ungroup() %>%
  distinct(year, sector, row_country, col_country, .keep_all = TRUE)
  
icio_matrix <- icio_matrix %>%  
  left_join(regions, by=c('row_country'='region')) %>%
  dplyr::rename(exporter=number) %>%
  left_join(regions, by=c('col_country'='region')) %>%
  dplyr::rename(importer = number) %>%
  dplyr::rename(importer_country = col_country) %>%
  dplyr::rename(exporter_country = row_country) %>%
  arrange(year, sector, importer, exporter) %>%
  select(-col_item, -exporter, -value)
icio_matrix$exporter_country <- paste("value", icio_matrix$exporter_country, sep="_")  
icio_matrix <- spread(icio_matrix, exporter_country, tot_value, fill = NA, convert = TRUE,  drop = TRUE, sep = NULL)
icio_matrix <- icio_matrix %>% select(year, importer_country, sector, importer, value_USA, everything())
icio_matrix <- icio_matrix %>% arrange(year, sector, importer)

write.table(icio_matrix, file = paste("1-Intermediate_Processed_Data//ICIO_countries.csv", sep=""), sep = ",", row.names = FALSE)

```

# Task 2

## Importing datasets to R

```{r}
regions <- read.csv("0-Raw_Data/regions.csv")
c_names <- regions %>% filter(status == "country")
c_names <- c_names$region

s_names <- regions %>% filter(status == "state") 
s_names <- s_names$region

n_s <- length(s_names)
n_c <- length(c_names) -1


# ICIO 
icio <- icio_full
```

## Trade flows between countries

```{r}
#colapsing by origin sectors 
icio <- icio %>%
  group_by(year, row_country, col_country, row_item) %>%
  mutate(val = sum(value)) %>%
  ungroup() %>%
  distinct(year, row_country, col_country, row_item, .keep_all = TRUE) %>%
  select(-value, -col_item) %>%
  dplyr::rename(value = val)
colnames(icio) <- c("year", "exporter", "importer", "sector", "value")  
icio <- spread(icio, exporter, value, fill = NA, convert = TRUE,  drop = TRUE, sep = NULL)
icio <- icio %>% arrange(year, sector, importer)
icio <- data.frame(icio)

for (yr in 2015:2018) {
  Step1 <- icio %>%
    filter(year == yr) %>%
    select(year, importer, sector, everything())
  #Exporting
  write.table(Step1, file = paste("1-Intermediate_Processed_Data//country_country_step_", yr,".csv", sep=""), sep = ",", row.names = FALSE)
}
```


# Task 3

## Importing datasets to R
```{r, warning=FALSE}

icio_inter <- icio_full %>%
  filter(row_item != 0, col_item != 0) #NOT CONSUMPTION

```


## Value added and labor shares for countries

Gross output for each region, for each sector

$$R_{i,k}= \sum_j X_{ij,k} \quad \forall j$$

```{r gross-output, warning=FALSE}
R <- icio_full %>%
  group_by(year, row_country, row_item) %>%
  mutate(val = sum(value)) %>%
  ungroup() %>%
  distinct(year, row_country, row_item, .keep_all = TRUE) %>%
  select(-col_country, -col_item, -value) %>%
  dplyr::rename(sector = row_item, region = row_country) %>%
  arrange(year, sector, region)

```

Calculating value added:

$$VA_{i,k}= R_{i,k}- \sum_s X_{j,sk} $$
```{r value-added, warning=FALSE}
#calculating value added for countries
sum_s_X_jsk <- icio_inter %>%
  group_by(year, col_country, col_item) %>%
  mutate(val = sum(value)) %>%
  ungroup() %>%
  distinct(year, col_country, col_item, .keep_all = TRUE) %>%
  select(-row_country, -row_item, -value) %>%
  dplyr::rename(region = col_country, sector = col_item) %>%
  arrange(year, sector, region)
VA <- R$val - sum_s_X_jsk$val
VA <- cbind(R[1:3], VA)
colnames(VA) <- c("year", "region", "sector", "value_added")

VA<-spread(VA, sector, value_added, fill = NA, convert = TRUE,  drop = TRUE, sep = NULL)
R <-spread(R, sector, val, fill = NA, convert = TRUE,  drop = TRUE, sep = NULL)
```

Calculating the share of value added in gross output:

$$\alpha_{i,k} = \frac{VA_{i,k}}{R_{i,k}}$$
```{r value-added-share, warning=FALSE}
labor_shares <- VA[, 3:dim(VA)[2]] / R[, 3:dim(R)[2]]
labor_shares[labor_shares>1] <- 1
labor_shares[labor_shares<(-1)] <- (-1)
labor_shares <- cbind(VA[, 1:2], labor_shares)

for (yr in 2015:2018) {
  VA_c <- VA %>% filter(year == yr)
  write.table(VA_c, file = paste("1-Intermediate_Processed_Data//value_added_countries", yr, ".csv", sep=""), sep = ",", row.names = FALSE)

  final <- labor_shares %>% filter(year == yr)
  write.table(final, file = paste("1-Intermediate_Processed_Data//labor_shares_countries", yr, ".csv", sep=""), sep = ",", row.names = FALSE)
}


```


# Task 4

## Constructing Input-Output matrices for countries

Define $\alpha_{j,ks}$ for region $j$ as the share of purchases of sector $s$ that come from sector $k$ in the total purchases of sector $s$ (this is the input output coefficient). We have that:

$$\alpha_{j,ks}=\dfrac{\sum_{i}X_{ij,ks}}{\sum_{r}\sum_{i}X_{ij,rs}}=\dfrac{X_{j,ks}}{\sum_{r}X_{j,rs}}$$

Note that $\sum_{k}\alpha_{j,ks}=1\quad\forall j$. 


```{r}
## Calculating shares

num <- icio_inter %>%
  group_by(year, col_country, col_item, row_item) %>%
  mutate(val = sum(value)) %>%
  ungroup() %>%
  distinct(year, col_country, col_item, row_item, .keep_all = TRUE) %>%
  select(-row_country, -value) %>%
  dplyr::rename(region = col_country, sector_o = row_item, sector_d = col_item)

num <- spread(num, sector_o, val, fill = NA, convert = TRUE,  drop = TRUE, sep = NULL)

io_shares <- num[, 4:dim(num)[2]] / rowSums(num[, 4:dim(num)[2]])
io_shares <- cbind(num[, 1:3], io_shares)

io <- melt(io_shares, id.vars=c("year", "region", "sector_d"), variable.name="sector_o", value.name = "share")
id_tot <- data.frame(c_names)
id_tot$id <- c(1:(n_c + 1))
colnames(id_tot) <- c("regions", "id")

io <- io %>%
  left_join(id_tot, by=c('region'='regions')) %>%
  mutate(region = id) %>%
  select(-id) %>%
  arrange(year, region, sector_d)

io <-spread(io, sector_d, share, fill = NA, convert = TRUE,  drop = TRUE, sep = NULL)
io <- io %>% 
  dplyr::rename(sector = sector_o) %>%
  left_join(id_tot, by=c('region'='id')) %>%
  mutate(region = regions) %>%
  select(-regions)
io$sector <- as.numeric(as.character(io$sector))
# #check: 8x38=304
# colSums(io[, 4:dim(io)[2]])

for (yr in 2015:2018) {
  final <- io %>%
    filter(year == yr)
  write.table(final, file = paste("1-Intermediate_Processed_Data//io_shares", yr, ".csv", sep=""), sep = ",", row.names = FALSE)
}

```

## Exporting all IO-Tables in a unique file
```{stata, collectcode=TRUE}

forvalue i=2015/2018{

import delimited "1-Intermediate_Processed_Data/io_shares`i'.csv", clear
rename region country
ds year country sector, not
local vari `r(varlist)'
local y=101
replace sector=sector+100
foreach v of local vari{
rename `v' io`y'
format io`y' %25.24f
local y=`y'+1
}
order year country sector io*, first
 
export excel using "2-Final_Data/io_allyears.xlsx" ///
	, sheet("IO`i'") firstrow(variables) sheetmodify
}	

 
```

# Task 5

Almost all information to construct the final base comes from data sets of the UN; specifically, the inputs are the following:

a) United Nations: Population Division. "Population of Urban Agglomerations with 300,000 Inhabitants or More in 2018, by country, 1950-2035 (thousands)." Base found in: https://population.un.org/wup/Download/

b) United Nations: Population Division. "Total Population - Both Sexes. De facto population in a country, area or region as of 1 July of the year indicated. Figures are presented in thousands." Base found in: https://population.un.org/wpp/Download/Standard/Population/

c) Republic of Cyprus: Statistical Service. "POPULATION BY DISTRICT, 1996-2017." Base found in: https://www.mof.gov.cy/mof/cystat/statistics.nsf/populationcondition_21main_en/populationcondition_21main_en?OpenForm&sub=1&sel=2

The final output is based mainly on input a), with some information of b) and c), since the countries named before didn't meet the requirement of providing information of at least two cities. For all these special cases (with the exception of Cyprus), b) gives the information needed since it has the cities' population by sex (we just take the aggregation of male and female); from b) we take the urban agglomeration category when possible to make it comparable with the main UN base we are using. For the specific case of Cyprus, the cities' information comes from the country's Statistical Service; since the population statistics available refer to the provinces (named after the main city in them), and not to the city itself, we take the province's urban population and not the whole province population. Finally, regarding the output, all population statistics have to be read in thousands of humans beings.  

***Note that for the case of Cyprus, a more precise statistic for the cities' population can be given through municipalities' population instead of districts (province), but only for 2011.  

## Importing dataset to R

```{r}

#updating US counties population from 2010 to 2018
coordinates_counties_all <- read.csv(paste("0-Raw_Data/Fips/county_pop_alldata.csv", sep=""), header = TRUE, sep = ",", dec = ".")
coordinates_counties_all <- coordinates_counties_all %>%
  mutate(FIPS = STATE*1000 + COUNTY) %>%
  dplyr::rename(Population_2018 = POPESTIMATE2018) %>%
  select(FIPS, Population_2018)

coordinates_counties <-as.data.frame(read_excel(paste("0-Raw_Data/Fips/us_states_coordinates_counties.xlsx", sep="")))
coordinates_counties <- coordinates_counties %>%
  select(-Population_2010) %>%
  left_join(coordinates_counties_all, by=c('FIPS'='FIPS')) %>%
  filter(is.na(Population_2018)==FALSE)

write.table(coordinates_counties,file=paste("1-Intermediate_Processed_Data//us_states_coordinates_counties_2018.csv", sep=""), sep = ",", row.names = FALSE)

#######

coordinates_counties <- read.csv(paste("1-Intermediate_Processed_Data/us_states_coordinates_counties_2018.csv", sep=""), header = TRUE, sep = ",", dec = ".")
state_code <- read.csv(paste("0-Raw_Data/Fips/state_codes.csv", sep=""), header = TRUE, sep = ",", dec = ".")
state_code$state <- gsub(" ", "", state_code$state, fixed = TRUE)

UN <- read_excel("0-Raw_Data//Population_geography//UN_coordinates.xls")
UN <- data.frame(UN)
UN <- UN %>%
  dplyr::rename(country=Country.or.area, city=Urban.Agglomeration)

regions <- read.csv("0-Raw_Data/regions.csv")
c_names <- regions %>% filter(status=="country", region!="USA")
c_names <- c_names$region

s_names <- regions %>% filter(status =="state") 
s_names <- s_names$region

n_s <- length(s_names)
n_c <- length(c_names)

#Full population
Full_pop <- read_excel("0-Raw_Data//Population_geography//Full_population.xlsx")
Full_pop <- data.frame(Full_pop)

all_country_names=c("Aruba", "Afghanistan", "Angola", "Anguilla", "Aland Islands", "Albania", "Andorra", "United Arab Emirates", "Argentina", "Armenia", "American_Samoa", "Antarctica", "French Southern Territories", "Antigua and Barbuda", "Australia", "Austria", "Azerbaijan", "Burundi", "Belgium", "Benin", "Bonaire, Sint Eustatius and Saba", "Burkina_Faso", "Bangladesh", "Bulgaria", "Bahrain", "Bahamas", "Bosnia and Herzegovina", "Saint Barthelemy", "Belarus", "Belize", "Bermuda", "Bolivia", "Brazil", "Barbados", "Brunei Darussalam", "Bhutan", "Bouvet Island", "Botswana", "Central African Republic", "Canada", "Cocos Islands", "Switzerland", "Chile", "China", "Cote d Ivoire", "Cameroon", "Democratic Republic of the Congo", "Congo", "Cook Islands", "Colombia", "Comoros", "Cabo Verde", "Costa Rica", "Cuba", "Curacao", "Christmas Island", "Cayman Islands", "Cyprus", "Czechia", "Germany", "Djibouti", "Dominica", "Denmark", "Dominican Republic", "Algeria", "Ecuador", "Egypt", "Eritrea", "Western Sahara", "Spain", "Estonia", "Ethiopia", "Finland", "Fiji", "Falkland Islands", "France", "Faroe Islands", "Micronesia", "Gabon", "United Kingdom", "Georgia", "Guernsey", "Ghana", "Gibraltar", "Guinea", "Guadeloupe", "Gambia", "Guinea-Bissau", "Equatorial Guinea", "Greece", "Grenada", "Greenland", "Guatemala", "French Guiana", "Guam", "Guyana", "Hong Kong", "Heard Island and McDonald Islands", "Honduras", "Croatia", "Haiti", "Hungary", "Indonesia", "Isle of Man", "India", "British Indian Ocean Territory", "Ireland", "Iran", "Iraq", "Iceland", "Israel", "Italy", "Jamaica", "Jersey", "Jordan", "Japan", "Kazakhstan", "Kenya", "Kyrgyzstan", "Cambodia", "Kiribati", "Saint Kitts and Nevis", "South Korea", "Kuwait", "Laos", "Lebanon", "Liberia", "Libya", "Saint Lucia", "Liechtenstein", "Sri Lanka", "Lesotho", "Lithuania", "Luxembourg", "Latvia", "Macao", "Saint Martin", "Morocco", "Monaco", "Moldova", "Madagascar", "Maldives", "Mexico", "Marshall Islands", "North Macedonia", "Mali", "Malta", "Myanmar", "Montenegro", "Mongolia", "Northern Mariana Islands", "Mozambique", "Mauritania", "Montserrat", "Martinique", "Mauritius", "Malawi", "Malaysia", "Mayotte", "Namibia", "New Caledonia", "Niger", "Norfolk Island", "Nigeria", "Nicaragua", "Niue", "Netherlands", "Norway", "Nepal", "Nauru", "New Zealand", "Oman", "Pakistan", "Panama", "Pitcairn", "Peru", "Philippines", "Palau", "Papua New Guinea", "Poland", "Puerto Rico", "North Korea", "Portugal", "Paraguay", "Palestine", "French Polynesia", "Qatar", "Reunion", "Romania", "Russia", "Rwanda", "Saudi Arabia", "Sudan", "Senegal", "Singapore", "South Georgia and the South Sandwich Islands", "Saint Helena, Ascension and Tristan da Cunha", "Svalbard and Jan Mayen", "Solomon Islands", "Sierra Leone", "El Salvador", "San Marino", "Somalia", "Saint Pierre and Miquelon", "Serbia", "South Sudan", "Sao Tome and Principe", "Suriname", "Slovakia", "Slovenia", "Sweden", "Eswatini", "Sint Maarten", "Seychelles", "Syria", "Turks and Caicos Islands", "Chad", "Togo", "Thailand", "Tajikistan", "Tokelau", "Turkmenistan", "Timor-Leste", "Tonga", "Trinidad and Tobago", "Tunisia", "Turkey", "Tuvalu", "Taiwan", "Tanzania", "Uganda", "Ukraine", "United States Minor Outlying Islands", "Uruguay", "United States of America", "Uzbekistan", "Holy See", "Saint Vincent and the Grenadines", "Venezuela", "British Virgin Islands", "American Virgin Islands", "Vietnam", "Vanuatu", "Wallis and Futuna", "Samoa", "Yemen", "South Africa", "Zambia", "Zimbabwe")

all_country_codes=c("ABW", "AFG", "AGO", "AIA", "ALA", "ALB", "AND", "ARE", "ARG", "ARM", "ASM", "ATA", "ATF", "ATG", "AUS", "AUT", "AZE", "BDI", "BEL", "BEN", "BES", "BFA", "BGD", "BGR", "BHR", "BHS", "BIH", "BLM", "BLR", "BLZ", "BMU", "BOL", "BRA", "BRB", "BRN", "BTN", "BVT", "BWA", "CAF", "CAN", "CCK", "CHE", "CHL", "CHN", "CIV", "CMR", "COD", "COG", "COK", "COL", "COM", "CPV", "CRI", "CUB", "CUW", "CXR", "CYM", "CYP", "CZE", "DEU", "DJI", "DMA", "DNK", "DOM", "DZA", "ECU", "EGY", "ERI", "ESH", "ESP", "EST", "ETH", "FIN", "FJI", "FLK", "FRA", "FRO", "FSM", "GAB", "GBR", "GEO", "GGY", "GHA", "GIB", "GIN", "GLP", "GMB", "GNB", "GNQ", "GRC", "GRD", "GRL", "GTM", "GUF", "GUM", "GUY", "HKG", "HMD", "HND", "HRV", "HTI", "HUN", "IDN", "IMN", "IND", "IOT", "IRL", "IRN", "IRQ", "ISL", "ISR", "ITA", "JAM", "JEY", "JOR", "JPN", "KAZ", "KEN", "KGZ", "KHM", "KIR", "KNA", "KOR", "KWT", "LAO", "LBN", "LBR", "LBY", "LCA", "LIE", "LKA", "LSO", "LTU", "LUX", "LVA", "MAC", "MAF", "MAR", "MCO", "MDA", "MDG", "MDV", "MEX", "MHL", "MKD", "MLI", "MLT", "MMR", "MNE", "MNG", "MNP", "MOZ", "MRT", "MSR", "MTQ", "MUS", "MWI", "MYS", "MYT", "NAM", "NCL", "NER", "NFK", "NGA", "NIC", "NIU", "NLD", "NOR", "NPL", "NRU", "NZL", "OMN", "PAK", "PAN", "PCN", "PER", "PHL", "PLW", "PNG", "POL", "PRI", "PRK", "PRT", "PRY", "PSE", "PYF", "QAT", "REU", "ROU", "RUS", "RWA", "SAU", "SDN", "SEN", "SGP", "SGS", "SHN", "SJM", "SLB", "SLE", "SLV", "SMR", "SOM", "SPM", "SRB", "SSD", "STP", "SUR", "SVK", "SVN", "SWE", "SWZ", "SXM", "SYC", "SYR", "TCA", "TCD", "TGO", "THA", "TJK", "TKL", "TKM", "TLS", "TON", "TTO", "TUN", "TUR", "TUV", "TWN", "TZA", "UGA", "UKR", "UMI", "URY", "USA", "UZB", "VAT", "VCT", "VEN", "VGB", "VIR", "VNM", "VUT", "WLF", "WSM", "YEM", "ZAF", "ZMB", "ZWE") 

codes <- data.frame(cbind(all_country_names, all_country_codes))

```


## Verification of how many cities for each country

```{r cities}

UN_0 <- UN
UN_0 <- UN_0 %>%
  dplyr::rename(Country.code.num = Country.Code, Country.name = country) %>%
  select(Index, Country.code.num, Country.name, everything())

#Verification of how many cities for each country
UN_0 <- UN_0 %>%
  left_join(codes, by=c('Country.name'='all_country_names')) %>%
  dplyr::rename(country = all_country_codes) %>%
  arrange(country) 

city_count <- UN_0 %>% #doesn't include count for RoW
  filter(country %in% c_names) %>%
  mutate(c = 1) %>%
  group_by(country) %>%
  mutate(city.count = sum(c)) %>%
  ungroup() %>%
  distinct(country, .keep_all = TRUE) %>%
  select(country, city.count)

```

## Coordinates

```{r main-variables}

UN_1 <- UN_0 %>%
  select(-Index, -Note, -City.Code)

UN_1 <- data.frame(melt(UN_1, id.vars = c("Country.code.num", "Country.name", "country", "city", "Latitude", "Longitude"),  variable.name ="year", value.name = "city.population"))
UN_1 <- UN_1 %>% select(year, everything())
UN_1$year <- substr(UN_1$year, 2, 5)


#Incorporation of country's population
Full_pop1 <- data.frame(melt(Full_pop, id.vars = "Country", variable.name ="year", value.name = "country.population"))
Full_pop1 <- Full_pop1 %>% select(year, everything())
Full_pop1$year <- substr(Full_pop1$year, 2, 5)

UN_1 <- UN_1 %>%
  left_join(Full_pop1, by=c('year'='year', 'Country.name'='Country')) 

#Verification that none NA is left in the variable: country.population
UN_2 <- UN_1 %>%
  filter(is.na(country.population))

UN_special <- UN_1

UN_1 <- UN_1 %>% mutate(country.name.original = Country.name, country.population.original = country.population)
UN_1 <- UN_1 %>% select(year, Country.code.num, Country.name, country.name.original, everything())


#Change other countries to Rest of the World (RoW)
countries_states <- c("USA", c_names)
UN_1 <- UN_1 %>%
  mutate(Country.name= ifelse(country %in% countries_states, Country.name, "Rest of the World")) %>%
  mutate(country = ifelse(country %in% countries_states, country, "RoW"))

#Total population for RoW
UN_temp <- UN_special %>%
  distinct(year, country, .keep_all = TRUE) %>%
  mutate(Country.name= ifelse(country %in% countries_states, Country.name, "Rest of the World")) %>%
  mutate(country = ifelse(country %in% countries_states, country, "RoW")) %>%
  filter(country == "RoW") %>%
  group_by(year) %>%
  mutate(tot.pop.RoW=sum(country.population)) %>%
  ungroup() %>%
  distinct(year, .keep_all = TRUE) %>%
  select(year, country, tot.pop.RoW)

UN_1 <- UN_1 %>%
  left_join(UN_temp, by=c('year'='year', 'country'='country')) %>%
  mutate(country.population= ifelse(country == "RoW", tot.pop.RoW, country.population)) %>%
  select(-tot.pop.RoW) %>%
  arrange(year, Country.code.num)

write.table(UN_1, file = paste("1-Intermediate_Processed_Data//country_coordinates.csv", sep=""), sep = ",", row.names = FALSE)


```

# Task 6 

## Distances between all regions

```{r}
country_coordinates <- UN_1

state_i <- data.frame(s_names)
colnames(state_i) <- c("iso_o")
temp <- data.frame(c_names)
colnames(temp) <- c("iso_o")
state_i <- rbind(state_i, temp)
state_i$state_i <- as.numeric(row.names(state_i))
colnames(state_i) <- c("iso_o", "state_i")
state_j <- state_i %>%
  dplyr::rename(state_j = state_i, iso_d = iso_o)


#	1. Imports coordinates by county and crosses each county with the others
# States   ***
state_0 <- coordinates_counties %>%
  filter(State != "DC") %>%
  mutate(code = State) %>%
  left_join(state_code, by=('code'='code')) %>%
  mutate(State = state) %>%
  mutate(state = as.integer(factor(State))) %>%
  rename(fips = FIPS)
 

# POPULATIONS
# COUNTRIES
country_coor <- country_coordinates %>%
  filter(country.name.original != "United States of America", year == 2016) %>%
  mutate(city.population = ifelse(country == "IRL" & city == "Cork", 124.391, city.population)) %>%  # missing value
  mutate(state = as.integer(factor(country)) + 50) %>% ##ojo RoW y RUS
  mutate(city = ifelse(country == "RoW", paste(country.name.original,city, sep="_"), city)) %>%
  mutate(city = ifelse(city == "Xinyi" & round(Latitude)==22, "Xinyi2", city)) %>%
  mutate(country_city = paste(country,city, sep="_")) %>%
  mutate(fips = as.integer(factor(country_city))) %>%
  dplyr::rename(pop = city.population) %>%
  mutate(fips = fips*1000000) %>%
  mutate(pop = pop/country.population)
country_pop <- country_coor %>%
  select(state, fips, pop)
country_coor <- country_coor %>%
  select(state, fips,  Latitude, Longitude) %>%
  dplyr::rename(lat = Latitude, lon = Longitude)

# STATES
pops_i <- state_0 %>%
  select(state, fips, Population_2018) %>%
  dplyr::rename(pop = Population_2018) %>%
  group_by(state) %>%
  mutate(tot_pop = sum(pop)) %>%
  ungroup() %>%
  mutate(pop = pop/tot_pop) %>%
  select(-tot_pop)
# appending countries
pops_i <- rbind(pops_i, country_pop)
pops_i <- pops_i %>%
  dplyr::rename(state_i = state, fips_r_in_i = fips, pop_r_i = pop)
pops_j <- pops_i %>%
  dplyr::rename(state_j = state_i, fips_s_in_j = fips_r_in_i, pop_s_j = pop_r_i)
pops_i <- data.frame(pops_i)
pops_j <- data.frame(pops_j)

#  COORDINATES   ***   
cities <- state_0 %>%
  select(state, fips, Latitude, Longitude, State, County) %>%
  dplyr::rename(lat = Latitude, lon = Longitude) %>%
  mutate(lon = as.numeric(substr(x=lon, start=2, stop=10))) %>% 
  mutate(lon = -lon) %>%
  mutate(lon = ifelse( State == "Alaska" & County == "Aleutians West[4]", 178.33881299999999, lon)) %>%	
  select(-State, -County)
cities <- rbind(cities, country_coor) 
colnames(cities) <- c("state_s", "fips_s", "lat_s", "lon_s")

cities_lat_lon <- as.matrix(cities %>% select(lat_s, lon_s))
cities_no_s <- data.frame(cities)
colnames(cities_no_s) <- c("state", "fips", "lat", "lon")

#	2. Calculates distance in km
# generate distances in km between cities
dist <- geodist (cities_lat_lon, measure = "geodesic")
dist <- data.frame(dist)
colnames(dist) <- paste(cities$state_s,cities$fips_s, sep="_")
dist <- cbind(cities, dist)

dist <- melt(dist, id.vars=c("state_s", "fips_s", "lat_s", "lon_s"), variable.name="state_fips", value.name = "d_rs")

dist <- dist %>%
  mutate(state = substr(x=state_fips, start=1, stop=2)) %>%
  mutate(fips = substr(x=state_fips, start=3, stop=15))
dist$state <- gsub("_", "", dist$state, fixed = TRUE)
dist$fips <- gsub("_", "", dist$fips, fixed = TRUE)
dist <- dist %>%
  mutate(state = as.numeric(state), fips = as.numeric(fips)) %>%
  arrange(state, state_s, fips, fips_s) %>%
  select(state, state_s, fips, fips_s, d_rs) 

#	3. Renames variables
dist <- dist %>%
  dplyr::rename(state_i = state, state_j = state_s, fips_r_in_i = fips, fips_s_in_j = fips_s)

#	4. Merging populations
dist <- dist %>%
  left_join(pops_i, by=c('state_i'='state_i', 'fips_r_in_i'='fips_r_in_i')) %>%
  left_join(pops_j, by=c('state_j'='state_j', 'fips_s_in_j'='fips_s_in_j'))

#	5. Applying formula
dist <- dist %>%
  mutate(theta = -1) %>%
  mutate(dist = (pop_r_i*pop_s_j*(d_rs^(theta)))) %>%
  mutate(dist = ifelse(d_rs == 0, 0, dist)) %>%
  group_by(state_i, state_j) %>%
  mutate(dist = sum(dist)) %>%
  ungroup() %>%
  distinct(state_i, state_j, .keep_all = TRUE) %>%
  mutate(dist1 = (dist^(1/theta))) %>%
  mutate(dist1 = ifelse(dist == 0, 0, dist1)) %>%
  select(-dist) %>%
  dplyr::rename(dist = dist1) %>%
  mutate(dist = dist/1000)

# Merging back states
dist <- dist %>%
  left_join(state_i, by=c('state_i'='state_i')) %>%
  left_join(state_j, by=c('state_j'='state_j')) %>%
  select(iso_o, iso_d, dist) %>%
  arrange(iso_o, iso_d)

final <- c()
for (yr in 2015:2018){
dist$year <- yr  
final <- rbind(final, dist)  
}


write.table(final, file = paste("1-Intermediate_Processed_Data/distances.csv", sep=""), sep = ",", row.names = FALSE)


```


# Task 7

To do this, we estimate:
$$
\ln X_{ij,t}=\lambda_t + \delta_{i}^{o}+\delta_{j}^{d}+\beta_{0}\iota_{ij}+\beta_{1}\ln dist_{ij}+\xi_{ij,t}
$$
where $\lambda_t$ is a time FE, $\delta_{i}^{o}$ is an origin FE, $\delta_{j}^{d}$ is a destination FE, and $\iota_{ij}$ is an indicator variable equal to 1 if $i=j$, and zero otherwise. As usual $X_{ij,t}$ is what country $i$ sells to country $j$ in year $t$. The coefficients of interest are $\beta_0$ and $\beta_1$ that we use to construct: $$\tilde{\tau}_{ij}=\exp(\hat{\beta}_{0}\iota_{ij}+\hat{\beta}_{1}\ln dist_{ij}).$$

We focus on countries only (including the US), and our focus is on separate regressions for services and agriculture.

## Importing dataset to R

```{r}
regions <- read.csv("0-Raw_Data/regions.csv")
regions$region <- as.character(regions$region)
c_names <- regions %>% filter(status == "country", region != "USA")
c_names <- c_names$region

s_names <- regions %>% filter(status == "state") 
s_names <- s_names$region

n_s <- length(s_names)
n_c <- length(c_names)

reclass_sectors <- read.csv("0-Raw_Data/sectors_icio.csv", fileEncoding="UTF-8-BOM")
n_sec <- length(unique(reclass_sectors$final_sector)) -4

reclass_sectors <- read.csv("0-Raw_Data/sectors_icio.csv", fileEncoding="UTF-8-BOM")
reclass_sectors <- reclass_sectors %>% 
  select(final_sector, ICIO_sector)

#ICIO
icio <- read.csv("1-Intermediate_Processed_Data//icio_full.csv")

#Trade distances
data_services <- read.csv("1-Intermediate_Processed_Data//distances.csv")
data_services <- data.frame(data_services)
data_servicesUS <- data_services %>% filter(year == 2018)

#US population in 2018
us_pop <- read.csv(paste("1-Intermediate_Processed_Data/us_states_coordinates_counties_2018.csv", sep=""), header = TRUE, sep = ",", dec = ".")
us_pop <- data.frame(us_pop)

#Trade distances
data_agriculture <- read.csv("1-Intermediate_Processed_Data//distances.csv")
data_agriculture <- data.frame(data_agriculture)
data_agricultureUS <- data_agriculture %>% filter(year == 2018)

#US population in 2018
us_pop <- read.csv(paste("1-Intermediate_Processed_Data/us_states_coordinates_counties_2018.csv", sep=""), header = TRUE, sep = ",", dec = ".")
us_pop <- data.frame(us_pop)
```

## Gravity for services
  
```{r services}
#only sector 13: origin
icio1 <- icio %>%
  filter(row_item == n_sec + 1) %>%
  group_by(year, row_country, col_country) %>%
  mutate(ag_value = sum(value)) %>% 
  ungroup() %>%  
  distinct(year, row_country, col_country, .keep_all=TRUE ) %>% 
  select(-col_item, -value, -row_item) %>%
  mutate(iota = 0) %>%
  mutate(iota = replace(iota, row_country == col_country, 1)) %>%
  unite(ori_dest, c("row_country", "col_country"), remove = FALSE) %>%
  mutate(dist = 0)

#preparing distances 
data_services1 <- data_services %>%
  unite(ori_dest, c("iso_o", "iso_d"), remove = FALSE) %>%
  mutate(ch_o = nchar(as.character(iso_o))) %>%
  mutate(ch_d = nchar(as.character(iso_d))) %>%
  filter(ch_d <= 3 & ch_o <= 3) %>%
  select(year, dist, ori_dest)

#add distances (country-country, excluding US)
icio2 <- icio1
for (i in 1:dim(data_services1)[1]){  
  icio2 <- icio2 %>%
    mutate(dist = replace(dist, ori_dest == data_services1$ori_dest[i] & year == data_services1$year[i], data_services1$dist[i]))
}
```

### Adding US-country distances 

```{r distance-1}

#creating weights
us_pop1 <- us_pop %>%
  select(State, Population_2018) %>%
  group_by(State) %>%
  mutate(state_pop = sum(Population_2018)) %>%
  ungroup() %>%
  select(-Population_2018) %>%
  distinct(State, .keep_all=TRUE) %>%
  filter(State != "DC")

data_servicesUS1 <- data_servicesUS %>%
  mutate(ch_d = nchar(as.character(iso_d))) %>%
  filter(ch_d > 3) %>%
  distinct(iso_d, .keep_all = TRUE)
us_pop1$State <- data_servicesUS1$iso_d
us_pop1 <- us_pop1 %>% mutate(tot_pop = sum(state_pop))
us_pop1 <- us_pop1 %>% mutate(pond = (state_pop/tot_pop)) #pond sum==1

# adding US-country distances
data_servicesUS2 <- data_services %>%
  mutate(ch_o = nchar(as.character(iso_o))) %>%
  mutate(ch_d = nchar(as.character(iso_d))) %>%
  filter(ch_d > 3 & ch_o == 3) %>% #only countries in origin and only states in destination 
  mutate(pond = 0)

data_servicesUS3 <- data_servicesUS2
for(i in 1:n_s){
  data_servicesUS3 <- data_servicesUS3 %>%
    mutate(pond = replace(pond, iso_d == us_pop1$State[i], us_pop1$pond[i]))
}

data_servicesUS3 <- data_servicesUS3 %>%
  mutate(component = pond*dist) %>%
  group_by(year, iso_o) %>%
  mutate(dist_US = sum(component)) %>%
  ungroup() %>%
  distinct(year, iso_o, .keep_all=TRUE) %>%
  mutate(iso_d = "USA") %>%
  unite(name1, c("iso_o", "iso_d"), remove = FALSE) %>%
  unite(name2, c("iso_d", "iso_o"), remove = FALSE) %>%
  select(year, name1, name2, dist_US)

#add distances (US-country)
icio3 <- icio2
for (i in 1:dim(data_servicesUS3)[1]){  
  icio3 <- icio3 %>%
    mutate(dist=replace(dist, (ori_dest == data_servicesUS3$name1[i] | ori_dest == data_servicesUS3$name2[i]) & year == data_servicesUS3$year[i], data_servicesUS3$dist_US[i]))
}

```

Specific US_US distance using the following formula:
$$dist\left(US,US\right) = \left(\sum_{r\, \in\, US} \sum_{s \, \in \, US} \left(\dfrac{pop_r}{pop_{US}}\right) \left(\dfrac{pop_s}{pop_{US}}\right) d_{rs}^{-1}\right)^{-1},$$

```{r distance-2}
# US-US distance
USserv <- data_services %>%
  mutate(ch_o = nchar(as.character(iso_o))) %>%
  mutate(ch_d = nchar(as.character(iso_d))) %>%
  filter(ch_d > 3 & ch_o > 3) %>%
  mutate(pond_r = 0, pond_s = 0) %>%
  mutate(dist = (1/dist))
  

for(i in 1:n_s){
  USserv <- USserv %>%
    mutate(pond_r = replace(pond_r, iso_o == us_pop1$State[i], us_pop1$pond[i])) %>%
    mutate(pond_s = replace(pond_s, iso_d == us_pop1$State[i], us_pop1$pond[i]))
}

USserv_final <- USserv %>%
  mutate(component = pond_r*pond_s*dist) %>%
  group_by(year) %>%
  mutate(dist_US_US = sum(component)) %>%
  ungroup() %>%
  distinct(year, .keep_all=TRUE) %>%
  mutate(dist_US_US = 1/dist_US_US) 

USserv_final$iso_o = "USA"
USserv_final$iso_d = "USA"
USserv <- USserv_final %>%
  unite(name1, c("iso_o", "iso_d"), remove = FALSE) 
  
#adding US_US  
for (i in 1:dim(USserv)[1]){  
  icio3 <- icio3 %>%
    mutate(dist = replace(dist, ori_dest == USserv$name1[i] & year == USserv$year[i], USserv$dist_US_US[i]))
}
```

### Final base for services regression

```{r serv-regression}
#taking out countries not of interest
final <- icio3 %>%
  filter(dist != 0) %>%
  dplyr::rename(X = ag_value) %>%
  arrange(year, row_country) %>%
  dplyr::rename(origin = row_country , destination = col_country)
write.csv(final, "1-Intermediate_Processed_Data//Country_Gravity_Services_ICIO.csv")
```

### Stata code for services

```{stata, collectcode=TRUE}
import delimited "1-Intermediate_Processed_Data//Country_Gravity_Services_ICIO.csv"


gen log_X=ln(x)

gen log_dist=ln(dist)

drop ori_dest x dist

*ssc install reghdfe

*ssc install ftools

*reghdfe, compile

egen ori=group(origin)
egen des=group(destin)

reghdfe log_X iota log_dist , absorb(year ori des) vce(robust)
reghdfe log_X iota log_dist , absorb(year#ori year#des) vce(robust)


gen b0 = _b[iota]

gen b1 = _b[log_dist]

gen phi = exp((b0*iota) + (b1*log_dist))
```

## Gravity for agriculture

```{r agric}
#only sector 14: origin
icio1 <- icio %>%
  filter(row_item == n_sec + 2) %>%
  group_by(year, row_country, col_country) %>%
  mutate(ag_value = sum(value)) %>% 
  ungroup() %>% 
  distinct(year, row_country, col_country, .keep_all=TRUE ) %>% 
  select(-col_item, -value, -row_item) %>%
  mutate(iota = 0) %>%
  mutate(iota=replace(iota, row_country == col_country, 1)) %>%
  unite(ori_dest, c("row_country", "col_country"), remove = FALSE) %>%
  mutate(dist = 0)

#preparing distances 
data_agriculture1 <- data_agriculture %>%
  unite(ori_dest, c("iso_o", "iso_d"), remove = FALSE) %>%
  mutate(ch_o = nchar(as.character(iso_o))) %>%
  mutate(ch_d = nchar(as.character(iso_d))) %>%
  filter(ch_d <= 3 & ch_o <= 3) %>%
  select(year, dist, ori_dest)

#add distances (country-country, excluding US)
icio2 <- icio1
for (i in 1:dim(data_agriculture1)[1]){  
  icio2 <- icio2 %>%
    mutate(dist = replace(dist, ori_dest == data_agriculture1$ori_dest[i] & year == data_agriculture1$year[i], data_agriculture1$dist[i]))
}

```

### Adding US-country distances
 
```{r dist-agri}

#creating weights
us_pop1 <- us_pop %>%
  select(State, Population_2018) %>%
  group_by(State) %>%
  mutate(state_pop = sum(Population_2018)) %>%
  ungroup() %>%
  select(-Population_2018) %>%
  distinct(State, .keep_all = TRUE) %>%
  filter(State != "DC")

data_agricultureUS1 <- data_agricultureUS %>%
  mutate(ch_d = nchar(as.character(iso_d))) %>%
  filter(ch_d > 3) %>%
  distinct(iso_d, .keep_all=TRUE)
us_pop1$State=data_agricultureUS1$iso_d
us_pop1 <- us_pop1 %>% mutate(tot_pop = sum(state_pop))
us_pop1 <- us_pop1 %>% mutate(pond = (state_pop/tot_pop)) #pond sum==1

# adding US-country distances
data_agricultureUS2 <- data_agriculture %>%
  mutate(ch_o = nchar(as.character(iso_o))) %>%
  mutate(ch_d = nchar(as.character(iso_d))) %>%
  filter(ch_d > 3 & ch_o == 3) %>% #only countries in origin and only states in destination 
  mutate(pond = 0)

data_agricultureUS3 <- data_agricultureUS2
for(i in 1:n_s){
  data_agricultureUS3 <- data_agricultureUS3 %>%
    mutate(pond=replace(pond, iso_d==us_pop1$State[i], us_pop1$pond[i]))
}

data_agricultureUS3 <- data_agricultureUS3 %>%
  mutate(component = pond*dist) %>%
  group_by(year, iso_o) %>%
  mutate(dist_US = sum(component)) %>%
  ungroup() %>%
  distinct(year, iso_o, .keep_all=TRUE) %>%
  mutate(iso_d="USA") %>%
  unite(name1, c("iso_o", "iso_d"), remove = FALSE) %>%
  unite(name2, c("iso_d", "iso_o"), remove = FALSE) %>%
  select(year, name1, name2, dist_US)

#add distances (US-country)
icio3 <- icio2
for (i in 1:dim(data_agricultureUS3)[1]){  
  icio3 <- icio3 %>%
    mutate(dist = replace(dist, (ori_dest == data_agricultureUS3$name1[i] | ori_dest == data_agricultureUS3$name2[i]) & year == data_agricultureUS3$year[i], data_agricultureUS3$dist_US[i]))
}

# US-US distance
USagri <- data_agriculture %>%
  mutate(ch_o = nchar(as.character(iso_o))) %>%
  mutate(ch_d = nchar(as.character(iso_d))) %>%
  filter(ch_d > 3 & ch_o > 3) %>%
  mutate(pond_r = 0, pond_s = 0) %>%
  mutate(dist = (1/dist))
  

for(i in 1:n_s){
  USagri <- USagri %>%
    mutate(pond_r = replace(pond_r, iso_o == us_pop1$State[i], us_pop1$pond[i])) %>%
    mutate(pond_s = replace(pond_s, iso_d == us_pop1$State[i], us_pop1$pond[i]))
}

USagri_final <- USagri %>%
  mutate(component = pond_r*pond_s*dist) %>%
  group_by(year) %>%
  mutate(dist_US_US = sum(component)) %>%
  ungroup() %>%
  distinct(year, .keep_all = TRUE) %>%
  mutate(dist_US_US = 1/dist_US_US) 

USagri_final$iso_o = "USA"
USagri_final$iso_d = "USA"
USagri <- USagri_final %>%
  unite(name1, c("iso_o", "iso_d"), remove = FALSE) 
  
#adding US_US  
for (i in 1:dim(USagri)[1]){  
  icio3 <- icio3 %>%
    mutate(dist = replace(dist, ori_dest == USagri$name1[i] & year == USagri$year[i], USagri$dist_US_US[i]))
}
```

### Final base for agriculture regression

```{r agri-regression}
#taking out countries not of interest
final <- icio3 %>%
  filter(dist != 0) %>%
  dplyr::rename(X = ag_value) %>%
  arrange(year, row_country) %>%
  dplyr::rename(origin = row_country , destination = col_country)
write.csv(final, "1-Intermediate_Processed_Data//Country_Gravity_Agriculture_ICIO.csv")
```

### Stata code for agriculture

```{stata, collectcode=TRUE}
import delimited "1-Intermediate_Processed_Data//Country_Gravity_Agriculture_ICIO.csv", clear

gen log_X=ln(x)

gen log_dist=ln(dist)

drop ori_dest x dist

egen ori=group(origin)
egen des=group(destin)

reghdfe log_X iota log_dist , absorb(year ori des) vce(robust)
reghdfe log_X iota log_dist , absorb(year#ori year#des) vce(robust)

```

