---
title: "4-Gravity_services"
output:
  html_notebook:
    toc: yes
    df_print: paged
    number_sections: yes    
  pdf_document:
    number_sections: yes
  word_document: default
header-includes:
- \usepackage{amsmath}
- \usepackage{amssymb}
- \usepackage{pifont}
- \usepackage{mathpazo}
---

## General information

1) Constructs the data of Revenue and Expenditure of services by state.

2) This code generates the inputs of the services gravity system and solves it.

## Input files

1. `0-Raw_Data/Fips/state_codes.csv`
2. `0-Raw_Data/Expenditure/SAPCE1_2014_2020_ALL_AREAS.csv`
3. `1-Intermediate_Processed_Data/ICIO_countries.csv`
4. `0-Raw_Data/SAGDP/`
5. `1-Intermediate_Processed_Data/distances.csv`
6. `1-Intermediate_Processed_Data/state_exp_gdp_services.csv`
7. `0-Raw_Data/regions.csv`
8. `0-Raw_Data/sectors_naics.csv`
9. `1-Intermediate_Processed_Data/country_country_step_.csv`
10. `1-Intermediate_Processed_Data/data_services.csv`
11. Coefficients from `Country_Gravity_ICIO`

## Output files

1. `1-Intermediate_Processed_Data/state_exp_gdp_services.csv`
2. `1-Intermediate_Processed_Data/data_services.csv`
3. `1-Intermediate_Processed_Data/matrix_B_.csv`
4. `1-Intermediate_Processed_Data/vector_lambda_.csv`



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
remove(list = ls())
```

```{r packages, warning=FALSE, include=FALSE, message=FALSE, results='hide', cache=FALSE}
# libraries
options( scipen=100, stringsAsFactors=FALSE );
# vector of libraries to be used
libs <- c( 'pracma', 'dplyr', 'tidyr', 'readstata13', 'haven', 'stringr', 'nleqslv', 'gdata', 'Statamarkdown', 'formattable', 'ggplot2', 'viridis', 'RColorBrewer', 'reshape2', 'foreign', 'readxl', 'writexl', 'rjson', 'read_excel', 'Matrix', 'geodist', 'pracma');
# Install libraries in case they are not installed
 for( i in length(libs)){
 if( !(libs[i] %in% installed.packages()) ) install.packages( libs )
 };
# load library
lapply( libs, require, character.only=TRUE );
```


# Task 1

## Importing datasets to R

```{r data, warning=FALSE}
state_code <- read.csv(paste("0-Raw_Data/Fips/state_codes.csv", sep=""), header = TRUE, sep = ",", dec = ".")
bea_raw <- read.csv(paste("0-Raw_Data/Expenditure/SAPCE1_2014_2020_ALL_AREAS.csv", sep=""), header = TRUE, sep = ",", dec = ".")
icio_countries <- read.csv(paste("1-Intermediate_Processed_Data/ICIO_countries.csv", sep=""), header = TRUE, sep = ",", dec = ".")

```

## Services revenue and expenditure

```{r}
## Step 1

# ICIO
icio_countries <- icio_countries %>% filter(sector==13)

# EXPORTS OF US
icio_us_all_exp <- icio_countries %>%
  select(year, importer_country, value_USA) %>%
  group_by(year) %>%
  mutate(val_ = sum(value_USA)) %>% 
  ungroup() %>%
  filter(importer_country == "USA") %>%
  select(year, val_)

# IMPORTS OF US
icio_all_us_imp <- icio_countries %>%
  filter(importer_country == "USA")
icio_all_us_imp$val_ <- rowSums(icio_all_us_imp[, 5:dim(icio_all_us_imp)[2]])  
icio_all_us_imp <- icio_all_us_imp %>% select(year, val_)

#	1. Constructs the data of GDP by state

all_states <- read.csv(paste("0-Raw_Data/SAGDP/SAGDP2N_2015_2018.csv", sep=""), header = TRUE, sep = ",", dec = ".")
all_states <- all_states %>%
  filter(GeoFips <= 90000 & GeoFips != 0) %>%
  filter(GeoName != "District of Columbia") %>%
  left_join(state_code, by=c("GeoName"="state"))
colnames(all_states) <- c("GeoFips", "state", "naics", "Description", "2015", "2016", "2017", "2018", "code")  

#  SERVICE SECTORS   ***  
#  	SECTORS USED AND THEIR MAPPING TO ICIO (indexed by ci)

# 13.	(NAICS XX) Construction (c18); 
# 14.	(NAICS 42-45) Wholesale and Retail Trade (c19ñc21); 
# 15.	(NAICS 481-488) Transport Services (c23ñc26);
# 16.	(NAICS 511ñ518) Information Services (c27); 
# 17.	(NAICS 521ñ525) Finance and Insurance (c28); 
# 18.	(NAICS 531-533)  Real Estate (c29ñc30); 
# 19.	(NAICS 61) Education (c32); 
# 20.	(NAICS 621ñ624) Health Care (c33); 
# 21.	(NAICS 721ñ722) Accommodation and Food Services (c22); 
# 22.	(NAICS 493, 541, 55, 561, 562, 711ñ713, 811-814) Other Services (c34).

all_states <- all_states %>%
  mutate(sector = NA) %>%
  mutate(sector = ifelse(naics == "11" | naics == "34" | naics == "35" | naics == "36" | naics == "45" | naics == "50" | naics == "59" | naics == "68" | naics == "75" | naics == "82",  13, sector)) %>%
  filter(is.na(sector) == FALSE) %>%
  mutate(sector = as.numeric(sector))

#  SETTING FINAL DATA   ***  
# only sending sector
all_states <- melt(all_states, id.vars=c("GeoFips", "state", "naics", "Description", "code", "sector"), variable.name="year", value.name = "revenue")

all_states <- all_states %>%
  mutate(revenue = as.numeric(revenue)) %>%
  mutate(revenue = ifelse(is.na(revenue) == TRUE, 0, revenue)) %>%
  filter(sector >= 13) %>%
  group_by(state, code, year) %>%
  mutate(revenue = sum(revenue)) %>%
  ungroup() %>%
  distinct(state, code, year, .keep_all = TRUE) %>%
  select(state, code, year, revenue) %>%
  arrange(state, code)
all_states$state <- gsub(" ", "", all_states$state, fixed = TRUE)


#	2. Constructs the data of EXPEDITURE by state

# IMPORTING BEA DATA
# KEEPING AND RENAMING RELEVANT VARIABLES
# keeping only states
# making it a number instead of string
bea_raw$GeoFips <- as.numeric(bea_raw$GeoFips)
# dropping regions (leaving only states)
# dropping DC
# keeping relevant variables
bea_raw <- bea_raw %>%
  filter(GeoFips <= 90000 & GeoFips != 0) %>%
  filter(GeoName != "District of Columbia") %>%
  select(-GeoFips) %>%
  dplyr::rename(state = GeoName)
colnames(bea_raw) <- c("state", "line", "description", "2014", "2015", "2016", "2017", "2018", "2019", "2020")

# KEEPING RELEVANT LINES
# keeping only services
bea_raw <- bea_raw %>%
  filter(line == 13) %>% #services
  select(-line, -description) %>%
  arrange(state)
bea_raw$state <- gsub(" ", "", bea_raw$state, fixed = TRUE)
bea_raw <- melt(bea_raw, id.vars=c("state"), variable.name="year", value.name = "expen")


#	3. Merging Expenditure with GDP
bea_raw <- bea_raw %>%
  filter(year!="2014" & year!="2019" & year!="2020") %>%
  left_join(all_states, by=c('year'='year', 'state'='state')) %>%
  arrange(state, code, year, expen, revenue) %>%
  select(state, code, year, expen, revenue)

#	4. Proportionality with ICIO
bea_raw <- bea_raw %>%
  mutate(year = as.numeric(as.character(year))) %>%
  group_by(year) %>%
  mutate(tot_expen = sum(expen)) %>%
  mutate(tot_revenue = sum(revenue)) %>%
  ungroup() %>%
  mutate(sh_expen = expen/tot_expen) %>%
  mutate(sh_revenue = revenue/tot_revenue) %>%
  left_join(icio_all_us_imp, by=c('year'='year')) %>%
  mutate(expen = sh_expen*val_) %>%
  select(-val_) %>%
  left_join(icio_us_all_exp, by=c('year'='year')) %>%
  mutate(revenue = sh_revenue*val_) %>%
  select(-val_) %>%
  select(state, year, revenue, expen) %>%
  arrange(year, state)
  
#check 
check <- bea_raw %>%
  group_by(year) %>%
  mutate(tot_expen = sum(expen)) %>%
  mutate(tot_revenue = sum(revenue)) %>%
  ungroup() %>%
  distinct(year, .keep_all = TRUE) %>%
  select(year, tot_expen, tot_revenue) %>%
  filter(year >= 2015 & year <= 2018)
round(sum(check$tot_expen - icio_all_us_imp$val_), 5) == 0
round(sum(check$tot_revenue - icio_us_all_exp$val_), 5) == 0

write.table(bea_raw, file = paste("1-Intermediate_Processed_Data/state_exp_gdp_services.csv", sep=""), sep = ",", row.names = FALSE)

```


FORMULA 
$d_{ij} = (\sum_{r \in i} \sum_{s \in j} (\dfrac{pop_r}{pop_i}) (\dfrac{pop_s}{pop_j}) d_{rs}^\theta)^{1/ \theta}$

```{r}
# Step 2

state_exp_gdp_services <- read.csv(paste("1-Intermediate_Processed_Data/state_exp_gdp_services.csv", sep=""), header = TRUE, sep = ",", dec = ".")

distances <- read.csv(paste("1-Intermediate_Processed_Data/distances.csv", sep=""), header = TRUE, sep = ",", dec = ".")


#  EXPENDITURE AND GDP AT THE COUNTRY LEVEL   ***  
# Services consumption
expen <- icio_countries %>%
  filter(importer_country != "USA") %>%
  dplyr::rename(country = importer_country)
expen$expen <- rowSums(expen[, 5:dim(expen)[2]])
expen <- expen %>% select(country, year, expen)

# Services production
country_gdp_exp_services <- melt(icio_countries, id.vars=c("year", "importer_country", "sector", "importer"), variable.name="country", value.name = "revenue")

country_gdp_exp_services <- country_gdp_exp_services %>%
  mutate(country = substr(x=country, start=7, stop=9)) %>%
  filter(country != "USA") %>%
  group_by(year, country) %>%
  mutate(revenue = sum(revenue)) %>%
  ungroup() %>%
  distinct(year, country, .keep_all = TRUE) %>%
  select(country, year, revenue) %>%
  left_join(expen, by=c('year'='year', 'country'='country'))

#  EXPENDITURE AND GDP   ***  
# STATES
revenue1 <- state_exp_gdp_services %>% 
  filter(year >= 2015) %>%
  dplyr::rename(iso_o = state, R_i = revenue) %>%
  select(iso_o, R_i, year)
expen1 <- state_exp_gdp_services %>% 
  filter(year >= 2015) %>%
  dplyr::rename(iso_d = state, E_j = expen) %>%
  select(iso_d, E_j, year)
# COUNTRIES
revenue2 <- country_gdp_exp_services %>% 
  dplyr::rename(iso_o = country, R_i = revenue) %>%
  select(iso_o, R_i, year)
expen2 <- country_gdp_exp_services %>% 
  dplyr::rename(iso_d = country, E_j = expen) %>%
  select(iso_d, E_j, year)

revenue <- rbind(revenue1, revenue2)
expen <- rbind(expen1, expen2)

# Merging EXPENDITURE AND GDP
final <- distances %>%
  left_join(revenue, by=c('year'='year', 'iso_o'='iso_o')) %>%
  left_join(expen, by=c('year'='year', 'iso_d'='iso_d')) %>%
  arrange(year, iso_o, iso_d) %>%
  select(year, iso_o, iso_d, dist, R_i, E_j) %>%
  mutate(iso_o =  gsub(" ", "", iso_o, fixed = TRUE)) %>%
  mutate(iso_d =  gsub(" ", "", iso_d, fixed = TRUE))


write.table(final, file = paste("1-Intermediate_Processed_Data/data_services.csv", sep=""), sep = ",", row.names = FALSE)

```


# Task 2 

## Theory

Start with the standard gravity equation: 

$$X_{ij}=\left(\frac{w_{i}\tau_{ij}}{P_{j}}\right)^{-\varepsilon}E_{j},$$

where $P_{j}^{-\varepsilon}=\sum_{i}\left(w_{i}\tau_{ij}\right)^{-\varepsilon}$. But $\sum_{j}X_{ij}=R_{i}$ and hence $\sum_{j}\left(\frac{w_{i}\tau_{ij}}{P_{j}} \right)^{-\varepsilon}E_{j}=R_{i}.$ This implies $w_{i}^{-\varepsilon} \Pi_{i}^{-\varepsilon}=R{i}$, where $\Pi_{i}^{-\varepsilon}= \sum_{j} \tau_{ij}^{ -\varepsilon}P_{j}^{\varepsilon}E_{j}$. Hence we have:

$$X_{ij}=\frac{\tau_{ij}^{-\varepsilon}}{\left(\Pi_{i}P_{j}\right)^{-\varepsilon}}R_{i}E_{j}=\tilde{\tau}_{ij}\tilde{\Pi}_{i}^{-1}\tilde{P}_{j}^{-1}R_{i}E_{j},$$

where $\tilde{P}_{j}\equiv P_{j}^{-\varepsilon}$ and $\tilde{\Pi}_{i}\equiv \Pi_{i}^{-\varepsilon}$, and $\tilde{\tau}_{ij}\equiv\tau_{ij}^{-\varepsilon}$. From ICIO we have $X_{ij}$ for some pairs. Define the set $S_{j}$ such that we know $X_{ij}$ whenever $i\in S_{j}$. Letting 
$$X_{S_{j},j}\equiv\sum_{i\in S_{j}}X_{ij}, \quad\lambda_{j}\equiv1-X_{S_{j},j}/E_{j}, \quad X_{i,S_{i}^{*}}=\sum_{j\in S_{i}^{*}}X_{ij}, \quad\lambda_{i}^{*}\equiv1-X_{i,S_{i}^{*}}/R_{i}, $$
the full system becomes:

$$
\tilde{P}_{j}  =  \sum_{i}\tilde{\tau}_{ij}\tilde{\Pi}_{i}^{-1}R_{i}\quad j\in US 
$$

$$
\tilde{\Pi}_{i}  = \sum_{j}\tilde{\tau}_{ij}\tilde{P}_{j}^{-1}E_{j}\quad i\in US 
$$

$$
\lambda_{j}\tilde{P}_{j} =  \sum_{i\notin S_{j}}\tilde{\tau}_{ij}\tilde{\Pi}_{i}^{-1}R_{i}\quad j\notin US 
$$

$$
\lambda_{i}^{*}\tilde{\Pi}_{i}  =  \sum_{j\notin S_{i}^{*}}\tilde{\tau}_{ij}\tilde{P}_{j}^{-1}E_{j}\quad i\notin US
$$

Define $P_{s}=(\tilde{P}_{1},...,\tilde{P}_{50})'$ for the states,
$P_{c}=(\tilde{P}_{51},...,\tilde{P}_{87})'$ for the countries. Similarly
for $\Pi_{s}$, $\Pi_{c}$,$\lambda_{c}$, and $\lambda_{c}^{*}$.
Define $\lambda=\left(1_{1\times50},\lambda_{c},1_{1\times50},\lambda_{c}^{*}\right)'$. Define $S=\left(P_{s},P_{c},\Pi_{s},\Pi_{c}\right)'$, and with
some abuse of notation $S^{-1}=\left(P_{s}^{-1},P_{c}^{-1},\Pi_{s}^{-1},\Pi_{c}^{-1}\right)'$ with dimensions $(s+c+s+c)\times(1)$. Define $\left(\mathcal{T} R\right)_{ss}$ as follows:

$$ \boldsymbol{\left(\mathcal{T} R\right)_{ss}}=\left(\begin{array}{ccc}
\tilde{\tau}_{s_{1}s_{1}}R_{s_{1}} & \cdots & \tilde{\tau}_{s_{k}s_{1}}R_{s_{k}}\\
\vdots & \ddots & \vdots\\
\tilde{\tau}_{s_{1}s_{k}}R_{s_{1}} & \cdots & \tilde{\tau}_{s_{k}s_{k}}R_{s_{k}}
\end{array}\right), $$

and define , $\left(\mathcal{T} R\right)_{sc} \,, \left(\mathcal{T} R\right)_{cs}$,$\left(\mathcal{T} E\right)_{ss}$,
$\left(\mathcal{T} E\right)_{sc} \,, \left(\mathcal{T} E\right)_{cs}$ analogously. The full system can be written
as:

$$ \lambda\circ S=\left(\begin{array}{cccc}
0 & 0 & \left(\mathcal{T} R\right)_{ss} & \left(\mathcal{T} R\right)_{sc}\\
0 & 0 & \left(\mathcal{T} R\right)_{cs} & 0\\
\left(\mathcal{T} E\right)_{ss} & \left(\mathcal{T} E\right)_{sc} & 0 & 0\\
\left(\mathcal{T} E\right)_{cs} & 0 & 0 & 0
\end{array}\right)\cdot S^{-1}, $$

or as:
$$ \lambda\circ S=B\cdot S^{-1}, $$

where $\circ$ is the element-by-element product and $B$ is the big matrix. Given $\left\{ E_{j}\right\}$  and $\left\{ R_{i}\right\}$  and $\left\{ \tilde{\tau}_{ij}\right\}$ (more on the computation of $\tilde{\tau}_{ij}$ below) ,  we can get $\left\{ \tilde{P}_{j}\right\}$  and $\left\{ \tilde{\Pi}_{i}\right\}$. The solution for $\left\{ \tilde{P}_{j}, \tilde{\Pi}_{i} \right\}$ is unique up to a constant. This indeterminacy requires a normalization. We thus impose $\tilde{P}_{1}=100$. Then we compute $\left\{ X_{ij}\right\}$  from $$X_{ij}=\tilde{\tau}_{ij}\tilde{\Pi}_{i}^{-1}\tilde{P}_{j}^{-1}R_{i}E_{j}.$$

## Importing data to R

```{r, warning=FALSE}
#COEFFIENTS FROM REGRESSIONS
own_dummy <- 4.525
dist_coeff <- -1.03

# number of countries (n_c) and number of states (n_s)
regions <- read.csv("0-Raw_Data/regions.csv")
regions$region <- as.character(regions$region)
c_names <- regions %>% filter(status=="country", region!="USA")
c_names <- c_names$region

s_names <- regions %>% filter(status=="state") 
s_names <- s_names$region

n_s <- length(s_names)
n_c <- length(c_names)

icio_base <- c()
for (yr in 2015:2018) {
  icio <- country.country <- read.csv(paste('1-Intermediate_Processed_Data//country_country_step_', yr, '.csv', sep= ""), header = TRUE, sep = ",", dec = ".")
  icio_base <- rbind(icio_base, icio)
}
data_gravity_services.base <- read.csv( file="1-Intermediate_Processed_Data//data_services.csv", header=TRUE, sep="," )

reclass_sectors <- read.csv("0-Raw_Data/sectors_naics.csv", fileEncoding="UTF-8-BOM")
n_sec <- length(unique(reclass_sectors$final_sector)) -3

#### epsilon to check between calculations
epsilon<-0.01;

#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
```

## Creating and solving gravity system for services

```{r inputs, warning=FALSE}
for (yr in 2015:2018) {
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
# 1. LOAD DATA
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  
icio <- icio_base
## path
## ICIO US
us.us.icio <- icio %>% 
  filter( year == yr, sector == n_sec + 1 ) %>% 
  filter( importer == 'USA' ) %>% 
  select( USA );
us.us.icio <- as.numeric(us.us.icio[1])
# US exports
us.exports.icio <- icio %>% 
  filter( year == yr, sector == n_sec + 1 ) %>% 
  filter( importer != 'USA' ) %>% 
  select( USA ); 
#us.exports.icio <-sum(us.exports.icio)
# US imports
us.imports.icio <- icio %>% 
  filter( year == yr, sector == n_sec + 1 ) %>% 
  filter( importer == 'USA' ) %>% 
  select( -year, -USA, -sector, -importer) 
#us.imports.icio <-sum(us.imports.icio)
## ICIO
icio <- icio_base %>% 
  filter( year == yr, sector == n_sec + 1 ) %>% 
  filter( importer != 'USA' ) %>% 
  select( -year, -USA, -sector, -importer  ) 
## Gravity data csv format with comma as the separator
data_gravity_services <- data_gravity_services.base %>% 
  filter( year == yr ) %>% 
  select( -year ) %>% 
  mutate( iso_o = as.character(iso_o), iso_d = as.character(iso_d) ) %>% 
  mutate( len_o = ifelse(nchar(iso_o)<4, 'COUNTRY', 'STATE'), 
          len_d = ifelse(nchar(iso_d)<4, 'COUNTRY', 'STATE') )
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
# 2. ID for STATES AND COUNTRIES
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
# States
id.states <- data_gravity_services %>% 
  filter( len_o == 'STATE' ) %>% 
  select( iso_o ) %>% 
  distinct( iso_o ) %>% 
  arrange( iso_o ) %>% 
  mutate( id = row_number() ) %>% 
  dplyr::rename( region=iso_o ) %>% 
  select( id, region )
# Countries
id.countries <- data_gravity_services %>% 
  filter( len_o == 'COUNTRY' ) %>% 
  select( iso_o ) %>% 
  distinct( iso_o ) %>% 
  arrange( iso_o ) %>% 
  mutate( id = n_s + row_number() ) %>% 
  dplyr::rename( region = iso_o ) %>% 
  select( id, region )
# Total
id.total <- rbind( id.states, id.countries )
#t(id.total)

#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
# 3. DISTANCES, PRODUCTION AND EXPENDITURE
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
##
## 3.1. DISTANCE MATRIX
##
dist.matrix <- data_gravity_services %>% 
  select( iso_o, iso_d, dist ) %>% 
  left_join( y = id.total, by=c('iso_o'='region') ) %>% 
  dplyr::rename( id_o = id ) %>% 
  left_join( y = id.total, by=c('iso_d'='region') ) %>% 
  dplyr::rename( id_d = id ) %>% 
  select( id_o, id_d, dist ) %>% 
  arrange( id_o, id_d ) %>% 
  spread( key = id_d, value = dist )
dist.matrix <- as.matrix( dist.matrix[,2:(n_c+n_s+1)] )

#Phis (taus)
#we use the coefficients from Country_Gravity_ICIO.
phi <- diag(x = own_dummy, n_c+n_s, n_c+n_s)
phi <- exp(phi)
phi <- phi*dist.matrix^dist_coeff
dist.matrix <- phi

##
## 3.2. GDP VECTOR
##
gdp.vector <- data_gravity_services %>% 
  select( iso_o, R_i ) %>% 
  distinct( iso_o, .keep_all=TRUE ) %>% 
  left_join( y = id.total, by=c('iso_o'='region') ) %>% 
  select( id, R_i ) %>% 
  arrange( id );
gdp.vector <- as.vector( gdp.vector$R_i )
##
## 3.3. EXPENDITURE VECTOR
##
exp.vector <- data_gravity_services %>% 
  select( iso_d, E_j ) %>% 
  distinct( iso_d, .keep_all=TRUE ) %>% 
  left_join( y = id.total, by=c('iso_d'='region') ) %>% 
  select( id, E_j ) %>% 
  arrange( id )
exp.vector <- as.vector( exp.vector$E_j )

#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
# 4. VECTOR LAMBDA AND MATRIX B
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

##
## 4.1 Matrices E and R (repeat gdp and expenditure vectors in ns+nc rows)
##

mat.Y <- matrix( data=rep(gdp.vector, (n_s + n_c)), nrow=(n_s + n_c), byrow=TRUE )
mat.X <- matrix( data=rep(exp.vector, (n_s + n_c)), nrow=(n_s + n_c), byrow=TRUE )

##
## 4.2 Submatrices
##

##phi*Y submatrices
phi.Y.ss <- dist.matrix[(1:n_s), (1:n_s)] * mat.Y[(1:n_s), (1:n_s)]
phi.Y.sc <- dist.matrix[(1:n_s), ((n_s+1):(n_s+n_c))] * mat.Y[(1:n_s), ((n_s+1):(n_s+n_c))]
phi.Y.cs <- dist.matrix[((n_s+1):(n_s+n_c)), (1:n_s)] * mat.Y[((n_s+1):(n_s+n_c)), (1:n_s)]
phi.Y.cc <- dist.matrix[((n_s+1):(n_s+n_c)), ((n_s+1):(n_s+n_c))] * 
  mat.Y[((n_s+1):(n_s+n_c)), ((n_s+1):(n_s+n_c))]
## phi*X submatrices
phi.X.ss <- dist.matrix[(1:n_s), (1:n_s)] * mat.X[(1:n_s), (1:n_s)]
phi.X.sc <- dist.matrix[(1:n_s), ((n_s+1):(n_s+n_c))] * mat.X[(1:n_s), ((n_s+1):(n_s+n_c))]
phi.X.cs <- dist.matrix[((n_s+1):(n_s+n_c)), (1:n_s)] * mat.X[((n_s+1):(n_s+n_c)), (1:n_s)]
phi.X.cc <- dist.matrix[((n_s+1):(n_s+n_c)), ((n_s+1):(n_s+n_c))] * 
  mat.X[((n_s+1):(n_s+n_c)), ((n_s+1):(n_s+n_c))]

##
## 4.3 ICIO matrix (d,o)
##
icio.matrix <- as.matrix( icio )

##
## 4.4 lambda vector
##

## imports and exports
imports.from.countries <- as.matrix( apply(X=icio.matrix, MARGIN=1, FUN=sum) )
exports.to.countries <- as.matrix( apply(X=icio.matrix, MARGIN=2, FUN=sum) )
## shares of imports and exports of countries
X_c <- as.matrix( exp.vector[(n_s + 1):(n_s + n_c)] )
Y_c <- as.matrix( gdp.vector[(n_s + 1):(n_s + n_c)] )
## lambdas
lambda.p <- (X_c-imports.from.countries) / X_c
lambda.pi <- (Y_c- exports.to.countries) / Y_c
## lambda vector
ones.ns <- matrix( data=rep(1, n_s), nrow = n_s, ncol = 1 )
lambda_vector <- rbind( ones.ns, lambda.p, ones.ns, lambda.pi )

##
## 4.5 Big zeros matrices
##
zeros.nc.nc <- matrix( data = rep(0, n_c*n_c), nrow = n_c, ncol = n_c )
zeros.big <- matrix( data = rep(0, (n_s + n_c)*(n_s + n_c)), nrow = (n_s + n_c), ncol = (n_s + n_c) )

##
## 4.6 MATRIX B
##

## Block for system 10
block.10 <- rbind( cbind(phi.Y.ss, phi.Y.sc),
                   cbind(phi.Y.cs, zeros.nc.nc) )
## Block for system 11
block.11 <- rbind( cbind(phi.X.ss, phi.X.sc),
                   cbind(phi.X.cs, zeros.nc.nc) )
## Relevant matrix
BIG.MAT <- rbind( cbind(zeros.big, block.10),
                  cbind(block.11, zeros.big) )

## Saving input matrices for the matlab code
write.csv(BIG.MAT, file= paste('1-Intermediate_Processed_Data//matrix_B_',yr, '.csv', sep= ""), row.names=FALSE, na="" );
write.csv(lambda_vector, file= paste('1-Intermediate_Processed_Data//vector_lambda_',yr, '.csv', sep= ""), row.names=FALSE, na="" );



## Gravity Systems

# Services gravity

BIG.MAT <- as.matrix(read.csv(paste('1-Intermediate_Processed_Data//matrix_B_',yr, '.csv', sep= "")))
lambda_vector <- as.matrix(read.csv(paste('1-Intermediate_Processed_Data//vector_lambda_',yr, '.csv', sep= "")))

# This function solves the equilibrum system to construct the data for
# RUV using a gravity approach 
gravity_system <- function(x){
    Bfirst <- ( ((lambda_vector*x) / (BIG.MAT%*%(x^(-1)))) -1)*100
    Bsecond <- x[1] - 100
    B <- as.matrix(rbind(Bfirst, Bsecond))  
}
start <- 100*matrix(1, 2*(n_c + n_s), 1)
if (yr > 2015){start <- as.matrix(sol)}

sol <- lsqnonlin(gravity_system, start)
sol <- as.matrix(sol$x)

#residuals
residuals <- (lambda_vector*sol) - (BIG.MAT %*% (sol^(-1)))
print(sum(residuals))

write.csv(sol, file= paste('1-Intermediate_Processed_Data//vector_solution_',yr, '.csv', sep= ""), row.names=FALSE, na="" );




## Solution system

#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
###SOLUTION FROM MATLAB
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
SOL <- as.matrix(read.csv(paste('1-Intermediate_Processed_Data//vector_solution_',yr, '.csv', sep="")))
lambda_before <- as.matrix(read.csv(paste('1-Intermediate_Processed_Data//vector_lambda_',yr, '.csv', sep="")))
B_before <-as.matrix(read.csv(paste('1-Intermediate_Processed_Data//matrix_B_',yr, '.csv', sep="")))
checking_sol <- lambda_before*SOL / B_before %*% SOL^{-1}-1;
if(norm(checking_sol, "2")>epsilon) 
  stop(paste("Solution from Matlab does not seem to work for year", yr))

#checking
#checking
value <-  lambda_vector*SOL / BIG.MAT %*% SOL^{-1}-1;
if(norm(value, "2")>epsilon)
  stop(paste("Solution from Matlab does not seem to work for year", yr))

x<- SOL;
##calculating Xij
  #x<-as.matrix( c(normalization,x) )
### Basic vector
P.s <- as.matrix( x[1:n_s] );
P.c <- as.matrix( x[(n_s+1):(n_s+n_c)] );
Pi.s <- as.matrix( x[(n_s+n_c+1):(2*n_s+n_c)] );
Pi.c <- as.matrix( x[(2*n_s+n_c+1):(2*(n_s+n_c))] );

### Inverted basic
P.s.inv <- P.s^(-1); 
P.c.inv <- P.c^(-1);
Pi.s.inv <- Pi.s^(-1);
Pi.c.inv <- Pi.c^(-1);

Pi.sq.mat <- matrix( data=rep(c(Pi.s.inv, Pi.c.inv)*gdp.vector, (n_s + n_c)),
                     nrow=(n_s+n_c), ncol=(n_s+n_c), byrow=TRUE );
P.sq.mat <- matrix( data=rep(c(P.s.inv, P.c.inv)*exp.vector, (n_s + n_c)),
                    nrow=(n_s+n_c), ncol=(n_s+n_c), byrow=FALSE );
#consistent with ICIO notation
Xij.matrix <- Pi.sq.mat * dist.matrix * P.sq.mat;
write.csv(Xij.matrix, file= paste('1-Intermediate_Processed_Data//Xij_matrix_services_', yr, '.csv', sep= ""), row.names=FALSE, na="" );
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
###CHECKS
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

#CHECK: US consumption from US =US consumption from US from ICIO
value1 <- sum(Xij.matrix[(1:n_s), (1:n_s)])
value2 <- icio_base %>%
  filter(sector == n_sec + 1, year == yr, importer == "USA") %>%
  select(-importer, -sector, -year)
value2=value2$USA;
#checking
if(abs(value1/value2-1)>epsilon) 
  stop(paste(" US consumption from US != value from ICIO for", yr, " value=", abs(value1/value2-1)))

#Check: states' exports to countries add to ICIO US exports to countries (sector 13)
value1 <- sum(Xij.matrix[((n_s + 1):(n_s + n_c)), (1:n_s)])
value2 <- icio_base %>%
  filter(year == yr, sector == n_sec + 1, importer != "USA")
value2 <- sum(value2$USA)
if(abs(value1/value2-1)>epsilon) 
  stop(paste("states' exports to countries != ICIO US exports to countries for", yr, " value=", abs(value1/value2-1)))

#Check: states' imports from countries add to ICIO US imports from countries (sector 13)
value1 <- sum(Xij.matrix[(1:n_s), ((n_s + 1):(n_s + n_c))])
value2 <- icio_base %>%
  filter(year == yr, sector == n_sec + 1, importer == "USA") %>%
  select(-year, -importer, -sector, - USA)
value2 <- sum(value2)
if(abs(value1/value2-1)>epsilon) 
  stop(paste("states' imports from countries != ICIO US imports from countries for", yr, " value=", abs(value1/value2-1)))


#Check: states' exports to states add to ICIO US exports to US (sector 13)
value1 <- sum(Xij.matrix[(1:n_s), (1:n_s)])
value2 <- icio_base %>%
  filter(year == yr, sector == n_sec + 1, importer=="USA")
value2 <- value2$USA[1]
if(abs(value1/value2-1)>epsilon) 
  stop(paste("states' exports to states != ICIO US exports to US for", yr, " value=", abs(value1/value2-1)))

}
```

