---
title: "6-Matrices_deficits_VA_shares_states"
output:
  html_notebook:
    toc: yes
    df_print: paged
    number_sections: yes    
  pdf_document:
    number_sections: yes
header-includes:
- \usepackage{amsmath}
- \usepackage{amssymb}
- \usepackage{pifont}
- \usepackage{mathpazo}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
remove(list = ls())
```

```{r warning=FALSE, include=FALSE, message=FALSE, results='hide', cache=FALSE}
# libraries
options( scipen=100, stringsAsFactors=FALSE );
# vector of libraries to be used
libs <- c( 'dplyr', 'tidyr', 'readstata13', 'haven', 'stringr', 'nleqslv', 'gdata', 'Statamarkdown', 'formattable', 'ggplot2', 'viridis', 'RColorBrewer', 'reshape2', 'foreign', 'readxl', 'writexl', 'rjson', 'read_excel', 'Matrix', 'geodist');
# Install libraries in case they are not installed
#for( i in length(libs)){ 
  #if( !(libs[i] %in% installed.packages()) ) install.packages( libs ) 
#};
# Carga de bibliotecas
lapply( libs, require, character.only=TRUE );
```

## General Instructions

1) The goal is to combine the previous steps in order to obtain a new Final Matrix for 2000 through 2007: $X_{ij,k}^{year}$.

2) The purpose of this task is to calculate region-sector level deficits. The deficit for a region $i$ in a sector $k$ would be defined as: 
$$
D_{i,k}=\sum_j X_{ij,k} - \sum_j X_{ji,k}
$$

3) This code calculates the share of value added in gross output for each US state, for each sector.


## Input files

1. `0-Raw_Data/regions.csv`
2. `0-Raw_Data/sectors.csv`
3. `1-Intermediate_Processed_Data//Xij_matrix_services_.csv`
4. `1-Intermediate_Processed_Data//Xij_matrix_agric_.csv`
5. `1-Intermediate_Processed_Data//country_country_step_.csv`
6. `1-Intermediate_Processed_Data//state_cfs_step_.csv`
7. `1-Intermediate_Processed_Data//state_country_step_e_.csv`
8. `1-Intermediate_Processed_Data//state_country_step_y_.csv`
9. `1-Intermediate_Processed_Data//final_matrix_.csv`
10. `1-Intermediate_Processed_Data//value_added_countries.csv`
11. `0-Raw_Data//Labor_shares//gdp_state_.csv`
12. `0-Raw_Data//Labor_shares//taxes_state_.csv`
13. `0-Raw_Data//Labor_shares//subsidies_state_.csv`

## Output files

1. `1-Intermediate_Processed_Data//final_matrix_.csv`
2. `2-Final_Data/bilat_matrix_allyears.xlsx`
3. `1-Intermediate_Processed_Data//deficits_.csv`
4. `2-Final_Data/deficits_allyears.xlsx`
5. `1-Intermediate_Processed_Data//labor_shares_states.csv`
6. `2-Final_Data/va_shares_allyears.xlsx`


# Task 1

## Importing datasets to R
```{r}

#General parameters
regions <- read.csv("0-Raw_Data/regions.csv")
regions$region <- as.character(regions$region)
c_names <- regions %>% filter(status == "country", region != "USA")
c_names <- c_names$region

s_names <- regions %>% filter(status == "state") 
s_names <- s_names$region

regions <- data.frame(rbind(as.matrix(s_names), as.matrix(c_names)))
colnames(regions) <- c("region")

n_s <- length(s_names)
n_c <- length(c_names)

#sectors
reclass_sectors <- read.csv("0-Raw_Data/sectors.csv")
n_sec <- length(unique(reclass_sectors$final_sector)) -3
reclass_sectors <- reclass_sectors %>% 
  select(final_sector, wiod_sector) %>%
  distinct(wiod_sector, .keep_all = TRUE)


services_all <- c()
agriculture_all <- c()
state_state_all <- c()
country_country_all <- c()
state_country_e_all <- c()
state_country_y_all <- c()

for (yr in 2000:2007) {
##Gravity step: services  
services <- read.csv(paste('1-Intermediate_Processed_Data//Xij_matrix_services_', yr, '.csv', sep= ""), header = TRUE, sep = ",", dec = ".")
services <- data.frame(services)
colnames(services) <- regions$region
services <- services %>%
  mutate(year = yr, sector = (n_sec + 1), importer = regions$region) %>%
  select(year, importer, sector, everything())
##Gravity step: agriculture  
agriculture <- read.csv(paste('1-Intermediate_Processed_Data//Xij_matrix_agric_', yr, '.csv', sep= ""), header = TRUE, sep = ",", dec = ".")
agriculture <- data.frame(agriculture)
colnames(agriculture) <- regions$region[1:n_s]
agriculture <- agriculture %>%
  mutate(year = yr, sector = (n_sec + 2), importer = regions$region[1:n_s]) %>%
  select(year, importer, sector, everything())
#country-country
country_country <- read.csv(paste('1-Intermediate_Processed_Data//country_country_step_', yr, '.csv', sep= ""), header = TRUE, sep = ",", dec = ".")
country_country <- country_country %>%
  filter(importer != "USA") %>%
  select(-USA)
#state-state
state_state <- read.csv(paste('1-Intermediate_Processed_Data//state_cfs_step_', yr, '.csv', sep= ""), header = TRUE, sep = ",", dec = ".")
#state-country-state
state_country_e <- read.csv(paste('1-Intermediate_Processed_Data//state_country_step_e_', yr, '.csv', sep= ""), header = TRUE, sep = ",", dec = ".")
state_country_y <- read.csv(paste('1-Intermediate_Processed_Data//state_country_step_y_', yr, '.csv', sep= ""), header = TRUE, sep = ",", dec = ".")
#appending
state_state_all <- rbind(state_state_all, state_state)
country_country_all <- rbind(country_country_all, country_country)
state_country_e_all <- rbind(state_country_e_all, state_country_e)
state_country_y_all <- rbind(state_country_y_all, state_country_y)
services_all <- rbind(services_all, services)
agriculture_all <- rbind(agriculture_all, agriculture)
}
```

## Including services and agriculture sectors

```{r}
#services state-state
serv_state <- services_all%>%
  filter(importer %in% s_names)
serv_state <- serv_state[ ,1:(n_s + 3)]
state_state_all <- rbind(state_state_all, serv_state)
state_state_all <- state_state_all %>%
  arrange(year, sector)
#agricultre state-state
agri_state <- agriculture_all%>%
  filter(importer %in% s_names)
agri_state <- agri_state[ ,1:(n_s + 3)]
state_state_all <- rbind(state_state_all, agri_state)
state_state_all <- state_state_all %>%
  arrange(year, sector)

#imports of states from countries: services
serv_state_country_e <- services_all%>%
  filter(importer %in% s_names)
temp <- serv_state_country_e[ ,1:3]
serv_state_country_e <- serv_state_country_e[ ,(n_s + 3 + 1):dim(serv_state_country_e)[2]]
serv_state_country_e <- cbind(temp, serv_state_country_e)
state_country_e_all <- rbind(state_country_e_all, serv_state_country_e)
state_country_e_all <- state_country_e_all %>%
  arrange(year, sector)
# #imports of states from countries: agriculture
# agri_state_country_e <- agriculture.all%>%
#   filter(importer %in% s.names)
# temp <- agri_state_country_e[ ,1:3]
# agri_state_country_e <- agri_state_country_e[ ,(n.s+3+1):dim(agri_state_country_e)[2]]
# agri_state_country_e <- cbind(temp, agri_state_country_e)
# state.country.e.all <- rbind(state.country.e.all, agri_state_country_e)
# state.country.e.all <- state.country.e.all %>%
#   arrange(year, sector)

#exports of states to countries: services
serv_state_country_y <- services_all%>%
  filter(importer %in% c_names)
serv_state_country_y <- serv_state_country_y[ ,1:(n_s + 3)]
state_country_y_all <- rbind(state_country_y_all, serv_state_country_y)
state_country_y_all <- state_country_y_all %>%
  arrange(year, sector)
# #exports of states to countries: agriculture
# agri_state_country_y <- agriculture.all%>%
#   filter(importer %in% c.names)
# agri_state_country_y <- agri_state_country_y[ ,1:(n.s+3)]
# state.country.y.all <- rbind(state.country.y.all, agri_state_country_y)
# state.country.y.all <- state.country.y.all %>%
#   arrange(year, sector)
```

## Putting all together

Constructing the full matrix $X_{ij,k}^{year}$ by putting together all elements using the steps described in the previous codes. 
```{r}

for (yr in 2000:2007) {

country_country <- subset(country_country_all, year == yr)
country_country <- subset(country_country, select = -year)  
  
state_state <- subset(state_state_all, year == yr)
state_state <- subset(state_state, select = -year)   
  
state_country_y <- subset(state_country_y_all, year == yr)
state_country_y <- subset(state_country_y, select= -year)   
  
state_country_e <- subset(state_country_e_all, year == yr)
state_country_e <- subset(state_country_e, select = -year) 


final_matrix=c()
#loop across sectors
for (k in 1:(n_sec + 2)) {
  state_country_e_0 <- state_country_e %>%
    filter(sector == k) %>%
    select(-importer, -sector)
  state_state_0 <- state_state %>%
    filter(sector == k)
  matrix_top <- cbind(state_state_0, state_country_e_0)
   
  country_country_0 <- country_country %>%
    filter(sector == k) %>%
    select(-importer, -sector)
  state_country_y_0 <- state_country_y %>%
    filter(sector == k)
  matrix_sub=cbind(state_country_y_0, country_country_0)
  
  colnames(matrix_top) <- colnames(matrix_sub) #matching names

  ma02 <- rbind(matrix_top, matrix_sub)
  
  final_matrix <- rbind(final_matrix, ma02 ) #Final 
  
}
final_matrix$sector = 100 + final_matrix$sector;

write.table(final_matrix, file = paste("1-Intermediate_Processed_Data//final_matrix_", yr, ".csv", sep=""), sep = ",", row.names = FALSE)
}

```


## Final checks

The final goal is to check that the final trade matrices are consistent with WIOD database.

```{r}
## Importing datasets to R
#Final WIOD
#WIOD
wiod_base <- c()
for (yr in 2000:2007) {
  wiot <- country.country <- read.csv(paste('1-Intermediate_Processed_Data//country_country_step_', yr, '.csv', sep= ""), header = TRUE, sep = ",", dec = ".")
  wiod_base <- rbind(wiod_base, wiot)
}

cdp_base <- c()
#Final matrices 
for (yr in 2000:2007) {
##Gravity step: services  
matrix <- read.csv(paste("1-Intermediate_Processed_Data//final_matrix_", yr, ".csv", sep=""), header = TRUE, sep = ",", dec = ".")
matrix$year <- yr
matrix <- matrix %>% select(year, everything())

cdp_base <- rbind(cdp_base, matrix)
}

#
epsilon <- 0.0001
```

```{r}
for (yr in 2000:2007) {

#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
#country-country  
wiot <- melt(wiod_base, id.vars=c("year", "importer", "sector"), variable.name="exporter", value.name = "value")
wiot_cc <- wiot %>%
  filter(year == yr, importer %in% c_names, exporter %in% c_names) %>%
  arrange(sector, exporter, importer)

cdp <- melt(cdp_base, id.vars=c("year", "importer", "sector"), variable.name="exporter", value.name = "value")
cdp_cc <- cdp %>%
  filter(year == yr, importer %in% c_names, exporter %in% c_names) %>%
  arrange(sector, exporter, importer)

if(abs(sum(cdp_cc$val)/sum(wiot_cc$value) -1)>epsilon) 
  stop(paste("countries' exports to countries != countries' WIOD exports for year", yr))
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
#state-to-country

wiot_cs <- wiot %>%
  filter(year == yr, importer %in% c_names, exporter %in% c("USA")) %>%
  arrange(sector, exporter, importer)


cdp_cs <- cdp %>%
  filter(year == yr, importer %in% c_names, exporter %in% s_names) %>%
  group_by(sector, importer) %>%
  mutate(val = sum(value)) %>%
  ungroup() %>%
  distinct(year, sector, importer, .keep_all = TRUE) %>%
  arrange(sector, exporter, importer)

if(abs(sum(cdp_cs$val)/sum(wiot_cs$value) -1)>epsilon) 
  stop(paste("states' exports to countries != US WIOD exports to countries for year", yr))
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
# country-to-state

wiot_sc <- wiot %>%
  filter(year == yr, importer %in% c("USA"), exporter %in% c_names) %>%
  arrange(sector, exporter, importer)


cdp_sc <- cdp %>%
  filter(year == yr, importer %in% s_names, exporter %in% c_names) %>%
  group_by(sector, exporter) %>%
  mutate(val = sum(value)) %>%
  ungroup() %>%
  distinct(year, sector, exporter, .keep_all = TRUE) %>%
  arrange(sector, exporter, importer)

if(abs(sum(cdp_sc$val)/sum(wiot_sc$value) -1)>epsilon) 
  stop(paste("countries' exports to states != countries' WIOD exports to US for year", yr))
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
#state-state

wiot_ss <- wiot %>%
  filter(year == yr, importer %in% c("USA"), exporter %in% c("USA")) %>%
  arrange(sector, exporter, importer)


cdp_ss <- cdp %>%
  filter(year == yr, importer %in% s_names, exporter %in% s_names) %>%
  group_by(sector) %>%
  mutate(val = sum(value)) %>%
  ungroup() %>%
  distinct(year, sector, .keep_all = TRUE) %>%
  arrange(sector, exporter, importer)

if(abs(sum(cdp_ss$val)/sum(wiot_ss$value) -1)>epsilon) 
  stop(paste("states' exports to states != US WIOD exports to US for year", yr))

#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


}
```


## Exporting all bilateral flows matrices in a unique file 
```{stata, collectcode=TRUE}

forvalue i=2000/2007{

import delimited "1-Intermediate_Processed_Data/final_matrix_`i'.csv", clear 
         
export excel using "2-Final_Data/bilat_matrix_allyears.xlsx" ///
	, sheet("year`i'") firstrow(variables) sheetmodify

}
```


# Task 2

$$
D_{i,k}=\sum_j X_{ij,k} - \sum_j X_{ji,k}
$$

## Calculating Deficits

```{r}
base <- cdp_base

tot_imp <- rowSums(base[, 4:dim(base)[2]])

base1 <- melt(base, id.vars=c("year", "importer", "sector"), variable.name="exporter", value.name = "value")
base1 <- spread(base1, importer, value, fill = NA, convert = TRUE,  drop = TRUE, sep = NULL)
tot_exp <- rowSums(base1[, 4:dim(base1)[2]])

#deficit
deficit <- data.frame(tot_exp - tot_imp)
deficit <- cbind(base[, 1:3], deficit)
colnames(deficit) <- c("year", "importer", "sector", "deficit")

for (yr in 2000:2007) {
  final <- deficit %>% filter(year == yr)
  write.table(final, file = paste("1-Intermediate_Processed_Data//deficits_", yr, ".csv", sep=""), sep = ",", row.names = FALSE)
}

```


## Exporting all aggregate deficits in a unique sheet
```{r}

deficit$pos <- as.numeric(row.names(deficit))

deficit <- deficit %>%
  dplyr::rename(region = importer) %>% 
  group_by(region) %>%
  mutate(pos = min(pos)) %>%
  ungroup() %>%
  group_by(year, region) %>% 
  mutate(deficit = sum(deficit)) %>%
  ungroup() %>%
  distinct(year, region, .keep_all = TRUE) %>%
  select(-sector) 

deficit <-spread(deficit, year, deficit, fill = NA, convert = TRUE,  drop = TRUE, sep = NULL)
deficit <- deficit %>% 
  arrange(pos) %>%
  select(-pos)
colnames(deficit) <- c("region", "deficit_2000", "deficit_2001", "deficit_2002", "deficit_2003", "deficit_2004", "deficit_2005", "deficit_2006", "deficit_2007")


write_xlsx(deficit, paste("2-Final_Data/deficits_allyears.xlsx"), col_names = TRUE)

 
```

# Task 3


We need to compute the share of value added in gross output the 50 for US states.  We can obtain data on sectoral and regional value added for the US from the Bureau of Economic Analysis (BEA). The final output would a matrix 50 $\times$ 13 with the shares of value added in gross output for each state, for each sector. Let's focus on the year 2000.

- Value added for each of the 50 U.S. states and 13 sectors can be obtained from the Bureau of Economic Analysis (BEA) by subtracting taxes and subsidies from GDP data 
- Gross outputs for each region $i$ is just $Y_{i,k}= \sum_j X_{ij,k}$ 
- In a few cases, gross output might be smaller than value added (probably due to some small discrepancies between trade and production data, etc). In such cases, just constrain value added to be equal to gross output. 

## Importing datasets to R
```{r}

base <- c()
VA_US <- c()
GDP <- c()
TAXES <- c()
SUBSIDIES <- c()
for (yr in 2000:2007) {
#Final matrices
m <- read.csv(paste("1-Intermediate_Processed_Data//final_matrix_", yr, ".csv", sep=""), header = TRUE, sep = ",", dec = ".")
m$year <- yr
m <- m %>% select(year, everything())
base <- rbind(base, m)

#Value added USA
m1 <- read.csv(paste("1-Intermediate_Processed_Data//value_added_countries", yr, ".csv", sep=""), header = TRUE, sep = ",", dec = ".")
VA_US <- rbind(VA_US, m1)

#State GDP
gdp <- read.csv(paste("0-Raw_Data//Labor_shares//gdp_state_", yr, ".csv", sep=""), header = FALSE, sep = ",", dec = ".")
gdp <- gdp[6:(dim(gdp)[1]-7),1:dim(gdp)[2]]
gdp <- data.frame(gdp)
gdp$year <- yr
gdp <- gdp %>% select(year, everything())
GDP <- rbind(GDP, gdp)  
#These databases were downloaded from BEA, Regional Data section; specifically, the base is named SAGDP2. After choosing this information, select the NAICS option, All areas, All statistics in table, and each year from 2000 to 2007, separately (to download the next year just go to the previous time section window and change the selected year). Download each base in a csv file.
#Each of this annual base needs an additional column that was added manually in excel. The column's name is "code" and contains the sectors' code in expanded form (1-24). If the base needs to be downloaded again in the future all that has to be done to the new one is copy and paste this "code" column (as it is) to the new base.

#State taxes
taxes <- read.csv(paste("0-Raw_Data//Labor_shares//taxes_state_", yr, ".csv", sep=""), header = FALSE, sep = ",", dec = ".")
taxes <- taxes[6:(dim(taxes)[1]-6),1:dim(taxes)[2]]
taxes <- data.frame(taxes)
taxes$year <- yr
taxes <- taxes %>% select(year, everything())
TAXES <- rbind(TAXES, taxes)  
#These databases were downloaded from BEA, Regional Data section; specifically, the base is named SAGDP6. After choosing this information, select the NAICS option, All areas, All statistics in table, and each year from 2000 to 2007, separately (to download the next year just go to the previous time section window and change the selected year). Download each base in a csv file.

#State subsidies
subsidies <- read.csv(paste("0-Raw_Data//Labor_shares//subsidies_state_", yr, ".csv", sep=""), header = FALSE, sep = ",", dec = ".")
subsidies <- subsidies[6:(dim(subsidies)[1]-6),1:dim(subsidies)[2]]
subsidies <- data.frame(subsidies)
subsidies$year <- yr
subsidies <- subsidies %>% select(year, everything())
SUBSIDIES <- rbind(SUBSIDIES, subsidies)  
#These databases were downloaded from BEA, Regional Data section; specifically, the base is named SAGDP5. After choosing this information, select the NAICS option, All areas, All statistics in table, and each year from 2000 to 2007, separately (to download the next year just go to the previous time section window and change the selected year). Download each base in a csv file.
}

VA_US <- VA_US %>% filter(region == "USA")
#taking out spaces between states' names: the option sub eliminates only the first space it finds 
GDP$V2 <- sub(' ', '', GDP$V2)
TAXES$V2 <- sub(' ', '', TAXES$V2)
SUBSIDIES$V2 <- sub(' ', '', SUBSIDIES$V2)

```


Gross output for each region, for each sector
$Y_{i,k}= \sum_j X_{ij,k} \quad for \quad all \quad j$
```{r}
base <- melt(base, id.vars=c("year", "importer", "sector"), variable.name="exporter", value.name = "value")
base <- spread(base, importer, value, fill = NA, convert = TRUE,  drop = TRUE, sep = NULL)
Y <- data.frame(rowSums(base[, 4:dim(base)[2]]))
Y <- cbind(base[, 1:3], Y)
colnames(Y) <- c("year", "sector", "region", "value")
Y <- Y %>% filter(region %in% s_names)
Y <- spread(Y, sector, value, fill = NA, convert = TRUE,  drop = TRUE, sep = NULL)

```


## Constructing Value-Added ("Labor Shares")

Calculating value added
$VA_{i,k}=GDP_{i,k}-tax_{i,k}-subsidy_{i,k}, \quad where \quad subsidy<0$
```{r}
value <- GDP
value$tax <- TAXES[, 6]
value$sub <- SUBSIDIES[, 6]
colnames(value) <- c("year", "id" ,"state", "line", "description", "gdp", "sector", "tax", "sub")
value$id <- as.numeric(value$id)
value$sector <- as.numeric(value$sector)
value$gdp <- as.numeric(value$gdp) #some NA generated because base is not clean yet (cleaned in next steps)
value$tax <- as.numeric(value$tax)
value$sub <- as.numeric(value$sub)

VA_s <- value %>%
  mutate(gdp = ifelse(is.na(gdp) == TRUE, 0, gdp)) %>%
  mutate(tax = ifelse(is.na(tax) == TRUE, 0, tax)) %>%
  mutate(sub = ifelse(is.na(sub) == TRUE, 0, sub)) %>%
  filter(sector != 0) %>%
  filter(state != "Districtof Columbia" & state != "UnitedStates *") %>%
  filter(id <= 56000) %>% #id's of interest
  left_join(reclass_sectors, by=c('sector'='wiod_sector')) %>%
  mutate(sector = final_sector) %>%
  select(-final_sector) %>%
  mutate(sector = sector + 100) %>%
  select(-id, -line, -description) %>%
  mutate(amount = (gdp) - (tax/1000) - (sub/1000)) %>%
  group_by(year, state, sector) %>%
  mutate(value_added = sum(as.numeric(as.character(amount)))) %>%
  ungroup() %>%
  distinct(year, state, sector, .keep_all=TRUE ) %>%
  select(-amount) %>%
  arrange(year, state, sector) %>%
  select(-gdp, -tax, -sub) %>%
  dplyr::rename(region = state)


VA_US <- melt(VA_US, id.vars=c("year", "region"), variable.name= "sector", value.name = "value")
VA_US <- VA_US %>% arrange(year)
VA_US <-as.matrix(VA_US[, 4])

#proportionality for states' VA
VA_s<-spread(VA_s, region, value_added, fill = NA, convert = TRUE,  drop = TRUE, sep = NULL)
VA_s_1 <- VA_s[ , (3:dim(VA_s)[2])] /rowSums(VA_s[ , (3:dim(VA_s)[2])])
# check <- rowSums(VA_s_1)
# print(check)
value1 <- VA_s_1*VA_US
# check1 <- VA_US - rowSums(value1)
# print(check1)
VA_s <- cbind(VA_s[, 1:2], value1)
VA_s <- melt(VA_s, id.vars=c("year", "sector"), variable.name="region", value.name = "value_added")
VA_s$region <- as.character(VA_s$region)

VA_s <- spread(VA_s, sector, value_added, fill = NA, convert = TRUE,  drop = TRUE, sep = NULL)

#some NA generated at the beginning of the chunk because base was not clean yet
```

share of value added in gross output
$Labor share = \frac{VA_{i,k}}{Y_{i,k}}$
```{r}
labor_shares <- VA_s[, 3:dim(VA_s)[2]] / Y[, 3:dim(Y)[2]]
labor_shares[labor_shares>1] <- 1
labor_shares[labor_shares<(-1)] <- (-1)
labor_shares <- cbind(VA_s[, 1:2], labor_shares)

for (yr in 2000:2007) {
  final <- labor_shares %>%
    filter(year == yr)
  write.table(final, file = paste("1-Intermediate_Processed_Data//labor_shares_states", yr, ".csv", sep=""), sep = ",", row.names = FALSE)
}


```


## Exporting all VA shares in a unique sheet
```{r}
all <- c()
for (yr in 2000:2007) {

labor_states <- read.csv(paste("1-Intermediate_Processed_Data//labor_shares_states", yr, ".csv", sep=""))
labor_countries <- read.csv(paste("1-Intermediate_Processed_Data//labor_shares_countries", yr, ".csv", sep=""))
labor_countries <- labor_countries %>% filter(region != "USA")
colnames(labor_states) <- c("year", "region", "sector_101", "sector_102", "sector_103", "sector_104", "sector_105", "sector_106", "sector_107", "sector_108", "sector_109", "sector_110", "sector_111", "sector_112", "sector_113", "sector_114")
colnames(labor_countries) <- c("year", "region", "sector_101", "sector_102", "sector_103", "sector_104", "sector_105", "sector_106", "sector_107", "sector_108", "sector_109", "sector_110", "sector_111", "sector_112", "sector_113", "sector_114")

all_shares <- rbind(labor_states, labor_countries)
all <- rbind(all, all_shares)

}

write_xlsx(all, paste("2-Final_Data/va_shares_allyears.xlsx"), col_names = TRUE)

```
